### 2008-01-25 docs/motivation/motivation.tex
Initial creation of the motivating paper.

\documentclass[conference]{IEEEtran}

\usepackage[noadjust]{cite}

%\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
\usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
%\fi

\usepackage{url}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{threeparttable}
\usepackage{multirow}

\begin{document}

\title{Using Altruistic Peers to Supplement Downloading Bandwidth}
\author{\IEEEauthorblockN{Cameron Dale}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: camerond@cs.sfu.ca}}

\maketitle

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background}
\label{background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{conclusions}

\bibliographystyle{IEEEtran}
\bibliography{./IEEEabrv,./all}

\end{document}

### 2008-02-09 docs/motivation/motivation.tex
Lots of work on the motivating paper, and a new figure.

@@
 
-%\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-%\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
 \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-%\fi
 
@@
 
-\title{Using Altruistic Peers to Supplement Downloading Bandwidth}
+\title{Using Altruistic Peers to Help With Free Web Downloads}
 \author{\IEEEauthorblockN{Cameron Dale}
@@
 \begin{abstract}
+A large amount of free content is available over the Internet from
+many different sources. Most of this content uses the traditional
+client-server model of requests from users. However, there is an
+opportunity to use peer-to-peer techniques to reduce the cost of
+much of this distribution, especially due to the altruistic nature
+of many of these users. We present a general technique for
+satisfying the needs of this P2P distribution, which is applicable
+to many of these distributors systems. Our method makes use of a DHT
+for storing the location of peers, using the cryptographic hash of
+the file as a key. We then go further and implement a solution for
+the distribution of Debian software packages.
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+There are a large number of free content distributors with using
+package distribution systems over the Internet to distribute content
+to their users. These distributors have developed many different
+methods for this distribution, but they mostly use a client-server
+model to satisfy user requests. The popularity and number of users
+results in a large number of requests, which usually requires a
+network of mirrors to handle. Due to the free nature of this
+content, many users are willing and able to contribute upload
+bandwidth to this distribution, but have no current way to do this.
+
+We present a new peer-to-peer distribution model to meet these
+demands. It is based on many previous implementations of successful
+peer-to-peer protocols, including distributed hash tables (DHT) and
+BitTorrent. The model relies on the pre-existence of cryptographic
+hashes of the content, which should uniquely identify it for a
+request from other peers. If the peer-to-peer download fails, then
+the original request to the server is used as a fallback to prevent
+any dissatisfaction from mirrors. The peer can then share this new
+content with others.
+
+First, we examine the opportunity that is available for many of
+these free content distributors. We present an overview of a system
+that would efficiently satisfy the demands of a large number of
+users, and should substantially reduce the currently substantial
+bandwidth requirements of hosting this content. We then present an
+example implementation based on the Debian package distribution
+system. This implementation will be used by a large number of users,
+and serves as an example for other free content distributors of the
+opportunity that can be met with such a system.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+There are many peer-to-peer implementations available today, but
+none is very well suited to this specific problem.
 
-\section{Background}
-\label{background}
+Many distributors make their content available in some form using
+BitTorrent \cite{COHEN03}, e.g. for the distribution of CD
+images. This is not an ideal situation though, as it requires the
+peers to download large amounts of content that they are not
+interested in, and prevents them from updating to new content
+without downloading another image containing a lot of the same
+content they already have. There have also been other attempted
+implementations, usually based on some form of BitTorrent, to
+address this problem. Unfortunately, BitTorrent is not ideally
+suited to an application such as this one, for the following
+reasons:
+\begin{itemize}
+ \item The packages are too small and there are too many to create
+       individual torrents for each one.
+ \item All the packages together are too large to track efficiently
+       as a single torrent.
+ \item BitTorrent's piece sizes are bigger than many of the
+       packages, which wastes bandwidth downloading parts of other
+       packages.
+ \item A small number of the packages can be updated every day,
+       which may require a new torrent, therefore splitting the
+       download population.
+ \item Incentives to share (upload) are no longer needed, as the
+       content is freely available for anyone to download without
+       sharing (seeds are also not needed).
+\end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Some other implementations have used BitTorrent almost unmodified,
+while others have only looked at using DHTs to replace the tracker
+in a BitTorrent system. apt-torrent \cite{apttorrent} creates
+torrents for some of the larger content available, but it ignores
+the smaller content, which is often the more requested content.
+DebTorrent \cite{debtorrent} makes widespread modifications to a
+traditional BitTorrent client, but also requires some changes to the
+distribution system to support it, and separates peers into groups
+based on their characteristics which limits the effectiveness of the
+sharing. Kenosis \cite{kenosis} is a P2P Remote Procedure Call
+client also based on the Kademlia DHT, but it is meant as a P2P
+primitive system on which other tools can be built, and so it has no
+file sharing functionality at all. Many have used a DHT as a drop-in
+replacement for the tracker in a BitTorrent system
+\cite{bittorrent-dht, azureus-dht}, but such systems only use the
+DHT to find peers for the same torrent, so the file sharing uses
+traditional BitTorrent and so is not ideal for the reasons listed
+above.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Situation}
+\label{situation}
+
+There are a large number of groups using the Internet to distribute
+their free content. This content is usually available from a free
+download site, which usually requires a network of mirrors to
+support the large number of requests. This content almost always
+supports some kind of file verification, usually cryptographic
+hashes, to verify the completed downloads accuracy or authenticity.
+
+\subsection{Examples}
+\label{examples}
+
+Most Linux distributions use a software package management system
+that fetches packages to be installed from a network of mirrors. The
+Debian project \cite{debian} (and other Debian-based distributions)
+uses the \texttt{apt} program, which downloads Debian packages in
+the \texttt{.deb} format from one of many mirrors. The program will
+first download index files that contain a listing of which packages
+are available, as well as important information such as their size,
+location, and a hash of their content. Red Hat's Fedora project
+\cite{fedora} (and other RPM-based distributions) use the
+\texttt{yum} program to obtain RPMs, and Gentoo \cite{gentoo} uses
+\texttt{portage} in a similar way.
+
+Other software vendors also use a similar system. CPAN \cite{cpan}
+distrbutes files containing software packages for the PERL
+programming language. CYGWIN \cite{cygwin} provides many of the
+standard Unix/Linux tools in a Windows environment, also using a
+package management tool that requests packages from websites. There
+are two software distribution systems for Mac OSX, fink and
+MacPorts, that also retrieve packages in this way.
+
+Also, some systems use direct web downloads, but with a hash
+verification file also available for download next to the desired
+file. These hash files usually have the same file name, but with an
+added extension identifying the hash used (e.g. \texttt{.md5} for
+the MD5 hash).
+
+Finally, there are other programs that use cryptographic hashing to
+identify files. Git is a version control system in which all files,
+commits, and tags, are identified by their SHA1 hash. These hashes
+are used to verify the origin of these items, but are also used when
+clients update their local files with new remote information.
+
+\subsection{Similarities}
+\label{similarities}
+
+The important things to note for each of the systems mentioned in
+section \ref{examples}, is that they all have the following in
+common:
+\begin{itemize}
+ \item The content is avaliable for anyone to freely download.
+ \item The content is divided into packages, each of which contains
+       a small independent part of all the content.
+ \item No peers are interested in downloading all of the content
+       available.
+ \item Hashes of the content and its size are available before the
+       download is attempted.
+ \item Requests for the content are sent by a tool, not directly by
+       the user, though the tool is responding to requests from the user.
+\end{itemize}
+
+We also expect that there are a number of users of these systems
+that are motivated by altruism to want to help out with this
+distribution. This is common in these systems due to the free nature
+of the content being delivered, which encourages some to want to
+help out in some way. A number of the systems are also used by
+groups that are staffed mostly, or sometimes completely, by
+volunteers. This also encourages users to want to \emph{give back}
+to the volunteer community that has created the content they are
+using.
+
+\subsection{Differences}
+\label{differences}
+
+Although at first it might seem that a single all-reaching solution
+is possible for this situation, there are some differences in each
+system that require independent solutions. The systems all use
+different tools for their distribution, so any implementation would
+have to be specific to the tool it is to integrate with. In
+particular, how to obtain requests from the tool or user, and how to
+find a hash that corresponds to the file being requested, is very
+specific to each system.
+
+Also, there may be no benefit in creating a single large solution to
+integrate all these problems. For one, though they sometimes
+distribute nearly identical content (e.g. the same software
+available in multiple Linux distributions), it is not exactly
+identical and has usually been tailored to the system. The small
+differences will make it impossible to distribute similar content
+across systems. And, although peer-to-peer systems scale very well
+with the number of peers in the system, there is some overhead
+involved. So, having a much larger system of peers would mean that
+requests could take longer to complete.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Opportunity}
+\label{opportunity}
+
+The situation described in section \ref{situation} presents a clear
+opportunity to use some form of peer-to-peer file sharing protocols.
+This sparse interest in a large number of packages, and constant
+updating, is well suited to the functionality provided by a
+Distributed Hash Table (DHT). DHTs require unique keys to store and
+retrieve strings of data, for which the cryptographic hashes used by
+these package management systems are perfect for. The stored and
+retrieved strings can then be pointers to the peers that have the
+content package that hashes to that key. A downloading peer can then
+lookup the package hash in the DHT, and then, if it is found,
+download the file from those peers and verify the content. If the
+package hash can not be found in the DHT, the peer will fallback to
+downloading from the original content location (i.e. the network of
+mirrors), and once complete will add a new entry to the DHT
+indicating that it has the package.
+
+\subsection{Implementation Options}
+\label{imp_options}
+
+There are several ways to implement the desired P2P functionality
+into the existing package management software. The functionality can
+be directly integrated into the software, though this can be
+difficult as the DHT should be running at all times, both for
+efficient lookups and to support uploading of already downloaded
+content. Many of the package management software implmentations use
+HTTP requests to download the files, which makes it possible to
+implement the P2P aspect as a standard HTTP caching proxy, which
+will get uncached requests first from the P2P system, and then
+fallback to the normal HTTP request from a server. For methods that
+don't use HTTP requests, other types of proxies may also be
+possible.
+
+\subsection{Downloading From Peers}
+\label{downloading}
+
+Downloading a file efficiently from a number of peers is where
+BitTorrent shines as a peer-to-peer application. Its method of
+breaking up larger files into sub-pieces, each with its own hash,
+makes it very easy to parallelize the downloading process and
+maximize the download speed. For very small packages (i.e. less than
+the sub-piece size), this parallel downloading is not necessary, or
+even desirable. However, this method can still be used in
+conjunction with the DHT, for the larger packages that are
+available.
+
+In addition to storing the file download location (which would still
+be used for small files), a peer will store a \emph{torrent string}
+containing the peer's hashes of the individual pieces for the larger
+files. These piece hashes could be compared ahead of time to
+determine which peers have the same piece hashes (they all should),
+and then used during the download to verify the downloaded pieces.
+
+For very large files (about 5 or more pieces), the torrent strings
+are too long to store in the DHT (a single UDP packet should be less
+than 1472 bytes to avoid fragmentation). Instead, the peers will
+store the torrent string for large files separately in the DHT, and
+only contain a reference to it in their stored value for the hash of
+the file. The reference would be a hash of the torrent string. If
+the torrent string is short enough to store in the DHT (i.e. less
+than 1472 bytes, or about 70 pieces for the SHA1 hash), then a
+lookup of that hash in the DHT would give the torrent-like string.
+Otherwise, a request to the peer for the hash (just like files are
+downloaded), should return the torrent string.
+
+% This can cause a problem with hash checking the returned data, as
+% hashes for the pieces are not known. Any file that fails a hash
+% check should be downloaded again, with each piece being downloaded
+% from different peers than it was previously. The peers are shifted
+% by 1, so that if a peers previously downloaded piece i, it now
+% downloads piece i+1, and the first piece is downloaded by the
+% previous downloader of the last piece, or preferably a previously
+% unused peer. As each piece is downloaded the running hash of the
+% file should be checked to determine the place at which the file
+% differs from the previous download.
+% 
+% If the hash check then passes, then the peer who originally provided
+% the bad piece can be assessed blame for the error. Otherwise, the
+% peer who originally provided the piece is probably at fault, since
+% he is now providing a later piece. This doesn't work if the
+% differing piece is the first piece, in which case it is downloaded
+% from a 3rd peer, with consensus revealing the misbehaving peer.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Sample Implementation}
+\label{implementation}
+
+A sample implementation has been created that functions as described
+in section \ref{opportunity}. This software, called
+\texttt{apt-dht}, interacts with the \texttt{apt} tool found in most
+Debian-based Linux distributions. Our tool uses SHA1 hashes to
+verify most downloaded files, including the large index files that
+contain the hashes of the individual packages. We chose this
+distribution system as it is familiar to us, and there are
+interesting statistics available for analyzing the popularity of the
+software packages \cite{popcon}.
+
+Since all requests from apt are in the form of HTTP requests to a
+server, the implementation takes the form of a caching HTTP proxy.
+Pointing a standard apt implementation to use the proxy is as simple
+as prepending the proxy location and port to the front of the mirror
+name (i.e. ``localhost:9977/mirrorname.debian.org/\ldots'').
+
+The DHT is based on Khashmir, which is an implementation of the
+Kademlia DHT using methods familiar to BitTorrent developers. The
+communication is all handled by UDP messages, and RPC requests are
+bencoded similar to torrent files. Khashmir uses the high-level
+Twisted event-driven networking engine, so Twisted is also used for
+all other networking needs.
+
+The torrent strings stored in the DHT are all bencoded dictionaries
+containing similar information to what is in a torrent file. This
+includes the piece size used, the length of the piece hash, and of
+course the hashes of all the sub-pieces of the content.
+
+Downloading is accomplished by sending simple HTTP requests to the
+peer's identified by lookups in the DHT to have the desired file.
+The HTTP server used for the proxy also doubles as the server
+listeneing for requests for downloads from other peers. All peers
+support HTTP/1.1, both in the server and the client, which allows
+for pipelining of multiple requests to a peer, and the requesting of
+smaller pieces of a large file using the Range request header.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{apt_dht_simulation-size_CDF.eps}
+\caption{The CDF of the size of packages in a Debian system, both
+for the actual size and adjusted size based on the popularity of
+the package.}
+\label{size_CDF}
+\end{figure}
+
+Figure \ref{size_CDF} shows the package size of the 22,298 packages
+available in Debian in January 2008. We can see that most of the
+packages are quite small, and so most will therefore not require
+sub-piece information to download. We have chosen a maximum piece
+size of 512 kB, which means that 17,515 (78\%) of the packages will
+not require this information. There are 3054 packages that will
+require 2 to 4 pieces, for which the torrent string can be stored
+directly with the package hash in the DHT. There are 1667 packages
+that will require a separate lookup in the DHT for the longer
+torrent string as they require 5 to 70 pieces. Finally, there are
+only 62 packages that require more than 70 pieces, and so will
+require a separate request to the user for the torrent string.
+
+\subsection{Current Status}
+\label{status}
+
+Though functional and useful, the current implementation is not
+complete, and is missing some of the downloading details necessary
+to make it complete. It operates as a caching HTTP proxy, serving
+requests from apt and other peers. Index files are identified by the
+current implementation, and parsed to find hashed of files for DHT
+lookups if they are later requested. Files are downloaded from
+peers, and those not available in peers are downloaded correctly
+from the mirror. Any downloaded files are hash checked to verify
+them, and then added to the DHT.
+
+However, the current implementation does not use the sub-piece
+hashing described in section \ref{downloading}, or any parallelizing
+of downloads. Instead, it currently chooses a peer at random from
+the list of possible peers, and downloads the entire file from that
+peer, regardless of the size of the file. This needs to change if
+both a) the file is large (more than 512 KB), and b) there are
+multiple peers with the file. The storage and retrieval of sub-piece
+hashes has not yet been finalized though, and until it is this final
+step is not possible.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \label{conclusions}
+
+We have designed a generally applicable peer-to-peer content
+distribution model to be used by many of the free content
+distributors operating today. It makes use of already existing
+features of most package management systems, to create a
+peer-to-peer distribution that should substantially reduce the costs
+of hosting the content.
+
+We have also implemented our design in software to be used in
+conjuction with Debian-based distribution of Linux software
+packages. It will soone be in use by many users of the Debian
+project's distribution, and so serves as an example of the
+possibilities that exist.
 

### 2008-02-13 docs/motivation/motivation.tex
Finalized initial motivation paper.

@@
 A large amount of free content is available over the Internet from
-many different sources. Most of this content uses the traditional
-client-server model of requests from users. However, there is an
+many different distributors. Most of this content uses the traditional
+client-server model to handle requests from users. However, there is an
 opportunity to use peer-to-peer techniques to reduce the cost of
@@
 
-There are a large number of free content distributors with using
+There are a large number of free content distributors using
 package distribution systems over the Internet to distribute content
 to their users. These distributors have developed many different
-methods for this distribution, but they mostly use a client-server
+methods for this distribution, but they almost exclusively use a client-server
 model to satisfy user requests. The popularity and number of users
@@
 demands. It is based on many previous implementations of successful
-peer-to-peer protocols, including distributed hash tables (DHT) and
+peer-to-peer protocols, especially distributed hash tables (DHT) and
 BitTorrent. The model relies on the pre-existence of cryptographic
@@
 the original request to the server is used as a fallback to prevent
-any dissatisfaction from mirrors. The peer can then share this new
-content with others.
+any dissatisfaction from users. The peer can then share this new
+content with others through the P2P system.
 
@@
 that would efficiently satisfy the demands of a large number of
-users, and should substantially reduce the currently substantial
+users, and should significantly reduce the currently substantial
 bandwidth requirements of hosting this content. We then present an
@@
  \item A small number of the packages can be updated every day,
-       which may require a new torrent, therefore splitting the
-       download population.
+       requiring a new torrent and splitting the
+       download population (even though peers in the new torrent
+       share 99\% of the files in common with peers in the old
+       torrent).
  \item Incentives to share (upload) are no longer needed, as the
@@
 torrents for some of the larger content available, but it ignores
-the smaller content, which is often the more requested content.
+the smaller content, which is often the most popular.
 DebTorrent \cite{debtorrent} makes widespread modifications to a
@@
 distribution system to support it, and separates peers into groups
-based on their characteristics which limits the effectiveness of the
+based on their interest which limits the effectiveness of the
 sharing. Kenosis \cite{kenosis} is a P2P Remote Procedure Call
@@
 Other software vendors also use a similar system. CPAN \cite{cpan}
-distrbutes files containing software packages for the PERL
-programming language. CYGWIN \cite{cygwin} provides many of the
-standard Unix/Linux tools in a Windows environment, also using a
+distributes files containing software packages for the PERL
+programming language, using SOAP RPC requests to find and download
+files. Cygwin \cite{cygwin} provides many of the
+standard Unix/Linux tools in a Windows environment, using a
 package management tool that requests packages from websites. There
@@
  \item The content is avaliable for anyone to freely download.
- \item The content is divided into packages, each of which contains
+ \item The content is divided into distinct units (packages), each of which contains
        a small independent part of all the content.
- \item No peers are interested in downloading all of the content
+ \item Users are typically not interested in downloading all of the content
        available.
@@
  \item Requests for the content are sent by a tool, not directly by
-       the user, though the tool is responding to requests from the user.
+       the user (though the tool is responding to requests from the user).
 \end{itemize}
@@
 identical and has usually been tailored to the system. The small
-differences will make it impossible to distribute similar content
+differences will change the hash of the files, and so
+will make it impossible to distribute similar content
 across systems. And, although peer-to-peer systems scale very well
 with the number of peers in the system, there is some overhead
-involved. So, having a much larger system of peers would mean that
+involved, so having a much larger system of peers would mean that
 requests could take longer to complete.
@@
 The situation described in section \ref{situation} presents a clear
-opportunity to use some form of peer-to-peer file sharing protocols.
+opportunity to use some form of peer-to-peer file-sharing protocol.
 This sparse interest in a large number of packages, and constant
@@
 retrieved strings can then be pointers to the peers that have the
-content package that hashes to that key. A downloading peer can then
-lookup the package hash in the DHT, and then, if it is found,
+content package that hashes to that key. A downloading peer can
+lookup the package hash in the DHT and, if it is found,
 download the file from those peers and verify the content. If the
@@
 mirrors), and once complete will add a new entry to the DHT
-indicating that it has the package.
+indicating that it has the content.
 
@@
 efficient lookups and to support uploading of already downloaded
-content. Many of the package management software implmentations use
+content, whereas the tools typically only run until the download request is complete.
+Many of the package management software implementations use
 HTTP requests to download the files, which makes it possible to
@@
 
-In addition to storing the file download location (which would still
+Since the package management system only stores a hash of the entire
+content, and not of sub-pieces of that content, we will need to be
+able to store and retrieve these sub-piece hashes using the P2P protocol.
+In addition to storing the file download location in the DHT (which would still
 be used for small files), a peer will store a \emph{torrent string}
-containing the peer's hashes of the individual pieces for the larger
+containing the peer's hashes of the sub-pieces of the larger
 files. These piece hashes could be compared ahead of time to
@@
 
-For very large files (about 5 or more pieces), the torrent strings
+For very large files (5 or more pieces), the torrent strings
 are too long to store in the DHT (a single UDP packet should be less
@@
 the torrent string is short enough to store in the DHT (i.e. less
-than 1472 bytes, or about 70 pieces for the SHA1 hash), then a
+than 1472 bytes, or 70 pieces for the SHA1 hash), then a
 lookup of that hash in the DHT would give the torrent-like string.
-Otherwise, a request to the peer for the hash (just like files are
-downloaded), should return the torrent string.
+Otherwise, a request to the peer for the hash (using the same
+method as file downloads), should return the torrent string.
 
@@
 \texttt{apt-dht}, interacts with the \texttt{apt} tool found in most
-Debian-based Linux distributions. Our tool uses SHA1 hashes to
+Debian-based Linux distributions. Apt uses SHA1 hashes to
 verify most downloaded files, including the large index files that
@@
 
-Since all requests from apt are in the form of HTTP requests to a
+Since all requests from apt are in the form of HTTP downloads from a
 server, the implementation takes the form of a caching HTTP proxy.
-Pointing a standard apt implementation to use the proxy is as simple
+Making a standard apt implementation use the proxy is then as simple
 as prepending the proxy location and port to the front of the mirror
@@
 
-The DHT is based on Khashmir, which is an implementation of the
-Kademlia DHT using methods familiar to BitTorrent developers. The
+The DHT is based on Khashmir \cite{khashmir}, which is an implementation of the
+Kademlia DHT \cite{kademlia} using methods familiar to BitTorrent
+developers. It is the same DHT used by most of the existing
+BitTorrent clients to implement trackerless operation. The
 communication is all handled by UDP messages, and RPC requests are
-bencoded similar to torrent files. Khashmir uses the high-level
-Twisted event-driven networking engine, so Twisted is also used for
+bencoded in the same way as torrent files. Khashmir uses the high-level
+Twisted event-driven networking engine \cite{twisted}, so Twisted is also used for
 all other networking needs.
@@
 Downloading is accomplished by sending simple HTTP requests to the
-peer's identified by lookups in the DHT to have the desired file.
+peer's identified (by lookups in the DHT) to have the desired file.
 The HTTP server used for the proxy also doubles as the server
-listeneing for requests for downloads from other peers. All peers
+listening for requests for downloads from other peers. All peers
 support HTTP/1.1, both in the server and the client, which allows
@@
 packages are quite small, and so most will therefore not require
-sub-piece information to download. We have chosen a maximum piece
+sub-piece information to download. We have chosen a piece
 size of 512 kB, which means that 17,515 (78\%) of the packages will
@@
 only 62 packages that require more than 70 pieces, and so will
-require a separate request to the user for the torrent string.
+require a separate request to a peer for the torrent string.
 
@@
 Though functional and useful, the current implementation is not
-complete, and is missing some of the downloading details necessary
-to make it complete. It operates as a caching HTTP proxy, serving
+complete, and is missing some of the downloading details necessary.
+It operates as a caching HTTP proxy, serving
 requests from apt and other peers. Index files are identified by the
-current implementation, and parsed to find hashed of files for DHT
-lookups if they are later requested. Files are downloaded from
-peers, and those not available in peers are downloaded correctly
+current implementation, and parsed to find the hashes of files for later DHT
+lookups if they are requested. Files are downloaded from
+peers, and those not available in peers are downloaded directly
 from the mirror. Any downloaded files are hash checked to verify
@@
 
+\section{Future Work}
+\label{future}
+
+There are some additional steps to be taken to succeed with this
+project. First, the storage method of sub-piece hashes needs to be
+finalized and properly implemented. The communication built into the
+Khashmir-based DHT also needs to be solidified. Some additional peer
+statistics data needs to be made available for each peer, so that
+some form of analysis is available once the system is deployed. The
+reference implementation then needs to be packaged for upload to the
+Debian package archive, so that it can benefit from the widest
+possible deployment among users.
+
+Finally, once deployed the system needs to be analyzed to determine
+its effectiveness at meeting the goals of this project. The analysis
+will consist of measuring, over time:
+\begin{itemize}
+ \item The number of peers using the system.
+ \item The amount of downloaded data by peers from other peers.
+ \item The amount of downloaded data by peers from mirrors/servers.
+ \item The bandwidth used to maintain the DHT.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusions}
@@
 conjuction with Debian-based distribution of Linux software
-packages. It will soone be in use by many users of the Debian
+packages. It will soon be in use by many users of the Debian
 project's distribution, and so serves as an example of the

### 2008-03-05 apt_dht_Khashmir/actions.py
Document the DHT's actions module.

@@ -1,6 +1,8 @@
 ## Copyright 2002-2004 Andrew Loewenstern, All Rights Reserved
 # see LICENSE.txt for license information
 
+"""Details of how to perform actions on remote peers."""
+
 from twisted.internet import reactor
 from twisted.python import log
 
@@ -8,9 +10,64 @@ from khash import intify
 from util import uncompact
 
 class ActionBase:
-    """ base class for some long running asynchronous proccesses like finding nodes or values """
+    """Base class for some long running asynchronous proccesses like finding nodes or values.
+    
+    @type caller: L{khashmir.Khashmir}
+    @ivar caller: the DHT instance that is performing the action
+    @type target: C{string}
+    @ivar target: the target of the action, usually a DHT key
+    @type config: C{dictionary}
+    @ivar config: the configuration variables for the DHT
+    @type action: C{string}
+    @ivar action: the name of the action to call on remote nodes
+    @type num: C{long}
+    @ivar num: the target key in integer form
+    @type queried: C{dictionary}
+    @ivar queried: the nodes that have been queried for this action,
+        keys are node IDs, values are the node itself
+    @type answered: C{dictionary}
+    @ivar answered: the nodes that have answered the queries
+    @type found: C{dictionary}
+    @ivar found: nodes that have been found so far by the action
+    @type sorted_nodes: C{list} of L{node.Node}
+    @ivar sorted_nodes: a sorted list of nodes by there proximity to the key
+    @type results: C{dictionary}
+    @ivar results: keys are the results found so far by the action
+    @type desired_results: C{int}
+    @ivar desired_results: the minimum number of results that are needed
+        before the action should stop
+    @type callback: C{method}
+    @ivar callback: the method to call with the results
+    @type outstanding: C{int}
+    @ivar outstanding: the number of requests currently outstanding
+    @type outstanding_results: C{int}
+    @ivar outstanding_results: the number of results that are expected from
+        the requests that are currently outstanding
+    @type finished: C{boolean}
+    @ivar finished: whether the action is done
+    @type sort: C{method}
+    @ivar sort: used to sort nodes by their proximity to the target
+    """
+    
     def __init__(self, caller, target, callback, config, action, num_results = None):
-        """Initialize the action."""
+        """Initialize the action.
+        
+        @type caller: L{khashmir.Khashmir}
+        @param caller: the DHT instance that is performing the action
+        @type target: C{string}
+        @param target: the target of the action, usually a DHT key
[…]

### 2008-03-05 apt_dht_Khashmir/db.py
Document the DHT's db module.

@@ -1,4 +1,6 @@
 
+"""An sqlite database for storing nodes and key/value pairs."""
+
 from datetime import datetime, timedelta
 from pysqlite2 import dbapi2 as sqlite
 from binascii import a2b_base64, b2a_base64
@@ -15,18 +17,32 @@ class khash(str):
     
 class dht_value(str):
     """Dummy class to convert all DHT values to base64 for storing in the DB."""
-    
+
+# Initialize the database to work with 'khash' objects (binary strings)
 sqlite.register_adapter(khash, b2a_base64)
 sqlite.register_converter("KHASH", a2b_base64)
 sqlite.register_converter("khash", a2b_base64)
+
+# Initialize the database to work with DHT values (binary strings)
 sqlite.register_adapter(dht_value, b2a_base64)
 sqlite.register_converter("DHT_VALUE", a2b_base64)
 sqlite.register_converter("dht_value", a2b_base64)
 
 class DB:
-    """Database access for storing persistent data."""
+    """An sqlite database for storing persistent node info and key/value pairs.
+    
+    @type db: C{string}
+    @ivar db: the database file to use
+    @type conn: L{pysqlite2.dbapi2.Connection}
+    @ivar conn: an open connection to the sqlite database
+    """
     
     def __init__(self, db):
+        """Load or create the database file.
+        
+        @type db: C{string}
+        @param db: the database file to use
+        """
         self.db = db
         try:
             os.stat(db)
@@ -39,8 +55,10 @@ class DB:
             sqlite.register_converter("text", str)
         else:
             self.conn.text_factory = str
-        
+
+    #{ Loading the DB
     def _loadDB(self, db):
+        """Open a new connection to the existing database file"""
         try:
             self.conn = sqlite.connect(database=db, detect_types=sqlite.PARSE_DECLTYPES)
         except:
@@ -48,16 +66,23 @@ class DB:
             raise DBExcept, "Couldn't open DB", traceback.format_exc()
         
     def _createNewDB(self, db):
+        """Open a connection to a new database and create the necessary tables."""
         self.conn = sqlite.connect(database=db, detect_types=sqlite.PARSE_DECLTYPES)
         c = self.conn.cursor()
-        c.execute("CREATE TABLE kv (key KHASH, value DHT_VALUE, last_refresh TIMESTAMP, PRIMARY KEY (key, value))")
+        c.execute("CREATE TABLE kv (key KHASH, value DHT_VALUE, last_refresh TIMESTAMP, "+
[…]

### 2008-03-05 apt_dht_Khashmir/ktable.py
Document the DHT's ktable module.

@@ -1,6 +1,8 @@
 ## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
 # see LICENSE.txt for license information
 
+"""The routing table and buckets for a kademlia-like DHT."""
+
 from datetime import datetime
 from bisect import bisect_left
 
@@ -11,8 +13,24 @@ import khash
 from node import Node, NULL_ID
 
 class KTable:
-    """local routing table for a kademlia like distributed hash table"""
+    """Local routing table for a kademlia-like distributed hash table.
+    
+    @type node: L{node.Node}
+    @ivar node: the local node
+    @type config: C{dictionary}
+    @ivar config: the configuration parameters for the DHT
+    @type buckets: C{list} of L{KBucket}
+    @ivar buckets: the buckets of nodes in the routing table
+    """
+    
     def __init__(self, node, config):
+        """Initialize the first empty bucket of everything.
+        
+        @type node: L{node.Node}
+        @param node: the local node
+        @type config: C{dictionary}
+        @param config: the configuration parameters for the DHT
+        """
         # this is the root node, a.k.a. US!
         assert node.id != NULL_ID
         self.node = node
@@ -20,14 +38,18 @@ class KTable:
         self.buckets = [KBucket([], 0L, 2L**self.config['HASH_LENGTH'])]
         
     def _bucketIndexForInt(self, num):
-        """the index of the bucket that should hold int"""
+        """Find the index of the bucket that should hold the node's ID number."""
         return bisect_left(self.buckets, num)
     
     def findNodes(self, id):
+        """Find the K nodes in our own local table closest to the ID.
+
+        @type id: C{string} of C{int} or L{node.Node}
+        @param id: the ID to find nodes that are close to
+        @raise TypeError: if id does not properly identify an ID
         """
-            return K nodes in our own local table closest to the ID.
-        """
-        
+
+        # Get the ID number from the input
         if isinstance(id, str):
             num = khash.intify(id)
         elif isinstance(id, Node):
@@ -40,7 +62,7 @@ class KTable:
         nodes = []
         i = self._bucketIndexForInt(num)
         
-        # if this node is already in our table then return it
+        # If this node is already in our table then return it
         try:
             index = self.buckets[i].l.index(num)
         except ValueError:
@@ -48,14 +70,16 @@ class KTable:
         else:
             return [self.buckets[i].l[index]]
             
-        # don't have the node, get the K closest nodes
[…]

### 2008-03-05 apt_dht_Khashmir/node.py
Document the DHT's node module.

@@ -1,6 +1,12 @@
 ## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
 # see LICENSE.txt for license information
 
+"""Represents a node in the DHT.
+
+@type NULL_ID: C{string}
+@var NULL_ID: the node ID to use until one is known
+"""
+
 from datetime import datetime, MINYEAR
 from types import InstanceType
 
@@ -13,8 +19,40 @@ from util import compact
 NULL_ID = 20 * '\0'
 
 class Node:
-    """encapsulate contact info"""
+    """Encapsulate a node's contact info.
+    
+    @type fails: C{int}
+    @ivar fails: number of times this node has failed in a row
+    @type lastSeen: C{datetime.datetime}
+    @ivar lastSeen: the last time a response was received from this node
+    @type id: C{string}
+    @ivar id: the node's ID in the DHT
+    @type num: C{long}
+    @ivar num: the node's ID in number form
+    @type host: C{string}
+    @ivar host: the IP address of the node
+    @type port: C{int}
+    @ivar port: the port of the node
+    @type token: C{string}
+    @ivar token: the last received token from the node
+    @type num_values: C{int}
+    @ivar num_values: the number of values the node has for the key in the
+        currently executing action
+    """
+    
     def __init__(self, id, host = None, port = None):
+        """Initialize the node.
+        
+        @type id: C{string} or C{dictionary}
+        @param id: the node's ID in the DHT, or a dictionary containing the
+            node's id, host and port
+        @type host: C{string}
+        @param host: the IP address of the node
+            (optional, but must be specified if id is not a dictionary)
+        @type port: C{int}
+        @param port: the port of the node
+            (optional, but must be specified if id is not a dictionary)
+        """
         self.fails = 0
         self.lastSeen = datetime(MINYEAR, 1, 1)
 
@@ -35,20 +73,29 @@ class Node:
         self._contactInfo = None
     
     def updateLastSeen(self):
+        """Updates the last contact time of the node and resets the number of failures."""
         self.lastSeen = datetime.now()
         self.fails = 0
         
     def updateToken(self, token):
+        """Update the token for the node."""
         self.token = token
     
     def updateNumValues(self, num_values):
+        """Update how many values the node has in the current search for a value."""
         self.num_values = num_values
     
     def msgFailed(self):
+        """Log a failed attempt to contact this node.
+        
+        @rtype: C{int}
[…]

### 2008-03-05 apt_dht_Khashmir/khashmir.py
Document the DHT's main khashmir module.

@@ -1,6 +1,8 @@
 ## Copyright 2002-2004 Andrew Loewenstern, All Rights Reserved
 # see LICENSE.txt for license information
 
+"""The main Khashmir program."""
+
 import warnings
 warnings.simplefilter("ignore", DeprecationWarning)
 
@@ -20,155 +22,257 @@ from khash import newID, newIDInRange
 from actions import FindNode, FindValue, GetValue, StoreValue
 import krpc
 
-# this is the base class, has base functionality and find node, no key-value mappings
 class KhashmirBase(protocol.Factory):
+    """The base Khashmir class, with base functionality and find node, no key-value mappings.
+    
+    @type _Node: L{node.Node}
+    @ivar _Node: the knode implementation to use for this class of DHT
+    @type config: C{dictionary}
+    @ivar config: the configuration parameters for the DHT
+    @type port: C{int}
+    @ivar port: the port to listen on
+    @type store: L{db.DB}
+    @ivar store: the database to store nodes and key/value pairs in
+    @type node: L{node.Node}
+    @ivar node: this node
+    @type table: L{ktable.KTable}
+    @ivar table: the routing table
+    @type token_secrets: C{list} of C{string}
+    @ivar token_secrets: the current secrets to use to create tokens
+    @type udp: L{krpc.hostbroker}
+    @ivar udp: the factory for the KRPC protocol
+    @type listenport: L{twisted.internet.interfaces.IListeningPort}
+    @ivar listenport: the UDP listening port
+    @type next_checkpoint: L{twisted.internet.interfaces.IDelayedCall}
+    @ivar next_checkpoint: the delayed call for the next checkpoint
+    """
+    
     _Node = KNodeBase
+    
     def __init__(self, config, cache_dir='/tmp'):
+        """Initialize the Khashmir class and call the L{setup} method.
+        
+        @type config: C{dictionary}
+        @param config: the configuration parameters for the DHT
+        @type cache_dir: C{string}
+        @param cache_dir: the directory to store all files in
+            (optional, defaults to the /tmp directory)
+        """
         self.config = None
         self.setup(config, cache_dir)
         
     def setup(self, config, cache_dir):
+        """Setup all the Khashmir sub-modules.
+        
+        @type config: C{dictionary}
+        @param config: the configuration parameters for the DHT
+        @type cache_dir: C{string}
+        @param cache_dir: the directory to store all files in
+        """
         self.config = config
         self.port = config['PORT']
[…]

### 2008-03-05 apt_dht/AptPackages.py
Various documentation fixes and additions.

@@ -203,7 +203,7 @@ class AptPackages:
         self.apt_config['Dir'] = self.cache_dir.path
         self.apt_config['Dir::State::status'] = self.cache_dir.preauthChild(self.apt_config['Dir::State']).preauthChild(self.apt_config['Dir::State::status']).path
         self.packages = PackageFileList(cache_dir)
-        self.loaded = 0
+        self.loaded = False
         self.loading = None
         self.unload_later = None
         
@@ -321,7 +321,7 @@ class AptPackages:
         else:
             self.srcrecords = None
 
-        self.loaded = 1
+        self.loaded = True
         return True
 
     def unload(self):
@@ -336,7 +336,7 @@ class AptPackages:
             del self.records
             del self.srcrecords
             del self.indexrecords
-            self.loaded = 0
+            self.loaded = False
 
     def cleanup(self):
         """Cleanup and close any loaded caches."""

### 2008-03-05 apt_dht/__init__.py

@@ -1,6 +1,13 @@
 
 """The main apt-dht modules.
 
+To run apt-dht, you probably want to do something like::
+
+  from apt_dht.apt_dht import AptDHT
+  myapp = AptDHT(myDHT)
+
+where myDHT is a DHT that implements interfaces.IDHT.
+
 Diagram of the interaction between the given modules::
   
   +---------------+    +-----------------------------------+    +-------------

### 2008-03-05 apt_dht_Khashmir/actions.py

@@ -82,7 +82,7 @@ class ActionBase:
         self.callback = callback
         self.outstanding = 0
         self.outstanding_results = 0
-        self.finished = 0
+        self.finished = False
     
         def sort(a, b, num=self.num):
             """Sort nodes relative to the ID we are looking for."""
@@ -111,7 +111,7 @@ class ActionBase:
         if self.desired_results and ((len(self.results) >= abs(self.desired_results)) or
                                      (self.desired_results < 0 and
                                       len(self.answered) >= self.config['STORE_REDUNDANCY'])):
-            self.finished=1
+            self.finished = True
             result = self.generateResult()
             reactor.callLater(0, self.callback, *result)
 
@@ -156,7 +156,7 @@ class ActionBase:
 
         # If no requests are outstanding, then we are done
         if self.outstanding == 0:
-            self.finished = 1
+            self.finished = True
             result = self.generateResult()
             reactor.callLater(0, self.callback, *result)
 

### 2008-03-05 apt_dht_Khashmir/bencode.py

@@ -172,7 +172,7 @@ decode_func['9'] = decode_string
 decode_func['u'] = decode_unicode
 decode_func['t'] = decode_datetime
   
-def bdecode(x, sloppy = 0):
+def bdecode(x, sloppy = False):
     """Bdecode a string of data.
     
     @type x: C{string}

### 2008-03-05 apt_dht_Khashmir/db.py

@@ -108,10 +108,7 @@ class DB:
         self.conn.commit()
         
     def getRoutingTable(self):
-        """Load routing table nodes from database.
-        
-        It's usually a good idea to call refreshTable(force=1) after loading the table.
-        """
+        """Load routing table nodes from database."""
         c = self.conn.cursor()
         c.execute("SELECT * FROM nodes")
         return c.fetchall()

### 2008-03-05 apt_dht_Khashmir/node.py

@@ -21,6 +21,8 @@ NULL_ID = 20 * '\0'
 class Node:
     """Encapsulate a node's contact info.
     
+    @ivar conn: the connection to the remote node (added externally)
+    @ivar table: the routing table (added externally)
     @type fails: C{int}
     @ivar fails: number of times this node has failed in a row
     @type lastSeen: C{datetime.datetime}

### 2008-03-06 docs/motivation/motivation.tex
Rename project to apt-p2p.

@@
 in section \ref{opportunity}. This software, called
-\texttt{apt-dht}, interacts with the \texttt{apt} tool found in most
+\texttt{apt-p2p}, interacts with the \texttt{apt} tool found in most
 Debian-based Linux distributions. Apt uses SHA1 hashes to
@@
 \centering
-\includegraphics[width=\columnwidth]{apt_dht_simulation-size_CDF.eps}
+\includegraphics[width=\columnwidth]{apt_p2p_simulation-size_CDF.eps}
 \caption{The CDF of the size of packages in a Debian system, both

### 2008-03-06 apt-dht.py

@@ -1,9 +1,9 @@
 #!/usr/bin/env python
 
-# Load apt-dht application
+# Load apt-p2p application
 #
-# There are two ways apt-dht can be started:
-#  1. twistd -y apt-dht
+# There are two ways apt-p2p can be started:
+#  1. twistd -y apt-p2p
 #     - twistd will load this file and execute the app
 #       in 'application' variable
 #  2. from command line
@@ -16,14 +16,14 @@ from twisted.internet import reactor
 from twisted.python import usage, log
 from twisted.web2 import channel
 
-from apt_dht.apt_dht_conf import config, version, DEFAULT_CONFIG_FILES
-from apt_dht.interfaces import IDHT
+from apt_p2p.apt_p2p_conf import config, version, DEFAULT_CONFIG_FILES
+from apt_p2p.interfaces import IDHT
 
 config_file = ''
 
 if __name__ == '__main__':
     # Parse command line parameters when started on command line
-    class AptDHTOptions(usage.Options):
+    class AptP2POptions(usage.Options):
         optFlags = [
             ['help', 'h', 'Print this help message'],
             ]
@@ -31,12 +31,12 @@ if __name__ == '__main__':
             ['config-file', 'c', '', "Configuration file"],
             ['log-file', 'l', '-', "File to log to, - for stdout"],
             ]
-        longdesc="apt-dht is a peer-to-peer downloader for apt users"
+        longdesc="apt-p2p is a peer-to-peer downloader for apt users"
         def opt_version(self):
-            print "apt-dht %s" % version.short()
+            print "apt-p2p %s" % version.short()
             sys.exit(0)
 
-    opts = AptDHTOptions()
+    opts = AptP2POptions()
     try:
         opts.parseOptions()
     except usage.UsageError, ue:
@@ -60,9 +60,9 @@ else:
     uid,gid = None,None
 
 log.msg('Starting application')
-application = service.Application("apt-dht", uid, gid)
+application = service.Application("apt-p2p", uid, gid)
 #print service.IProcess(application).processName
-#service.IProcess(application).processName = 'apt-dht'
+#service.IProcess(application).processName = 'apt-p2p'
 
 log.msg('Starting DHT')
 DHT = __import__(config.get('DEFAULT', 'DHT')+'.DHT', globals(), locals(), ['DHT'])
@@ -71,8 +71,8 @@ myDHT = DHT.DHT()
 
[…]

### 2008-03-06 apt_dht/AptPackages.py

@@ -285,7 +285,7 @@ class AptPackages:
             file = self.packages[f]
             if f.split('/')[-1] == "Release":
                 self.addRelease(f, file)
-            fake_uri='http://apt-dht'+f
+            fake_uri='http://apt-p2p'+f
             fake_dirname = '/'.join(fake_uri.split('/')[:-1])
             if f.endswith('Sources'):
                 deb_src_added = True
@@ -434,7 +434,7 @@ class TestAptPackages(unittest.TestCase):
     
     def setUp(self):
         """Initializes the cache with files found in the traditional apt location."""
-        self.client = AptPackages(FilePath('/tmp/.apt-dht'), 300)
+        self.client = AptPackages(FilePath('/tmp/.apt-p2p'), 300)
     
         # Find the largest index files that are for 'main'
         self.packagesFile = os.popen('ls -Sr /var/lib/apt/lists/ | grep -E "_main_.*Packages$" | tail -n 1').read().rstrip('\n')

### 2008-03-06 apt_dht/CacheManager.py

@@ -200,7 +200,7 @@ class CacheManager:
     @ivar all_dirs: all the directories that have cached files in them
     @type db: L{db.DB}
     @ivar db: the database to use for tracking files and hashes
-    @type manager: L{apt_dht.AptDHT}
+    @type manager: L{apt_p2p.AptP2P}
     @ivar manager: the main program object to send requests to
     @type scanning: C{list} of L{twisted.python.filepath.FilePath}
     @ivar scanning: all the directories that are currectly being scanned or waiting to be scanned
@@ -216,7 +216,7 @@ class CacheManager:
         @type other_dirs: C{list} of L{twisted.python.filepath.FilePath}
         @param other_dirs: the other directories that have shared files in them
             (optional, defaults to only using the cache directory)
-        @type manager: L{apt_dht.AptDHT}
+        @type manager: L{apt_p2p.AptP2P}
         @param manager: the main program object to send requests to
             (optional, defaults to not calling back with cached files)
         """
@@ -430,7 +430,7 @@ class TestMirrorManager(unittest.TestCase):
     client = None
     
     def setUp(self):
-        self.client = CacheManager(FilePath('/tmp/.apt-dht'))
+        self.client = CacheManager(FilePath('/tmp/.apt-p2p'))
         
     def tearDown(self):
         for p in self.pending_calls:

### 2008-03-06 apt_dht/HTTPDownloader.py

@@ -15,7 +15,7 @@ from twisted.web2 import version as web2_version
 from twisted.trial import unittest
 from zope.interface import implements
 
-from apt_dht_conf import version
+from apt_p2p_conf import version
 
 class Peer(ClientFactory):
     """A manager for all HTTP requests to a single peer.
@@ -158,7 +158,7 @@ class Peer(ClientFactory):
         """Get the common HTTP headers for all requests."""
         headers = http_headers.Headers()
         headers.setHeader('Host', self.host)
-        headers.setHeader('User-Agent', 'apt-dht/%s (twisted/%s twisted.web2/%s)' % 
+        headers.setHeader('User-Agent', 'apt-p2p/%s (twisted/%s twisted.web2/%s)' % 
                           (version.short(), twisted_version.short(), web2_version.short()))
         return headers
     

### 2008-03-06 apt_dht/HTTPServer.py

@@ -10,7 +10,7 @@ from twisted.web2 import server, http, resource, channel, stream
 from twisted.web2 import static, http_headers, responsecode
 
 from policies import ThrottlingFactory
-from apt_dht_Khashmir.bencode import bencode
+from apt_p2p_Khashmir.bencode import bencode
 
 class FileDownloader(static.File):
     """Modified to make it suitable for apt requests.
@@ -19,7 +19,7 @@ class FileDownloader(static.File):
     freshness before being sent. Requests for unfound and stale files are
     forwarded to the main program for downloading.
     
-    @type manager: L{apt_dht.AptDHT}
+    @type manager: L{apt_p2p.AptP2P}
     @ivar manager: the main program to query 
     """
     
@@ -134,7 +134,7 @@ class TopLevel(resource.Resource):
     @ivar directory: the directory to check for cached files
     @type db: L{db.DB}
     @ivar db: the database to use for looking up files and hashes
-    @type manager: L{apt_dht.AptDHT}
+    @type manager: L{apt_p2p.AptP2P}
     @ivar manager: the main program object to send requests to
     @type factory: L{twisted.web2.channel.HTTPFactory} or L{policies.ThrottlingFactory}
     @ivar factory: the factory to use to server HTTP requests
@@ -150,7 +150,7 @@ class TopLevel(resource.Resource):
         @param directory: the directory to check for cached files
         @type db: L{db.DB}
         @param db: the database to use for looking up files and hashes
-        @type manager: L{apt_dht.AptDHT}
+        @type manager: L{apt_p2p.AptP2P}
         @param manager: the main program object to send requests to
         """
         self.directory = directory

### 2008-03-06 apt_dht/Hash.py

@@ -299,10 +299,10 @@ class TestHashObject(unittest.TestCase):
                 found = True
                 break
         self.failUnless(found == True)
-        h.set(hashType, 'c722df87e1acaa64b27aac4e174077afc3623540', '19')
+        h.set(hashType, '3bba0a5d97b7946ad2632002bf9caefe2cb18e00', '19')
         h.new()
-        h.update('apt-dht is the best')
-        self.failUnless(h.hexdigest() == 'c722df87e1acaa64b27aac4e174077afc3623540')
+        h.update('apt-p2p is the best')
+        self.failUnless(h.hexdigest() == '3bba0a5d97b7946ad2632002bf9caefe2cb18e00')
         self.failUnlessRaises(HashError, h.update, 'gfgf')
         self.failUnless(h.verify() == True)
         
@@ -315,10 +315,10 @@ class TestHashObject(unittest.TestCase):
                 found = True
                 break
         self.failUnless(found == True)
-        h.set(hashType, '2a586bcd1befc5082c872dcd96a01403', '19')
+        h.set(hashType, '6b5abdd30d7ed80edd229f9071d8c23c', '19')
         h.new()
-        h.update('apt-dht is the best')
-        self.failUnless(h.hexdigest() == '2a586bcd1befc5082c872dcd96a01403')
+        h.update('apt-p2p is the best')
+        self.failUnless(h.hexdigest() == '6b5abdd30d7ed80edd229f9071d8c23c')
         self.failUnlessRaises(HashError, h.update, 'gfgf')
         self.failUnless(h.verify() == True)
         
@@ -331,10 +331,10 @@ class TestHashObject(unittest.TestCase):
                 found = True
                 break
         self.failUnless(found == True)
-        h.set(hashType, '55b971f64d9772f733de03f23db39224f51a455cc5ad4c2db9d5740d2ab259a7', '19')
+        h.set(hashType, '47f2238a30a0340faa2bf01a9bdc42ba77b07b411cda1e24cd8d7b5c4b7d82a7', '19')
         h.new()
-        h.update('apt-dht is the best')
-        self.failUnless(h.hexdigest() == '55b971f64d9772f733de03f23db39224f51a455cc5ad4c2db9d5740d2ab259a7')
+        h.update('apt-p2p is the best')
+        self.failUnless(h.hexdigest() == '47f2238a30a0340faa2bf01a9bdc42ba77b07b411cda1e24cd8d7b5c4b7d82a7')
[…]

### 2008-03-06 apt_dht/MirrorManager.py

@@ -133,7 +133,7 @@ class TestMirrorManager(unittest.TestCase):
     client = None
     
     def setUp(self):
-        self.client = MirrorManager(FilePath('/tmp/.apt-dht'), 300)
+        self.client = MirrorManager(FilePath('/tmp/.apt-p2p'), 300)
         
     def test_extractPath(self):
         """Test extracting the site and base directory from various mirrors."""

### 2008-03-06 apt_dht/__init__.py

@@ -1,17 +1,17 @@
 
-"""The main apt-dht modules.
+"""The main apt-p2p modules.
 
-To run apt-dht, you probably want to do something like::
+To run apt-p2p, you probably want to do something like::
 
-  from apt_dht.apt_dht import AptDHT
-  myapp = AptDHT(myDHT)
+  from apt_p2p.apt_p2p import AptP2P
+  myapp = AptP2P(myDHT)
 
 where myDHT is a DHT that implements interfaces.IDHT.
 
 Diagram of the interaction between the given modules::
   
   +---------------+    +-----------------------------------+    +-------------
-  |     AptDHT    |    |               DHT                 |    |  Internet
+  |     AptP2P    |    |               DHT                 |    |  Internet
   |               |--->|join                            DHT|----|--\    
   |               |--->|loadConfig                         |    |  | Another
   |               |--->|getValue                           |    |  | Node

### 2008-03-06 apt_dht/apt_dht.py

@@ -18,7 +18,7 @@ from twisted.web2 import server, http, http_headers, static
 from twisted.python import log, failure
 from twisted.python.filepath import FilePath
 
-from apt_dht_conf import config
+from apt_p2p_conf import config
 from PeerManager import PeerManager
 from HTTPServer import TopLevel
 from MirrorManager import MirrorManager
@@ -32,7 +32,7 @@ TORRENT_PIECES = 70
 
 download_dir = 'cache'
 
-class AptDHT:
+class AptP2P:
     """The main code object that does all of the work.
     
     Contains all of the sub-components that do all the low-level work, and
@@ -65,11 +65,11 @@ class AptDHT:
         @type dht: L{interfaces.IDHT}
         @param dht: the DHT instance to use
         """
-        log.msg('Initializing the main apt_dht application')
+        log.msg('Initializing the main apt_p2p application')
         self.cache_dir = FilePath(config.get('DEFAULT', 'cache_dir'))
         if not self.cache_dir.child(download_dir).exists():
             self.cache_dir.child(download_dir).makedirs()
-        self.db = DB(self.cache_dir.child('apt-dht.db'))
+        self.db = DB(self.cache_dir.child('apt-p2p.db'))
         self.dht = dht
         self.dht.loadConfig(config, config.get('DEFAULT', 'DHT'))
         self.dht.join().addCallbacks(self.joinComplete, self.joinError)

### 2008-03-06 apt_dht/apt_dht_conf.py

@@ -22,7 +22,7 @@ class ConfigError(Exception):
     def __str__(self):
         return repr(self.message)
 
-version = versions.Version('apt-dht', 0, 0, 0)
+version = versions.Version('apt-p2p', 0, 0, 0)
 
 # Set the home parameter
 home = os.path.expandvars('${HOME}')
@@ -31,8 +31,8 @@ if home == '${HOME}' or not os.path.isdir(home):
     if not os.path.isdir(home):
         home = os.path.abspath(os.path.dirname(sys.argv[0]))
 
-DEFAULT_CONFIG_FILES=['/etc/apt-dht/apt-dht.conf',
-                      home + '/.apt-dht/apt-dht.conf']
+DEFAULT_CONFIG_FILES=['/etc/apt-p2p/apt-p2p.conf',
+                      home + '/.apt-p2p/apt-p2p.conf']
 
 DEFAULTS = {
 
@@ -40,7 +40,7 @@ DEFAULTS = {
     'PORT': '9977',
     
     # Directory to store the downloaded files in
-    'CACHE_DIR': home + '/.apt-dht/cache',
+    'CACHE_DIR': home + '/.apt-p2p/cache',
     
     # Other directories containing packages to share with others
     # WARNING: all files in these directories will be hashed and available
@@ -65,7 +65,7 @@ DEFAULTS = {
     # Which DHT implementation to use.
     # It must be possile to do "from <DHT>.DHT import DHT" to get a class that
     # implements the IDHT interface.
-    'DHT': 'apt_dht_Khashmir',
+    'DHT': 'apt_p2p_Khashmir',
 
     # Whether to only run the DHT (for providing only a bootstrap node)
     'DHT-ONLY': 'no',
@@ -120,7 +120,7 @@ DHT_DEFAULTS = {
     'SPEW': 'yes',
 }
 
-class AptDHTConfigParser(SafeConfigParser):
+class AptP2PConfigParser(SafeConfigParser):
     """Adds 'gettime' and 'getstringlist' to ConfigParser objects.
     
     @ivar time_multipliers: the 'gettime' suffixes and the multipliers needed
@@ -159,7 +159,7 @@ class AptDHTConfigParser(SafeConfigParser):
         return option.upper()
 
 # Initialize the default config parameters
-config = AptDHTConfigParser(DEFAULTS)
+config = AptP2PConfigParser(DEFAULTS)
[…]

### 2008-03-06 apt_dht/db.py

@@ -310,12 +310,12 @@ class TestDB(unittest.TestCase):
     timeout = 5
     db = FilePath('/tmp/khashmir.db')
     hash = '\xca\xec\xb8\x0c\x00\xe7\x07\xf8~])\x8f\x9d\xe5_B\xff\x1a\xc4!'
-    directory = FilePath('/tmp/apt-dht/')
-    file = FilePath('/tmp/apt-dht/khashmir.test')
+    directory = FilePath('/tmp/apt-p2p/')
+    file = FilePath('/tmp/apt-p2p/khashmir.test')
     testfile = 'tmp/khashmir.test'
-    dirs = [FilePath('/tmp/apt-dht/top1'),
-            FilePath('/tmp/apt-dht/top2/sub1'),
-            FilePath('/tmp/apt-dht/top2/sub2/')]
+    dirs = [FilePath('/tmp/apt-p2p/top1'),
+            FilePath('/tmp/apt-p2p/top2/sub1'),
+            FilePath('/tmp/apt-p2p/top2/sub2/')]
 
     def setUp(self):
         if not self.file.parent().exists():

### 2008-03-06 apt_dht/interfaces.py

@@ -1,5 +1,5 @@
 
-"""Some interfaces that are used by the apt-dht classes."""
+"""Some interfaces that are used by the apt-p2p classes."""
 
 from zope.interface import Interface
 

### 2008-03-06 apt_dht/util.py

@@ -1,5 +1,5 @@
 
-"""Some utitlity functions for use in the apt-dht program.
+"""Some utitlity functions for use in the apt-p2p program.
 
 @var isLocal: a compiled regular expression suitable for testing if an
     IP address is from a known local or private range

### 2008-03-06 apt_dht_Khashmir/DHT.py

@@ -13,11 +13,11 @@ from twisted.python import log
 from twisted.trial import unittest
 from zope.interface import implements
 
-from apt_dht.interfaces import IDHT
+from apt_p2p.interfaces import IDHT
 from khashmir import Khashmir
 from bencode import bencode, bdecode
 
-khashmir_dir = 'apt-dht-Khashmir'
+khashmir_dir = 'apt-p2p-Khashmir'
 
 class DHTError(Exception):
     """Represents errors that occur in the DHT."""
@@ -51,7 +51,7 @@ class DHT:
     @type retrieved: C{dictionary}
     @ivar retrieved: keys are the keys for which getValue requests are active,
         values are list of the values returned so far
-    @type config_parser: L{apt_dht.apt_dht_conf.AptDHTConfigParser}
+    @type config_parser: L{apt_p2p.apt_p2p_conf.AptP2PConfigParser}
     @ivar config_parser: the configuration info for the main program
     @type section: C{string}
     @ivar section: the section of the configuration info that applies to the DHT
@@ -76,7 +76,7 @@ class DHT:
         self.retrieved = {}
     
     def loadConfig(self, config, section):
-        """See L{apt_dht.interfaces.IDHT}."""
+        """See L{apt_p2p.interfaces.IDHT}."""
         self.config_parser = config
         self.section = section
         self.config = {}
@@ -104,7 +104,7 @@ class DHT:
                 self.config[k] = self.config_parser.get(section, k)
     
     def join(self):
-        """See L{apt_dht.interfaces.IDHT}."""
+        """See L{apt_p2p.interfaces.IDHT}."""
         if self.config is None:
             raise DHTError, "configuration not loaded"
         if self.joining:
@@ -172,7 +172,7 @@ class DHT:
         return self.foundAddrs
         
     def leave(self):
-        """See L{apt_dht.interfaces.IDHT}."""
+        """See L{apt_p2p.interfaces.IDHT}."""
         if self.config is None:
             raise DHTError, "configuration not loaded"
         
@@ -201,7 +201,7 @@ class DHT:
         return key
 
     def getValue(self, key):
-        """See L{apt_dht.interfaces.IDHT}."""
+        """See L{apt_p2p.interfaces.IDHT}."""
         if self.config is None:
             raise DHTError, "configuration not loaded"
         if not self.joined:
@@ -232,7 +232,7 @@ class DHT:
             del self.retrieving[key]
 
     def storeValue(self, key, value):
[…]

### 2008-03-06 apt_dht_Khashmir/__init__.py

@@ -1,5 +1,5 @@
 
-"""The apt-dht implementation of the Khashmir DHT.
+"""The apt-p2p implementation of the Khashmir DHT.
 
 These modules implement a modified Khashmir, which is a kademlia-like
 Distributed Hash Table available at::
@@ -8,11 +8,11 @@ Distributed Hash Table available at::
 
 The protocol for the implementation's communication is described here::
 
-  http://www.camrdale.org/apt-dht/protocol.html
+  http://www.camrdale.org/apt-p2p/protocol.html
 
 To run the DHT you probably want to do something like::
 
-  from apt_dht_Khashmir import DHT
+  from apt_p2p_Khashmir import DHT
   myDHT = DHT.DHT()
   myDHT.loadConfig(config, section)
   myDHT.join()

### 2008-03-06 apt_dht_Khashmir/util.py

@@ -1,7 +1,7 @@
 ## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
 # see LICENSE.txt for license information
 
-"""Some utitlity functions for use in apt-dht's khashmir DHT."""
+"""Some utitlity functions for use in apt-p2p's khashmir DHT."""
 
 from twisted.trial import unittest
 

### 2008-03-06 setup.py

@@ -12,14 +12,14 @@ c = changelog.Changelog(file=data, max_blocks=1)
 del data
 
 setup(
-    name = "apt-dht",
+    name = "apt-p2p",
     version = c.full_version,
     author = "Cameron Dale",
     author_email = "<camrdale@gmail.com>",
-    url = "http://www.camrdale.org/apt-dht.html",
+    url = "http://www.camrdale.org/apt-p2p.html",
     license = "GPL",
 
-    packages = ["apt_dht", "apt_dht_Khashmir"],
+    packages = ["apt_p2p", "apt_p2p_Khashmir"],
 
-    scripts = ['apt-dht.py']
+    scripts = ['apt-p2p.py']
     )

### 2008-03-06 test.py

@@ -1,9 +1,9 @@
 #!/usr/bin/env python
 
-"""Automated tests of the apt-dht functionality.
+"""Automated tests of the apt-p2p functionality.
 
 This script runs several automatic tests of some of the functionality in
-the apt-dht program.
+the apt-p2p program.
 
 @type tests: C{dictionary}
 @var tests: all of the tests that can be run.
@@ -15,11 +15,11 @@ the apt-dht program.
     
     The bootstrap nodes keys are integers, which must be in the range 1-9.
     The values are the dictionary of string formatting values for creating
-    the apt-dht configuration file (see L{apt_dht_conf_template} below).
+    the apt-p2p configuration file (see L{apt_p2p_conf_template} below).
     
     The downloaders keys are also integers in the range 1-99. The values are
-    the dictionary of string formatting values for creating the apt-dht
-    configuration file (see L{apt_dht_conf_template} below).
+    the dictionary of string formatting values for creating the apt-p2p
+    configuration file (see L{apt_p2p_conf_template} below).
     
     The apt-get commands' list elements are tuples with 2 elements: the
     downloader to run the command on, and the list of command-line
@@ -312,7 +312,7 @@ Debug
 
 }
 """
-apt_dht_conf_template = """
+apt_p2p_conf_template = """
 [DEFAULT]
 
 # Port to listen on for all requests (TCP and UDP)
@@ -344,12 +344,12 @@ KEY_REFRESH = 57m
 # Which DHT implementation to use.
 # It must be possile to do "from <DHT>.DHT import DHT" to get a class that
 # implements the IDHT interface.
-DHT = apt_dht_Khashmir
+DHT = apt_p2p_Khashmir
 
 # Whether to only run the DHT (for providing only a bootstrap node)
 DHT-ONLY = %(DHT-ONLY)s
 
-[apt_dht_Khashmir]
+[apt_p2p_Khashmir]
 # bootstrap nodes to contact to join the DHT
 BOOTSTRAP = %(BOOTSTRAP)s
 
@@ -579,7 +579,7 @@ def start_downloader(bootstrap_addresses, num_down, options = {},
     @param num_down: the number of the downloader to use
     @type options: C{dictionary}
     @param options: the dictionary of string formatting values for creating
-        the apt-dht configuration file (see L{apt_dht_conf_template} above).
+        the apt-p2p configuration file (see L{apt_p2p_conf_template} above).
[…]

### 2008-03-08 apt_p2p_Khashmir/ktable.py
Add new node to table when refreshing old node.
Fixes a bug introduced by commit:

    6ce38471: fix peer update problem connection refused in store value

@@ -169,11 +169,10 @@ class KTable:
                 # It responded, so update it
                 node.updateLastSeen()
                 # move node to end of bucket
-                xnode = self.buckets[i].l[it]
                 del(self.buckets[i].l[it])
                 # note that we removed the original and replaced it with the new one
                 # utilizing this nodes new contact info
-                self.buckets[i].l.append(xnode)
+                self.buckets[i].l.append(node)
                 self.buckets[i].touch()
             return
         

### 2008-03-08 apt_p2p_Khashmir/ktable.py
Don't use the tables findNodes when you can just bisect for it.

@@ -41,8 +41,8 @@ class KTable:
         """Find the index of the bucket that should hold the node's ID number."""
         return bisect_left(self.buckets, num)
     
-    def findNodes(self, id):
-        """Find the K nodes in our own local table closest to the ID.
+    def _nodeNum(self, id):
+        """Takes different types of input and converts to the node ID number.
 
         @type id: C{string} of C{int} or L{node.Node}
         @param id: the ID to find nodes that are close to
@@ -51,27 +51,27 @@ class KTable:
 
         # Get the ID number from the input
         if isinstance(id, str):
-            num = khash.intify(id)
+            return khash.intify(id)
         elif isinstance(id, Node):
-            num = id.num
+            return id.num
         elif isinstance(id, int) or isinstance(id, long):
-            num = id
+            return id
         else:
-            raise TypeError, "findNodes requires an int, string, or Node"
+            raise TypeError, "requires an int, string, or Node input"
             
-        nodes = []
+    def findNodes(self, id):
+        """Find the K nodes in our own local table closest to the ID.
+
+        @type id: C{string} of C{int} or L{node.Node}
+        @param id: the ID to find nodes that are close to
+        """
+
+        # Get the ID number from the input
+        num = self._nodeNum(id)
+            
+        # Get the K closest nodes from the appropriate bucket
         i = self._bucketIndexForInt(num)
-        
-        # If this node is already in our table then return it
-        try:
-            index = self.buckets[i].l.index(num)
-        except ValueError:
-            pass
-        else:
-            return [self.buckets[i].l[index]]
-            
-        # Don't have the node, get the K closest nodes from the appropriate bucket
-        nodes = nodes + self.buckets[i].l
+        nodes = list(self.buckets[i].l)
         
         # Make sure we have enough
         if len(nodes) < self.config['K']:
@@ -156,7 +156,7 @@ class KTable:
         if node.id == self.node.id: return
 
         # Get the bucket for this node
-        i = self. _bucketIndexForInt(node.num)
+        i = self._bucketIndexForInt(node.num)
 
         # Check to see if node is in the bucket already
         try:
@@ -212,11 +212,18 @@ class KTable:
         @rtype: C{datetime.datetime}
         @return: the old lastSeen time of the node, or None if it's not in the table
         """
+        # Get the bucket number
+        num = self._nodeNum(id)
+        i = self._bucketIndexForInt(num)
+
+        # Check to see if node is in the bucket
         try:
[…]

### 2008-03-08 apt_p2p_Khashmir/ktable.py
Fixed justSeenNode in KTable to update the bucket properly.
It now moves the seen node to the end and marks the bucket as
having been updated.

@@ -226,6 +226,12 @@ class KTable:
             n = self.buckets[i].l[it]
             tstamp = n.lastSeen
             n.updateLastSeen()
+            
+            # Move the node to the end and touch the bucket
+            del(self.buckets[i].l[it])
+            self.buckets[i].l.append(n)
+            self.buckets[i].touch()
+            
             return tstamp
     
     def invalidateNode(self, n):

### 2008-04-26 test.py
Update the config file for the change in default key expiry length.

@@ -466,7 +466,7 @@ UNLOAD_PACKAGES_CACHE = 5m
 
 # Refresh the DHT keys after this much time has passed.
 # This should be a time slightly less than the DHT's KEY_EXPIRE value.
-KEY_REFRESH = 57m
+KEY_REFRESH = 2.5h
 
 # The user name to try and run as (leave blank to run as current user)
 USERNAME = 
@@ -513,7 +513,7 @@ MIN_PING_INTERVAL = 15m
 BUCKET_STALENESS = 1h
 
 # expire entries older than this
-KEY_EXPIRE = 1h
+KEY_EXPIRE = 3h
 
 # Timeout KRPC requests to nodes after this time.
 KRPC_TIMEOUT = 14s

### 2008-04-26 apt_p2p/apt_p2p_conf.py
Update the version number and changelog for the initial release.

@@ -23,7 +23,7 @@ class ConfigError(Exception):
     def __str__(self):
         return repr(self.message)
 
-version = versions.Version('apt-p2p', 0, 0, 0)
+version = versions.Version('apt-p2p', 0, 1, 0)
 
 # Set the home parameter
 home = os.path.expandvars('${HOME}')

### 2008-04-26 apt-p2p.py
Fix some documentation errors.

@@ -1,13 +1,14 @@
-#!/usr/bin/env python
+#!/usr/bin/python
 
-# Load apt-p2p application
-#
-# There are two ways apt-p2p can be started:
-#  1. twistd -y apt-p2p
-#     - twistd will load this file and execute the app
-#       in 'application' variable
-#  2. from command line
-#     - __name__ will be '__main__'
+"""Load the apt-p2p application.
+
+There are two ways apt-p2p can be started:
+  1. twistd -y apt-p2p
+     - twistd will load this file and execute the app
+       in 'application' variable
+  2. from command line
+     - __name__ will be '__main__'
+"""
 
 import pwd,sys
 

### 2008-04-26 apt_p2p/HTTPServer.py

@@ -90,7 +90,7 @@ class FileDownloader(static.File):
 class PiecesUploader(static.Data):
     """Modified to identify it for peer requests.
     
-    Uses the modified L{Streams.PieceUploadStream} to stream the pieces for throttling.
+    Uses the modified L{Streams.PiecesUploadStream} to stream the pieces for throttling.
     """
 
     def render(self, req):

### 2008-04-26 apt_p2p/apt_p2p.py

@@ -179,8 +179,6 @@ class AptP2P(protocol.Factory):
         @param url: the URI of the actual mirror request
         @type orig_resp: L{twisted.web2.http.Response}
         @param orig_resp: the response from the cache to be sent to apt
-        @rtype: L{twisted.internet.defer.Deferred}
-        @return: a deferred that will be called back with the correct response
         """
         log.msg('Checking if %s is still fresh' % url)
         modtime = orig_resp.headers.getHeader('Last-Modified')

### 2008-04-26 apt_p2p/stats.py

@@ -34,8 +34,8 @@ class StatsLogger:
     def __init__(self, db):
         """Initialize the statistics.
         
-        @type store: L{db.DB}
-        @param store: the database for the Apt-P2P downloader
+        @type db: L{db.DB}
+        @param db: the database for the Apt-P2P downloader
         """
         # Database
         self.db = db

### 2008-04-26 test.py

@@ -1,4 +1,4 @@
-#!/usr/bin/env python
+#!/usr/bin/python
 
 """Automated tests of the apt-p2p functionality.
 
@@ -29,7 +29,8 @@ the apt-p2p program.
 @var CWD: the working directory the script was run from
 @type apt_conf_template: C{string}
 @var apt_conf_template: the template to use for the apt.conf file
-
+@type apt_p2p_conf_template: C{string}
+@var apt_p2p_conf_template: the template to use for the apt-p2p.conf file
 """
 
 from time import sleep, time

### 2008-04-26 setup.py
Remove the changelog parsing from setup.py and just hardcode the version.

@@ -1,19 +1,11 @@
-#!/usr/bin/env python
+#!/usr/bin/python
 
 import sys
 from distutils.core import setup
 
-from debian_bundle import changelog
-
-f = open('debian/changelog')
-data = f.read()
-f.close()
-c = changelog.Changelog(file=data, max_blocks=1)
-del data
-
 setup(
     name = "apt-p2p",
-    version = c.full_version,
+    version = "0.1.0",
     author = "Cameron Dale",
     author_email = "<camrdale@gmail.com>",
     url = "http://www.camrdale.org/apt-p2p.html",

### 2008-04-26 apt_p2p/policies.py
Clean up the copyrights mentioned in the code.

@@ -1,7 +1,3 @@
-# -*- test-case-name: twisted.test.test_policies -*-
-# Copyright (c) 2001-2007 Twisted Matrix Laboratories.
-# See LICENSE for details.
-
 
 """
 Resource limiting policies.

### 2008-04-26 apt_p2p_Khashmir/actions.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2004 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """Details of how to perform actions on remote peers."""
 

### 2008-04-26 apt_p2p_Khashmir/khash.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """Functions to deal with hashes (node IDs and keys).
 

### 2008-04-26 apt_p2p_Khashmir/khashmir.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2004 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """The main Khashmir program."""
 

### 2008-04-26 apt_p2p_Khashmir/knode.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2004 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """Represents a khashmir node in the DHT."""
 

### 2008-04-26 apt_p2p_Khashmir/krpc.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """The KRPC communication protocol implementation.
 

### 2008-04-26 apt_p2p_Khashmir/ktable.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """The routing table and buckets for a kademlia-like DHT.
 

### 2008-04-26 apt_p2p_Khashmir/node.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """Represents a node in the DHT.
 

### 2008-04-26 apt_p2p_Khashmir/util.py

@@ -1,5 +1,3 @@
-## Copyright 2002-2003 Andrew Loewenstern, All Rights Reserved
-# see LICENSE.txt for license information
 
 """Some utitlity functions for use in apt-p2p's khashmir DHT."""
 

### 2008-04-26 apt_p2p/Hash.py
Fix: no files were added to the DHT due to missing hashing results.

@@ -200,6 +200,7 @@ class HashObject:
         
     def verify(self):
         """Verify that the added file data hash matches the expected hash."""
+        self.digest()
         if self.result is None and self.fileHash is not None and self.expHash is not None:
             self.result = (self.fileHash == self.expHash and self.size == self.expSize)
         return self.result

### 2008-04-28 apt_p2p_Khashmir/DHT.py
Don't add local IP addresses to the routing table (with config option to override).

@@ -115,7 +115,7 @@ class DHT:
                        'KRPC_TIMEOUT', 'KRPC_INITIAL_DELAY']:
                 self.config[k] = self.config_parser.gettime(section, k)
             # The booleans in the config file
-            elif k in ['SPEW']:
+            elif k in ['SPEW', 'LOCAL_OK']:
                 self.config[k] = self.config_parser.getboolean(section, k)
             # Everything else is a string
             else:
@@ -335,7 +335,7 @@ class TestSimpleDHT(unittest.TestCase):
     DHT_DEFAULTS = {'PORT': 9977,
                     'CHECKPOINT_INTERVAL': 300, 'CONCURRENT_REQS': 4,
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
-                    'MAX_FAILURES': 3,
+                    'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
                     'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': False, }
@@ -456,7 +456,7 @@ class TestMultiDHT(unittest.TestCase):
     DHT_DEFAULTS = {'PORT': 9977,
                     'CHECKPOINT_INTERVAL': 300, 'CONCURRENT_REQS': 4,
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
-                    'MAX_FAILURES': 3,
+                    'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
                     'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': False, }

### 2008-04-28 apt_p2p_Khashmir/khashmir.py

@@ -1,5 +1,9 @@
 
-"""The main Khashmir program."""
+"""The main Khashmir program.
+
+@var isLocal: a compiled regular expression suitable for testing if an
+    IP address is from a known local or private range
+"""
 
 import warnings
 warnings.simplefilter("ignore", DeprecationWarning)
@@ -8,7 +12,7 @@ from datetime import datetime, timedelta
 from random import randrange, shuffle
 from sha import sha
 from copy import copy
-import os
+import os, re
 
 from twisted.internet.defer import Deferred
 from twisted.internet import protocol, reactor
@@ -22,6 +26,11 @@ from khash import newID, newIDInRange
 from actions import FindNode, FindValue, GetValue, StoreValue
 from stats import StatsLogger
 import krpc
+
+isLocal = re.compile('^(192\.168\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(10\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?([1][6-9])|([2][0-9])|([3][0-1])\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(127\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})$')
 
 class KhashmirBase(protocol.Factory):
     """The base Khashmir class, with base functionality and find node, no key-value mappings.
@@ -191,6 +200,10 @@ class KhashmirBase(protocol.Factory):
         @param contacted: whether the new node is known to be good, i.e.
             responded to a request (optional, defaults to True)
         """
+        # Don't add any local nodes to the routing table
+        if not self.config['LOCAL_OK'] and isLocal.match(node.host):
+            return
+
         old = self.table.insertNode(node, contacted=contacted)
         if (old and old.id != self.node.id and
             (datetime.now() - old.lastSeen) > 
@@ -517,7 +530,7 @@ class SimpleTests(unittest.TestCase):
     DHT_DEFAULTS = {'PORT': 9977,
                     'CHECKPOINT_INTERVAL': 300, 'CONCURRENT_REQS': 4,
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
-                    'MAX_FAILURES': 3,
[…]

### 2008-04-28 test.py

@@ -513,6 +513,11 @@ MIN_PING_INTERVAL = 15m
 # refresh buckets that haven't been touched in this long
 BUCKET_STALENESS = 1h
 
+# Whether it's OK to add nodes to the routing table that use an IP
+# address from a known local/private range.
+# If not specified here, the LOCAL_OK value in the DEFAULT section will be used.
+LOCAL_OK = yes
+
 # expire entries older than this
 KEY_EXPIRE = 3h
 

### 2008-04-29 apt_p2p_Khashmir/actions.py
Don't try and add ourself to the routing table during an action.

@@ -184,7 +184,8 @@ class ActionBase:
 
     def gotResponse(self, dict, node, expected_results, df):
         """Receive a response from a remote node."""
-        reactor.callLater(0, self.caller.insertNode, node)
+        if node.id != self.caller.node.id:
+            reactor.callLater(0, self.caller.insertNode, node)
         if self.finished or self.answered.has_key(node.id):
             # a day late and a dollar short
             return
@@ -197,7 +198,8 @@ class ActionBase:
     def actionFailed(self, err, node, expected_results, df):
         """Receive an error from a remote node."""
         log.msg("action %s failed on %s/%s: %s" % (self.action, node.host, node.port, err.getErrorMessage()))
-        self.caller.table.nodeFailed(node)
+        if node.id != self.caller.node.id:
+            self.caller.table.nodeFailed(node)
         self.outstanding -= 1
         self.outstanding_results -= expected_results
         self.schedule()

### 2008-04-29 apt_p2p_Khashmir/DHT.py
Only add nodes to the routing table that have responded to a request.
If they haven't, ping them before adding them.
Still ping the old node first to see if it's stale.
Move the temporary functions for ping/join to be module functions.

@@ -338,7 +338,7 @@ class TestSimpleDHT(unittest.TestCase):
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
                     'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
-                    'KEY_EXPIRE': 3600, 'SPEW': False, }
+                    'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def setUp(self):
         self.a = DHT()
@@ -459,7 +459,7 @@ class TestMultiDHT(unittest.TestCase):
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
                     'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
-                    'KEY_EXPIRE': 3600, 'SPEW': False, }
+                    'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def setUp(self):
         self.l = []

### 2008-04-29 apt_p2p_Khashmir/khashmir.py

@@ -164,7 +164,7 @@ class KhashmirBase(protocol.Factory):
             (optional, defaults to doing nothing with the results)
         @type errback: C{method}
         @param errback: the method to call if an error occurs
-            (optional, defaults to calling the callback with None)
+            (optional, defaults to calling the callback with the error)
         """
         n = self.Node(NULL_ID, host, port)
         self.sendJoin(n, callback=callback, errback=errback)
@@ -190,7 +190,7 @@ class KhashmirBase(protocol.Factory):
         
         If all you have is a host/port, then use L{addContact}, which calls this
         method after receiving the PONG from the remote node. The reason for
-        the seperation is we can't insert a node into the table without its
+        the separation is we can't insert a node into the table without its
         node ID. That means of course the node passed into this method needs
         to be a properly formed Node object with a valid ID.
 
@@ -202,30 +202,57 @@ class KhashmirBase(protocol.Factory):
         """
         # Don't add any local nodes to the routing table
         if not self.config['LOCAL_OK'] and isLocal.match(node.host):
+            log.msg('Not adding local node to table: %s/%s' % (node.host, node.port))
             return
-
+        
         old = self.table.insertNode(node, contacted=contacted)
-        if (old and old.id != self.node.id and
+
+        if (isinstance(old, self._Node) and old.id != self.node.id and
             (datetime.now() - old.lastSeen) > 
              timedelta(seconds=self.config['MIN_PING_INTERVAL'])):
-            
-            def _staleNodeHandler(err, oldnode = old, newnode = node, self = self, start = datetime.now()):
-                """The pinged node never responded, so replace it."""
-                log.msg("action ping failed on %s/%s: %s" % (oldnode.host, oldnode.port, err.getErrorMessage()))
-                self.stats.completedAction('ping', start)
-                self.table.replaceStaleNode(oldnode, newnode)
-            
-            def _notStaleNodeHandler(dict, old = old, self = self, start = datetime.now()):
-                """Got a pong from the old node, so update it."""
-                self.stats.completedAction('ping', start)
-                if dict['id'] == old.id:
-                    self.table.justSeenNode(old.id)
             
             # Bucket is full, check to see if old node is still available
             self.stats.startedAction('ping')
             df = old.ping(self.node.id)
[…]

### 2008-04-29 apt_p2p_Khashmir/ktable.py

@@ -173,6 +173,7 @@ class KTable:
             # Remove the stale node
             del(self.buckets[i].l[it])
             removed = True
+            log.msg('Removed node from routing table: %s/%s' % (stale.host, stale.port))
         
         # Insert the new node
         if new and self._bucketIndexForInt(new.num) == i and len(self.buckets[i].l) < K:
@@ -183,7 +184,8 @@ class KTable:
     def insertNode(self, node, contacted = True):
         """Try to insert a node in the routing table.
         
-        This inserts the node, returning None if successful, otherwise returns
+        This inserts the node, returning True if successful, False if the
+        node could have been added if it responds to a ping, otherwise returns
         the oldest node in the bucket if it's full. The caller is then
         responsible for pinging the returned node and calling replaceStaleNode
         if it doesn't respond. contacted means that yes, we contacted THEM and
@@ -194,11 +196,13 @@ class KTable:
         @type contacted: C{boolean}
         @param contacted: whether the new node is known to be good, i.e.
             responded to a request (optional, defaults to True)
-        @rtype: L{node.Node}
-        @return: None if successful (the bucket wasn't full), otherwise returns the oldest node in the bucket
+        @rtype: L{node.Node} or C{boolean}
+        @return: True if successful (the bucket wasn't full), False if the
+            node could have been added if it was contacted, otherwise
+            returns the oldest node in the bucket
         """
         assert node.id != NULL_ID
-        if node.id == self.node.id: return
+        if node.id == self.node.id: return True
 
         # Get the bucket for this node
         i = self._bucketIndexForInt(node.num)
@@ -219,16 +223,18 @@ class KTable:
                 # utilizing this nodes new contact info
                 self.buckets[i].l.append(node)
                 self.buckets[i].touch()
-            return
+            return True
         
         # We don't have this node, check to see if the bucket is full
         if len(self.buckets[i].l) < K:
             # Not full, append this node and return
             if contacted:
                 node.updateLastSeen()
-            self.buckets[i].l.append(node)
-            self.buckets[i].touch()
-            return
+                self.buckets[i].l.append(node)
+                self.buckets[i].touch()
+                log.msg('Added node to routing table: %s/%s' % (node.host, node.port))
+                return True
+            return False
             
         # Bucket is full, check to see if the local node is not in the bucket
         if not (self.buckets[i].min <= self.node < self.buckets[i].max):

### 2008-04-29 docs/abstract/abstract.tex
Create the Kile project for the SIGCOMM poster abstract.

--- 2008-03-06 docs/motivation/motivation.tex
@@
 Burnaby, British Columbia, Canada\\
-Email: camerond@cs.sfu.ca}}
+Email: camerond@cs.sfu.ca}
+\and
+\IEEEauthorblockN{Jiangchuan Liu}
+\IEEEauthorblockA{School of Computing Science\\
+Simon Fraser University\\
+Burnaby, British Columbia, Canada\\
+Email: jcliu@cs.sfu.ca}}
 
@@
 the distribution of Debian software packages.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-\label{intro}
 
@@
 opportunity that can be met with such a system.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Related Work}
-\label{related}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Situation}
-\label{situation}
-
 There are a large number of groups using the Internet to distribute
@@
 hashes, to verify the completed downloads accuracy or authenticity.
-
-\subsection{Examples}
-\label{examples}
 
@@
 
-\subsection{Similarities}
-\label{similarities}
-
-The important things to note for each of the systems mentioned in
-section \ref{examples}, is that they all have the following in
+The important things to note for each of the systems mentioned
+is that they all have the following in
 common:
@@
 
-\subsection{Differences}
-\label{differences}
-
 Although at first it might seem that a single all-reaching solution
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Opportunity}
-\label{opportunity}
-
-The situation described in section \ref{situation} presents a clear
+The situation presents a clear
 opportunity to use some form of peer-to-peer file-sharing protocol.
@@
 
-\subsection{Implementation Options}
-\label{imp_options}
-
 There are several ways to implement the desired P2P functionality
@@
 possible.
-
-\subsection{Downloading From Peers}
-\label{downloading}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Sample Implementation}
-\label{implementation}
-
 A sample implementation has been created that functions as described
-in section \ref{opportunity}. This software, called
+previously. This software, called
 \texttt{apt-p2p}, interacts with the \texttt{apt} tool found in most
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{apt_p2p_simulation-size_CDF.eps}
-\caption{The CDF of the size of packages in a Debian system, both
-for the actual size and adjusted size based on the popularity of
-the package.}
-\label{size_CDF}
-\end{figure}
-
-Figure \ref{size_CDF} shows the package size of the 22,298 packages
+Looking at the package size of the 22,298 packages
 available in Debian in January 2008. We can see that most of the
@@
 
-\subsection{Current Status}
-\label{status}
-
 Though functional and useful, the current implementation is not
@@
 
-However, the current implementation does not use the sub-piece
-hashing described in section \ref{downloading}, or any parallelizing
-of downloads. Instead, it currently chooses a peer at random from
-the list of possible peers, and downloads the entire file from that
-peer, regardless of the size of the file. This needs to change if
-both a) the file is large (more than 512 KB), and b) there are
-multiple peers with the file. The storage and retrieval of sub-piece
-hashes has not yet been finalized though, and until it is this final
-step is not possible.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Future Work}
-\label{future}
-
-There are some additional steps to be taken to succeed with this
-project. First, the storage method of sub-piece hashes needs to be
-finalized and properly implemented. The communication built into the
-Khashmir-based DHT also needs to be solidified. Some additional peer
-statistics data needs to be made available for each peer, so that
-some form of analysis is available once the system is deployed. The
-reference implementation then needs to be packaged for upload to the
-Debian package archive, so that it can benefit from the widest
-possible deployment among users.
-
 Finally, once deployed the system needs to be analyzed to determine
@@
 \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusions}
-\label{conclusions}
 
@@
 
+\end{abstract}
+
 \bibliographystyle{IEEEtran}

### 2008-04-29 apt_p2p_Khashmir/krpc.py
Don't log KRPC errors for receiving an invalid store_value token.

@@ -366,7 +366,8 @@ class KRPC:
                     ret = f(*(), **msg[ARG])
                 except KrpcError, e:
                     log.msg('Got a Krpc error while running: krpc_%s' % msg[REQ])
-                    log.err(e)
+                    if e[0] != KRPC_ERROR_INVALID_TOKEN:
+                        log.err(e)
                     self.stats.errorAction(msg[REQ])
                     olen = self._sendResponse(msg[REQ], addr, msg[TID], ERR,
                                               [e[0], e[1]])

### 2008-05-01 docs/abstract/abstract.tex
Massive update to most of the abstract.

@@
 
-\title{Using Altruistic Peers to Help With Free Web Downloads}
+\title{Using Altruistic Peers to Help With Free Content Downloads}
 \author{\IEEEauthorblockN{Cameron Dale}
@@
 
-\begin{abstract}
-A large amount of free content is available over the Internet from
-many different distributors. Most of this content uses the traditional
-client-server model to handle requests from users. However, there is an
-opportunity to use peer-to-peer techniques to reduce the cost of
-much of this distribution, especially due to the altruistic nature
-of many of these users. We present a general technique for
-satisfying the needs of this P2P distribution, which is applicable
-to many of these distributors systems. Our method makes use of a DHT
-for storing the location of peers, using the cryptographic hash of
-the file as a key. We then go further and implement a solution for
-the distribution of Debian software packages.
+% \begin{abstract}
+% A large amount of free content is available over the Internet from
+% many different distributors. Most of this content uses the traditional
+% client-server model to handle requests from users. However, there is an
+% opportunity to use peer-to-peer techniques to reduce the cost of
+% much of this distribution, especially due to the altruistic nature
+% of many of these users. We present a general technique for
+% satisfying the needs of this P2P distribution, which is applicable
+% to many of these distributors systems. Our method makes use of a DHT
+% for storing the location of peers, using the cryptographic hash of
+% the file as a key. We then go further and implement a solution for
+% the distribution of Debian software packages.
+% 
+% We have designed a generally applicable peer-to-peer content
+% distribution model to be used by many of the free content
+% distributors operating today. It makes use of already existing
+% features of most package management systems, to create a
+% peer-to-peer distribution that should substantially reduce the costs
+% of hosting the content.
+% 
+% We have also implemented our design in software to be used in
+% conjuction with Debian-based distribution of Linux software
+% packages. It will soon be in use by many users of the Debian
+% project's distribution, and so serves as an example of the
+% opportunities that exist.
+% \end{abstract}
 
-There are a large number of free content distributors using
+\section{Introduction}
+\label{intro}
+
+There are a large number of free content distributors using custom
 package distribution systems over the Internet to distribute content
 to their users. These distributors have developed many different
-methods for this distribution, but they almost exclusively use a client-server
-model to satisfy user requests. The popularity and number of users
-results in a large number of requests, which usually requires a
-network of mirrors to handle. Due to the free nature of this
-content, many users are willing and able to contribute upload
-bandwidth to this distribution, but have no current way to do this.
+methods for this distribution, but they almost exclusively use a
+client-server model to satisfy user requests. The contents' size,
+popularity and number of users results in a large number of
+requests, which usually requires a network of mirrors to handle. Due
+to the free nature of this content, many users are willing and able
+to contribute upload bandwidth to this distribution, but currently
+have no way to do this.
+
+There are many peer-to-peer implementations available today, but
+none is very well suited to this specific problem. Many distributors
+make their content available in some form using BitTorrent, but it
+is not ideally suited to an application such as this one. BitTorrent
+is designed for large static collections of files, in which all
+peers are interested in most or all of the content. But the packages
+in this case are small, constantly being updated, and users are only
+interested in a small subset of them. Other proposed implementations
+suffer from some of the same, and other, problems, and so none has
+been well-used to address this problem.
 
 We present a new peer-to-peer distribution model to meet these
-demands. It is based on many previous implementations of successful
-peer-to-peer protocols, especially distributed hash tables (DHT) and
+demands. It is based on the lessons learned from many previous
+implementations of successful peer-to-peer protocols, and makes use
+of a distributed hash table (DHT) and borrows many ideas from
 BitTorrent. The model relies on the pre-existence of cryptographic
-hashes of the content, which should uniquely identify it for a
-request from other peers. If the peer-to-peer download fails, then
-the original request to the server is used as a fallback to prevent
-any dissatisfaction from users. The peer can then share this new
-content with others through the P2P system.
+hashes of the content, which uniquely identify it for a request from
+other peers. If the peer-to-peer download fails, then the original
+request to the server is used as a fallback to prevent dissatisfied
+users. The peer can then share this new content with others through
+the P2P system. This system would efficiently satisfy the demands of
+a large number of users, and should significantly reduce the
+currently substantial bandwidth costs of hosting this content.
 
-First, we examine the opportunity that is available for many of
-these free content distributors. We present an overview of a system
-that would efficiently satisfy the demands of a large number of
-users, and should significantly reduce the currently substantial
-bandwidth requirements of hosting this content. We then present an
-example implementation based on the Debian package distribution
-system. This implementation will be used by a large number of users,
-and serves as an example for other free content distributors of the
-opportunity that can be met with such a system.
+We also present a sample implementation based on the Debian package
+distribution system. This implementation is simple, makes use of
+mostly well-known technologies, and so serves as an example for
+other free content distributors of the opportunity that can be met
+with such a system.
 
-There are many peer-to-peer implementations available today, but
-none is very well suited to this specific problem.
-
-Many distributors make their content available in some form using
-BitTorrent \cite{COHEN03}, e.g. for the distribution of CD
-images. This is not an ideal situation though, as it requires the
-peers to download large amounts of content that they are not
-interested in, and prevents them from updating to new content
-without downloading another image containing a lot of the same
-content they already have. There have also been other attempted
-implementations, usually based on some form of BitTorrent, to
-address this problem. Unfortunately, BitTorrent is not ideally
-suited to an application such as this one, for the following
-reasons:
-\begin{itemize}
- \item The packages are too small and there are too many to create
-       individual torrents for each one.
- \item All the packages together are too large to track efficiently
-       as a single torrent.
- \item BitTorrent's piece sizes are bigger than many of the
-       packages, which wastes bandwidth downloading parts of other
-       packages.
- \item A small number of the packages can be updated every day,
-       requiring a new torrent and splitting the
-       download population (even though peers in the new torrent
-       share 99\% of the files in common with peers in the old
-       torrent).
- \item Incentives to share (upload) are no longer needed, as the
-       content is freely available for anyone to download without
-       sharing (seeds are also not needed).
-\end{itemize}
-
-Some other implementations have used BitTorrent almost unmodified,
-while others have only looked at using DHTs to replace the tracker
-in a BitTorrent system. apt-torrent \cite{apttorrent} creates
-torrents for some of the larger content available, but it ignores
-the smaller content, which is often the most popular.
-DebTorrent \cite{debtorrent} makes widespread modifications to a
-traditional BitTorrent client, but also requires some changes to the
-distribution system to support it, and separates peers into groups
-based on their interest which limits the effectiveness of the
-sharing. Kenosis \cite{kenosis} is a P2P Remote Procedure Call
-client also based on the Kademlia DHT, but it is meant as a P2P
-primitive system on which other tools can be built, and so it has no
-file sharing functionality at all. Many have used a DHT as a drop-in
-replacement for the tracker in a BitTorrent system
-\cite{bittorrent-dht, azureus-dht}, but such systems only use the
-DHT to find peers for the same torrent, so the file sharing uses
-traditional BitTorrent and so is not ideal for the reasons listed
-above.
+\section{Problem}
 
 There are a large number of groups using the Internet to distribute
-their free content. This content is usually available from a free
+their free content. This content is made available from a free
 download site, which usually requires a network of mirrors to
 support the large number of requests. This content almost always
-supports some kind of file verification, usually cryptographic
-hashes, to verify the completed downloads accuracy or authenticity.
+supports some kind of file verification, using cryptographic hashes,
+to verify the completed downloads accuracy or authenticity.
 
 Most Linux distributions use a software package management system
-that fetches packages to be installed from a network of mirrors. The
-Debian project \cite{debian} (and other Debian-based distributions)
-uses the \texttt{apt} program, which downloads Debian packages in
-the \texttt{.deb} format from one of many mirrors. The program will
-first download index files that contain a listing of which packages
-are available, as well as important information such as their size,
-location, and a hash of their content. Red Hat's Fedora project
-\cite{fedora} (and other RPM-based distributions) use the
-\texttt{yum} program to obtain RPMs, and Gentoo \cite{gentoo} uses
-\texttt{portage} in a similar way.
+that fetches packages to be installed from a network of mirrors.
+Debian-based distributions uses the \texttt{apt} program, which
+downloads Debian packages from one of many mirrors. RPM-based
+distributions use \texttt{yum}, and Gentoo uses \texttt{portage},
+which both operate in a similar way. Other free software
+distributors also use a similar system: CPAN distributes files
+containing software packages for the PERL programming language,
+using SOAP RPC requests to find and download files; Cygwin provides
+many of the standard Unix/Linux tools in a Windows environment,
+using a package management tool that requests packages from
+websites; there are two software distribution systems for Mac OSX,
+fink and MacPorts, that also retrieve packages in this way. Also,
+some systems use direct web downloads, but with a hash verification
+file also available for download next to the desired file. These
+hash files usually have the same file name, but with an added
+extension identifying the hash used (e.g. \texttt{.md5} for the MD5
+hash). Finally, there are other programs that make use of
+cryptographic hashing to identify files: e.g. Git is a version
+control system in which all files and changes are identified by
+their SHA1 hash.
 
-Other software vendors also use a similar system. CPAN \cite{cpan}
-distributes files containing software packages for the PERL
-programming language, using SOAP RPC requests to find and download
-files. Cygwin \cite{cygwin} provides many of the
-standard Unix/Linux tools in a Windows environment, using a
-package management tool that requests packages from websites. There
-are two software distribution systems for Mac OSX, fink and
-MacPorts, that also retrieve packages in this way.
+These systems all share some commonalities: the content is avaliable
+for anyone to freely download, the content is divided into distinct
+units (packages), users are typically not interested in downloading
+all of the content available, and hashes of the content are
+available before the download is attempted. We also expect that
+there are a number of users of these systems that are motivated by
+altruism to want to help out with this distribution. This is common
+in these systems due to the free nature of the content being
+delivered.
 
-Also, some systems use direct web downloads, but with a hash
-verification file also available for download next to the desired
-file. These hash files usually have the same file name, but with an
-added extension identifying the hash used (e.g. \texttt{.md5} for
-the MD5 hash).
+% Although at first it might seem that a single all-reaching solution
+% is possible for this problem, there are some differences in each
+% system that require independent solutions. The systems all use
+% different tools for their distribution, so any implementation would
+% have to be specific to the tool it is to integrate with. In
+% particular, how to obtain requests from the tool or user, and how to
+% find a hash that corresponds to the file being requested, is very
+% specific to each system. Also, there may be no benefit in creating a
+% single large solution to integrate all these systems since, even
+% though they sometimes distribute nearly identical content, it is not
+% identical as it has been tailored to the system, which will change
+% the hash of the files and so will make it impossible to distribute
+% similar content across systems.
 
-Finally, there are other programs that use cryptographic hashing to
-identify files. Git is a version control system in which all files,
-commits, and tags, are identified by their SHA1 hash. These hashes
-are used to verify the origin of these items, but are also used when
-clients update their local files with new remote information.
+\section{Solution}
 
-The important things to note for each of the systems mentioned
-is that they all have the following in
-common:
-\begin{itemize}
- \item The content is avaliable for anyone to freely download.
- \item The content is divided into distinct units (packages), each of which contains
-       a small independent part of all the content.
- \item Users are typically not interested in downloading all of the content
-       available.
- \item Hashes of the content and its size are available before the
-       download is attempted.
- \item Requests for the content are sent by a tool, not directly by
-       the user (though the tool is responding to requests from the user).
-\end{itemize}
+This situation presents a clear opportunity to use some form of
+peer-to-peer file-sharing protocol. The sparse interest in a large
+number of packages undergoing constant updating is well suited to
+the functionality provided by a Distributed Hash Table (DHT). DHTs
+require unique keys to store and retrieve strings of data, which the
+cryptographic hashes used by these package management systems are
+perfect for. The stored and retrieved strings can then be pointers
+to the peers that have the content package that hashes to that key.
+A downloading peer can lookup the package hash in the DHT and, if it
+is found, download the file from those peers and verify the content.
+If the package hash can not be found in the DHT, the peer will
+fallback to downloading from the original content location, and once
+complete will add a new entry to the DHT indicating that it now has
+the content. The servers or mirrors thus act as \emph{seeds} for the
+P2P system without any modification to them.
 
-We also expect that there are a number of users of these systems
-that are motivated by altruism to want to help out with this
-distribution. This is common in these systems due to the free nature
-of the content being delivered, which encourages some to want to
-help out in some way. A number of the systems are also used by
-groups that are staffed mostly, or sometimes completely, by
-volunteers. This also encourages users to want to \emph{give back}
-to the volunteer community that has created the content they are
-using.
+This desired P2P functionality could be integrated into the existing
+package management software in 2 ways. The functionality can be
+directly integrated into the software, though this could be
+difficult as the DHT should be running at all times, whereas the
+tools typically only run until the download request is complete.
+Alternatively, since many of the package management software
+implementations use HTTP requests to download the files, this makes
+it possible to implement the P2P aspect as a standard HTTP caching
+proxy, which will get uncached requests first from the P2P system,
+and then fallback to the normal HTTP request from a server. For
+methods that don't use HTTP requests, other types of proxies may
+also be possible.
 
-Although at first it might seem that a single all-reaching solution
-is possible for this situation, there are some differences in each
-system that require independent solutions. The systems all use
-different tools for their distribution, so any implementation would
-have to be specific to the tool it is to integrate with. In
-particular, how to obtain requests from the tool or user, and how to
-find a hash that corresponds to the file being requested, is very
-specific to each system.
+Downloading a large file efficiently from a number of peers is where
+BitTorrent shines as a peer-to-peer application. Its method of
+breaking up larger files into pieces, each with its own hash, makes
+it very easy to parallelize the downloading process and maximize the
+download speed. Since the package management system only makes
+available a hash of the entire content, the piece hashes will have
+to be found through the P2P protocol. These hashes can be added to
+the DHT, or requested from other peers.
 
-Also, there may be no benefit in creating a single large solution to
-integrate all these problems. For one, though they sometimes
-distribute nearly identical content (e.g. the same software
-available in multiple Linux distributions), it is not exactly
-identical and has usually been tailored to the system. The small
-differences will change the hash of the files, and so
-will make it impossible to distribute similar content
-across systems. And, although peer-to-peer systems scale very well
-with the number of peers in the system, there is some overhead
-involved, so having a much larger system of peers would mean that
-requests could take longer to complete.
+For files smaller than the piece size, no piece hashes are
+necessary. For medium size files of only a few pieces, the piece
+hashes are short enough to be stored in the DHT with the pointer to
+the peer to download from. For large files of 10's of pieces, the
+piece hashes are generally too long to store with the pointer to
+download peers. Instead, the peers will store the piece hashes for
+large files separately in the DHT using as a key the hash of the
+piece hashes, and include this key in their stored value for the
+hash of the file. Peers downloading these large files can then
+retrieve the piece hashes using a second DHT request. For very large
+files for which the piece hashes are too large to store at all in
+the DHT, a request to the peer for the hash (using the same method
+as file downloads) should return the piece hashes.
 
-The situation presents a clear
-opportunity to use some form of peer-to-peer file-sharing protocol.
-This sparse interest in a large number of packages, and constant
-updating, is well suited to the functionality provided by a
-Distributed Hash Table (DHT). DHTs require unique keys to store and
-retrieve strings of data, for which the cryptographic hashes used by
-these package management systems are perfect for. The stored and
-retrieved strings can then be pointers to the peers that have the
-content package that hashes to that key. A downloading peer can
-lookup the package hash in the DHT and, if it is found,
-download the file from those peers and verify the content. If the
-package hash can not be found in the DHT, the peer will fallback to
-downloading from the original content location (i.e. the network of
-mirrors), and once complete will add a new entry to the DHT
-indicating that it has the content.
+\section{Sample Implementation}
 
-There are several ways to implement the desired P2P functionality
-into the existing package management software. The functionality can
-be directly integrated into the software, though this can be
-difficult as the DHT should be running at all times, both for
-efficient lookups and to support uploading of already downloaded
-content, whereas the tools typically only run until the download request is complete.
-Many of the package management software implementations use
-HTTP requests to download the files, which makes it possible to
-implement the P2P aspect as a standard HTTP caching proxy, which
-will get uncached requests first from the P2P system, and then
-fallback to the normal HTTP request from a server. For methods that
-don't use HTTP requests, other types of proxies may also be
-possible.
+To demonstrate the ease and effectiveness of this solution we have
+created a sample implementation called \texttt{apt-p2p} which
+interacts with the \texttt{apt} tool found in most Debian-based
+Linux distributions. \texttt{apt} uses SHA1 hashes to verify most
+downloaded files, including the large index files that contain
+information such as the hashes of the individual packages. Since all
+requests from \texttt{apt} are in the form of HTTP downloads from a
+server, the implementation takes the form of a caching HTTP proxy.
 
-Downloading a file efficiently from a number of peers is where
-BitTorrent shines as a peer-to-peer application. Its method of
-breaking up larger files into sub-pieces, each with its own hash,
-makes it very easy to parallelize the downloading process and
-maximize the download speed. For very small packages (i.e. less than
-the sub-piece size), this parallel downloading is not necessary, or
-even desirable. However, this method can still be used in
-conjunction with the DHT, for the larger packages that are
-available.
+The DHT used is based on Khashmir, which is an implementation of the
+Kademlia DHT \cite{kademlia} used by most of the existing BitTorrent
+clients to implement trackerless operation. It has been modified to
+better support the retrieval of multiple values per key, and to
+improve the response time so as to satisfy users' demands for speed.
+The values stored in the DHT are all bencoded dictionaries similar
+to torrent files. Peers store their download location (IP address
+and TCP port), as well as the piece strings (or hashes of piece
+strings) as described in the previous section. Downloading is
+accomplished by simple HTTP requests to the peers' identified to
+have the desired file. All peers support HTTP/1.1, both in the
+server and the client, which allows for pipelining of multiple
+requests to a peer, and the requesting of smaller pieces of a large
+file using the Range request header.
 
-Since the package management system only stores a hash of the entire
-content, and not of sub-pieces of that content, we will need to be
-able to store and retrieve these sub-piece hashes using the P2P protocol.
-In addition to storing the file download location in the DHT (which would still
-be used for small files), a peer will store a \emph{torrent string}
-containing the peer's hashes of the sub-pieces of the larger
-files. These piece hashes could be compared ahead of time to
-determine which peers have the same piece hashes (they all should),
-and then used during the download to verify the downloaded pieces.
+We have chosen a piece size of 512 kB, which means that 17,515
+(79\%) of the almost 23 thousand packages will not require piece
+information as they are smaller than a single piece. There are 3054
+packages (13\%) that will require 2 to 4 pieces, for which the piece
+hashes are stored directly with the peer download information in the
+DHT. There are 1667 packages (7\%) that will require 5 to 70 pieces,
+and so use a separate lookup in the DHT for the piece hashes.
+Finally, there are only 62 packages (0.3\%) that require more than
+70 pieces, and so for them a separate request to a peer for the
+piece hashes is needed.
 
-For very large files (5 or more pieces), the torrent strings
-are too long to store in the DHT (a single UDP packet should be less
-than 1472 bytes to avoid fragmentation). Instead, the peers will
-store the torrent string for large files separately in the DHT, and
-only contain a reference to it in their stored value for the hash of
-the file. The reference would be a hash of the torrent string. If
-the torrent string is short enough to store in the DHT (i.e. less
-than 1472 bytes, or 70 pieces for the SHA1 hash), then a
-lookup of that hash in the DHT would give the torrent-like string.
-Otherwise, a request to the peer for the hash (using the same
-method as file downloads), should return the torrent string.
+We have deployed the DHT from this implementation on PlanetLab to
+test it's speed and effectiveness at delivering results. We find
+that, even on the overloaded PlanetLab network, the responses to
+lookups of keys take less than 10 seconds on average. Since each
+package download requires a lookup before it can proceed, it may
+take at least 10 seconds to process a request. However, due to
+\texttt{apt} pipelining up to 10 requests at a time to our proxy,
+the user will see an average response time (after a short startup)
+of under a second, which should be responsive enough to satisfy most
+users.
 
-% This can cause a problem with hash checking the returned data, as
-% hashes for the pieces are not known. Any file that fails a hash
-% check should be downloaded again, with each piece being downloaded
-% from different peers than it was previously. The peers are shifted
-% by 1, so that if a peers previously downloaded piece i, it now
-% downloads piece i+1, and the first piece is downloaded by the
-% previous downloader of the last piece, or preferably a previously
-% unused peer. As each piece is downloaded the running hash of the
-% file should be checked to determine the place at which the file
-% differs from the previous download.
-% 
-% If the hash check then passes, then the peer who originally provided
-% the bad piece can be assessed blame for the error. Otherwise, the
-% peer who originally provided the piece is probably at fault, since
-% he is now providing a later piece. This doesn't work if the
-% differing piece is the first piece, in which case it is downloaded
-% from a 3rd peer, with consensus revealing the misbehaving peer.
-
-A sample implementation has been created that functions as described
-previously. This software, called
-\texttt{apt-p2p}, interacts with the \texttt{apt} tool found in most
-Debian-based Linux distributions. Apt uses SHA1 hashes to
-verify most downloaded files, including the large index files that
-contain the hashes of the individual packages. We chose this
-distribution system as it is familiar to us, and there are
-interesting statistics available for analyzing the popularity of the
-software packages \cite{popcon}.
-
-Since all requests from apt are in the form of HTTP downloads from a
-server, the implementation takes the form of a caching HTTP proxy.
-Making a standard apt implementation use the proxy is then as simple
-as prepending the proxy location and port to the front of the mirror
-name (i.e. ``localhost:9977/mirrorname.debian.org/\ldots'').
-
-The DHT is based on Khashmir \cite{khashmir}, which is an implementation of the
-Kademlia DHT \cite{kademlia} using methods familiar to BitTorrent
-developers. It is the same DHT used by most of the existing
-BitTorrent clients to implement trackerless operation. The
-communication is all handled by UDP messages, and RPC requests are
-bencoded in the same way as torrent files. Khashmir uses the high-level
-Twisted event-driven networking engine \cite{twisted}, so Twisted is also used for
-all other networking needs.
-
-The torrent strings stored in the DHT are all bencoded dictionaries
-containing similar information to what is in a torrent file. This
-includes the piece size used, the length of the piece hash, and of
-course the hashes of all the sub-pieces of the content.
-
-Downloading is accomplished by sending simple HTTP requests to the
-peer's identified (by lookups in the DHT) to have the desired file.
-The HTTP server used for the proxy also doubles as the server
-listening for requests for downloads from other peers. All peers
-support HTTP/1.1, both in the server and the client, which allows
-for pipelining of multiple requests to a peer, and the requesting of
-smaller pieces of a large file using the Range request header.
-
-Looking at the package size of the 22,298 packages
-available in Debian in January 2008. We can see that most of the
-packages are quite small, and so most will therefore not require
-sub-piece information to download. We have chosen a piece
-size of 512 kB, which means that 17,515 (78\%) of the packages will
-not require this information. There are 3054 packages that will
-require 2 to 4 pieces, for which the torrent string can be stored
-directly with the package hash in the DHT. There are 1667 packages
-that will require a separate lookup in the DHT for the longer
-torrent string as they require 5 to 70 pieces. Finally, there are
-only 62 packages that require more than 70 pieces, and so will
-require a separate request to a peer for the torrent string.
-
-Though functional and useful, the current implementation is not
-complete, and is missing some of the downloading details necessary.
-It operates as a caching HTTP proxy, serving
-requests from apt and other peers. Index files are identified by the
-current implementation, and parsed to find the hashes of files for later DHT
-lookups if they are requested. Files are downloaded from
-peers, and those not available in peers are downloaded directly
-from the mirror. Any downloaded files are hash checked to verify
-them, and then added to the DHT.
-
-Finally, once deployed the system needs to be analyzed to determine
-its effectiveness at meeting the goals of this project. The analysis
-will consist of measuring, over time:
-\begin{itemize}
- \item The number of peers using the system.
- \item The amount of downloaded data by peers from other peers.
- \item The amount of downloaded data by peers from mirrors/servers.
- \item The bandwidth used to maintain the DHT.
-\end{itemize}
-
-We have designed a generally applicable peer-to-peer content
-distribution model to be used by many of the free content
-distributors operating today. It makes use of already existing
-features of most package management systems, to create a
-peer-to-peer distribution that should substantially reduce the costs
-of hosting the content.
-
-We have also implemented our design in software to be used in
-conjuction with Debian-based distribution of Linux software
-packages. It will soon be in use by many users of the Debian
-project's distribution, and so serves as an example of the
-possibilities that exist.
-
-\end{abstract}
+This sample implementation has also been made available to all
+Debian users as an installable software package. Once it becomes
+more widely used, we plan to collect more usage information on the
+system through crawls and data analysis.
 

### 2008-05-01 apt_p2p/util.py
Fix the broken local IP regexes.

@@ -12,7 +12,9 @@ from twisted.trial import unittest
 
 isLocal = re.compile('^(192\.168\.[0-9]{1,3}\.[0-9]{1,3})|'+
                      '(10\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})|'+
-                     '(172\.0?([1][6-9])|([2][0-9])|([3][0-1])\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?1[6-9]\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?2[0-9]\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?3[0-1]\.[0-9]{1,3}\.[0-9]{1,3})|'+
                      '(127\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})$')
 
 def findMyIPAddr(addrs, intended_port, local_ok = False):

### 2008-05-01 apt_p2p_Khashmir/khashmir.py

@@ -29,7 +29,9 @@ import krpc
 
 isLocal = re.compile('^(192\.168\.[0-9]{1,3}\.[0-9]{1,3})|'+
                      '(10\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})|'+
-                     '(172\.0?([1][6-9])|([2][0-9])|([3][0-1])\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?1[6-9]\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?2[0-9]\.[0-9]{1,3}\.[0-9]{1,3})|'+
+                     '(172\.0?3[0-1]\.[0-9]{1,3}\.[0-9]{1,3})|'+
                      '(127\.[0-9]{1,3}\.[0-9]{1,3}\.[0-9]{1,3})$')
 
 class KhashmirBase(protocol.Factory):

### 2008-05-02 docs/abstract/abstract.tex
Final abstract for submission.

@@
 
-\title{Using Altruistic Peers to Help With Free Content Downloads}
+\title{Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads}
 \author{\IEEEauthorblockN{Cameron Dale}
-\IEEEauthorblockA{School of Computing Science\\
+\IEEEauthorblockA{School of Computing Science (student)\\
 Simon Fraser University\\
@@
 \IEEEauthorblockN{Jiangchuan Liu}
-\IEEEauthorblockA{School of Computing Science\\
+\IEEEauthorblockA{School of Computing Science (faculty)\\
 Simon Fraser University\\
@@
 
-% \begin{abstract}
-% A large amount of free content is available over the Internet from
-% many different distributors. Most of this content uses the traditional
-% client-server model to handle requests from users. However, there is an
-% opportunity to use peer-to-peer techniques to reduce the cost of
-% much of this distribution, especially due to the altruistic nature
-% of many of these users. We present a general technique for
-% satisfying the needs of this P2P distribution, which is applicable
-% to many of these distributors systems. Our method makes use of a DHT
-% for storing the location of peers, using the cryptographic hash of
-% the file as a key. We then go further and implement a solution for
-% the distribution of Debian software packages.
-% 
-% We have designed a generally applicable peer-to-peer content
-% distribution model to be used by many of the free content
-% distributors operating today. It makes use of already existing
-% features of most package management systems, to create a
-% peer-to-peer distribution that should substantially reduce the costs
-% of hosting the content.
-% 
-% We have also implemented our design in software to be used in
-% conjuction with Debian-based distribution of Linux software
-% packages. It will soon be in use by many users of the Debian
-% project's distribution, and so serves as an example of the
-% opportunities that exist.
-% \end{abstract}
-
 \section{Introduction}
@@
 package distribution systems over the Internet to distribute content
-to their users. These distributors have developed many different
+to their users: most Linux distributions (e.g. Debian, Red Hat,
+Gentoo), Cygwin, CPAN, etc.
+These distributors have developed many different
 methods for this distribution, but they almost exclusively use a
@@
 popularity and number of users results in a large number of
-requests, which usually requires a network of mirrors to handle. Due
+requests, which usually requires a complicated network of mirrors to handle. Due
 to the free nature of this content, many users are willing and able
@@
 none is very well suited to this specific problem. Many distributors
-make their content available in some form using BitTorrent, but it
+make their content available in some form using BitTorrent \cite{COHEN03}, but it
 is not ideally suited to an application such as this one. BitTorrent
@@
 interested in a small subset of them. Other proposed implementations
-suffer from some of the same, and other, problems, and so none has
+suffer from some of the same, and other, problems and so none has
 been well-used to address this problem.
@@
 demands. It is based on the lessons learned from many previous
-implementations of successful peer-to-peer protocols, and makes use
-of a distributed hash table (DHT) and borrows many ideas from
-BitTorrent. The model relies on the pre-existence of cryptographic
+implementations of successful peer-to-peer protocols, such as
+Distributed Hash Tables and BitTorrent.
+The model relies on the pre-existence of cryptographic
 hashes of the content, which uniquely identify it for a request from
@@
 mostly well-known technologies, and so serves as an example for
-other free content distributors of the opportunity that can be met
+other free content distributors of the opportunity that can be easily met
 with such a system.
@@
 support the large number of requests. This content almost always
-supports some kind of file verification, using cryptographic hashes,
-to verify the completed downloads accuracy or authenticity.
+supports some kind of file verification using cryptographic hashes
+to verify the download's accuracy or authenticity.
 
@@
 using a package management tool that requests packages from
-websites; there are two software distribution systems for Mac OSX,
+websites; two software distribution systems exist for Mac OSX,
 fink and MacPorts, that also retrieve packages in this way. Also,
@@
 available before the download is attempted. We also expect that
-there are a number of users of these systems that are motivated by
+there are many users of these systems that are motivated by
 altruism to want to help out with this distribution. This is common
@@
 
-This situation presents a clear opportunity to use some form of
-peer-to-peer file-sharing protocol. The sparse interest in a large
+This situation presents a clear opportunity to use a new
+peer-to-peer file sharing protocol.
+A downloading peer can lookup the package hash using the protocol and, if it
+is found, download the file from those peers and verify the content.
+If the package hash can not be found, the peer will
+fallback to downloading from the original content location, and once
+complete will announce to other peers indicating that it now has
+the content. The original servers or mirrors thus act as \emph{seeds} for the
+P2P system without any modification to them.
+
+This functionality could be directly integrated into the
+package management software, although this would be
+difficult as the protocol should be running at all times, whereas the
+package software typically only runs until the download request is complete.
+Alternatively, since many of the package management software
+implementations use HTTP requests to download the files, it
+it possible to implement the P2P aspect as as HTTP caching
+proxy. The proxy will get uncached requests first from the P2P system,
+and then fallback to the normal HTTP request from a server. For
+methods that don't use HTTP requests, other types of proxies
+(possibly custom-made) may also be possible.
+
+The sparse interest in a large
 number of packages undergoing constant updating is well suited to
@@
 require unique keys to store and retrieve strings of data, which the
-cryptographic hashes used by these package management systems are
+cryptographic hashes used by the package management systems are
 perfect for. The stored and retrieved strings can then be pointers
-to the peers that have the content package that hashes to that key.
-A downloading peer can lookup the package hash in the DHT and, if it
-is found, download the file from those peers and verify the content.
-If the package hash can not be found in the DHT, the peer will
-fallback to downloading from the original content location, and once
-complete will add a new entry to the DHT indicating that it now has
-the content. The servers or mirrors thus act as \emph{seeds} for the
-P2P system without any modification to them.
+to the peers that have the content package that hashes to that key,
+as well as any other information needed by the protocol to facilitate
+the download.
 
-This desired P2P functionality could be integrated into the existing
-package management software in 2 ways. The functionality can be
-directly integrated into the software, though this could be
-difficult as the DHT should be running at all times, whereas the
-tools typically only run until the download request is complete.
-Alternatively, since many of the package management software
-implementations use HTTP requests to download the files, this makes
-it possible to implement the P2P aspect as a standard HTTP caching
-proxy, which will get uncached requests first from the P2P system,
-and then fallback to the normal HTTP request from a server. For
-methods that don't use HTTP requests, other types of proxies may
-also be possible.
-
-Downloading a large file efficiently from a number of peers is where
-BitTorrent shines as a peer-to-peer application. Its method of
-breaking up larger files into pieces, each with its own hash, makes
-it very easy to parallelize the downloading process and maximize the
-download speed. Since the package management system only makes
+To download larger files efficiently from a number of peers,
+the file needs to be broken up into pieces, each with its own hash.
+This method, inspired by BitTorrent, makes it very easy to
+parallelize the downloading process and maximize the download speed.
+Since the package management systems only make
 available a hash of the entire content, the piece hashes will have
-to be found through the P2P protocol. These hashes can be added to
-the DHT, or requested from other peers.
+to be found through the P2P protocol.
 
@@
 
-The DHT used is based on Khashmir, which is an implementation of the
-Kademlia DHT \cite{kademlia} used by most of the existing BitTorrent
+The DHT used is based on Khashmir, which is an implementation of
+Kademlia \cite{kademlia} used by most of the existing BitTorrent
 clients to implement trackerless operation. It has been modified to
-better support the retrieval of multiple values per key, and to
+better support the retrieval of multiple values (peers) per key, and to
 improve the response time so as to satisfy users' demands for speed.
@@
 strings) as described in the previous section. Downloading is
-accomplished by simple HTTP requests to the peers' identified to
-have the desired file. All peers support HTTP/1.1, both in the
-server and the client, which allows for pipelining of multiple
+accomplished by simple HTTP requests for the hash of the file,
+which is sent to the peers identified from the DHT lookup to
+have the desired file. All peers support HTTP/1.1, both as
+servers and clients, which allows for pipelining of multiple
 requests to a peer, and the requesting of smaller pieces of a large
@@
 We have chosen a piece size of 512 kB, which means that 17,515
-(79\%) of the almost 23 thousand packages will not require piece
+(79\%) of the almost 23,000 Debian packages will not require piece
 information as they are smaller than a single piece. There are 3054
@@
 Finally, there are only 62 packages (0.3\%) that require more than
-70 pieces, and so for them a separate request to a peer for the
-piece hashes is needed.
+70 pieces, and so need a separate request to a peer for the
+piece hashes.
 
-We have deployed the DHT from this implementation on PlanetLab to
+We have deployed the DHT part of this implementation on PlanetLab to
 test it's speed and effectiveness at delivering results. We find

### 2008-05-03 apt_p2p/apt_p2p_conf.py
Lower the timeout to 9 sec to reduce delays.

@@ -124,14 +124,14 @@ DHT_DEFAULTS = {
     'KEY_EXPIRE': '3h', # 3 hours
     
     # Timeout KRPC requests to nodes after this time.
-    'KRPC_TIMEOUT': '14s',
+    'KRPC_TIMEOUT': '9s',
     
     # KRPC requests are resent using exponential backoff starting with this delay.
     # The request will first be resent after the delay set here.
     # The request will be resent again after twice the delay set here. etc.
-    # e.g. if TIMEOUT is 14 sec., and INITIAL_DELAY is 2 sec., then requests will
+    # e.g. if TIMEOUT is 9 sec., and INITIAL_DELAY is 2 sec., then requests will
     # be resent at times 0, 2 (2 sec. later), and 6 (4 sec. later), and then will
-    # timeout at 14.
+    # timeout at 9.
     'KRPC_INITIAL_DELAY': '2s',
 
     # whether to spew info about the requests/responses in the protocol

### 2008-05-03 apt_p2p_Khashmir/DHT.py

@@ -337,7 +337,7 @@ class TestSimpleDHT(unittest.TestCase):
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
-                    'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
+                    'KRPC_TIMEOUT': 9, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def setUp(self):
@@ -458,7 +458,7 @@ class TestMultiDHT(unittest.TestCase):
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
-                    'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
+                    'KRPC_TIMEOUT': 9, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def setUp(self):

### 2008-05-03 apt_p2p_Khashmir/khashmir.py

@@ -563,7 +563,7 @@ class SimpleTests(unittest.TestCase):
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
-                    'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
+                    'KRPC_TIMEOUT': 9, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def setUp(self):
@@ -637,7 +637,7 @@ class MultiTest(unittest.TestCase):
                     'STORE_REDUNDANCY': 3, 'RETRIEVE_VALUES': -10000,
                     'MAX_FAILURES': 3, 'LOCAL_OK': True,
                     'MIN_PING_INTERVAL': 900,'BUCKET_STALENESS': 3600,
-                    'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2,
+                    'KRPC_TIMEOUT': 9, 'KRPC_INITIAL_DELAY': 2,
                     'KEY_EXPIRE': 3600, 'SPEW': True, }
 
     def _done(self, val):

### 2008-05-03 apt_p2p_Khashmir/krpc.py

@@ -253,7 +253,7 @@ class KrpcRequest(defer.Deferred):
         """Check for a unrecoverable timeout, otherwise resend."""
         self.later = None
         delay = datetime.now() - self.start
-        if delay > timedelta(seconds = self.config.get('KRPC_TIMEOUT', 14)):
+        if delay > timedelta(seconds = self.config.get('KRPC_TIMEOUT', 9)):
             log.msg('%r timed out after %0.2f sec' %
                     (self.tid, delay.seconds + delay.microseconds/1000000.0))
             self.protocol.timeOut(self.tid, self.method)
@@ -584,7 +584,7 @@ def make(port):
     from stats import StatsLogger
     af = Receiver()
     a = hostbroker(af, StatsLogger(None, None),
-                   {'KRPC_TIMEOUT': 14, 'KRPC_INITIAL_DELAY': 2, 'SPEW': False})
+                   {'KRPC_TIMEOUT': 9, 'KRPC_INITIAL_DELAY': 2, 'SPEW': False})
     a.protocol = KRPC
     p = reactor.listenUDP(port, a)
     return af, a, p

### 2008-05-03 test.py

@@ -522,14 +522,14 @@ LOCAL_OK = yes
 KEY_EXPIRE = 3h
 
 # Timeout KRPC requests to nodes after this time.
-KRPC_TIMEOUT = 14s
+KRPC_TIMEOUT = 9s
 
 # KRPC requests are resent using exponential backoff starting with this delay.
 # The request will first be resent after the delay set here.
 # The request will be resent again after twice the delay set here. etc.
-# e.g. if TIMEOUT is 14 sec., and INITIAL_DELAY is 2 sec., then requests will
+# e.g. if TIMEOUT is 9 sec., and INITIAL_DELAY is 2 sec., then requests will
 # be resent at times 0, 2 (2 sec. later), and 6 (4 sec. later), and then will
-# timeout at 14.
+# timeout at 9.
 KRPC_INITIAL_DELAY = 2s
 
 # whether to spew info about the requests/responses in the protocol

### 2008-05-03 apt_p2p/apt_p2p_conf.py
Update the version numbers and changelog for new release.

@@ -23,7 +23,7 @@ class ConfigError(Exception):
     def __str__(self):
         return repr(self.message)
 
-version = versions.Version('apt-p2p', 0, 1, 0)
+version = versions.Version('apt-p2p', 0, 1, 1)
 
 # Set the home parameter
 home = os.path.expandvars('${HOME}')

### 2008-05-03 setup.py

@@ -5,7 +5,7 @@ from distutils.core import setup
 
 setup(
     name = "apt-p2p",
-    version = "0.1.0",
+    version = "0.1.1",
     author = "Cameron Dale",
     author_email = "<camrdale@gmail.com>",
     url = "http://www.camrdale.org/apt-p2p.html",

### 2008-05-03 apt_p2p_Khashmir/krpc.py
Make sure full timeouts occur even if a resend timeout has not.

@@ -212,8 +212,10 @@ class KrpcRequest(defer.Deferred):
     @ivar delay: the last timeout delay sent
     @type start: C{datetime}
     @ivar start: the time to request was started at
-    @type later: L{twisted.internet.interfaces.IDelayedCall}
-    @ivar later: the pending call to timeout the last sent request
+    @type laterNextTimeout: L{twisted.internet.interfaces.IDelayedCall}
+    @ivar laterNextTimeout: the pending call to timeout the last sent request
+    @type laterFinalTimeout: L{twisted.internet.interfaces.IDelayedCall}
+    @ivar laterFinalTimeout: the pending call to timeout the entire request
     """
     
     def __init__(self, protocol, newTID, method, data, config):
@@ -237,32 +239,38 @@ class KrpcRequest(defer.Deferred):
         self.config = config
         self.delay = self.config.get('KRPC_INITIAL_DELAY', 2)
         self.start = datetime.now()
-        self.later = None
+        self.laterNextTimeout = None
+        self.laterFinalTimeout = reactor.callLater(self.config.get('KRPC_TIMEOUT', 9), self.finalTimeout)
         reactor.callLater(0, self.send)
         
     def send(self):
         """Send the request to the remote node."""
-        assert not self.later, 'There is already a pending request'
-        self.later = reactor.callLater(self.delay, self.timeOut)
+        assert not self.laterNextTimeout, 'There is already a pending request'
+        self.laterNextTimeout = reactor.callLater(self.delay, self.nextTimeout)
         try:
             self.protocol.sendData(self.method, self.data)
         except:
             log.err()
 
-    def timeOut(self):
+    def nextTimeout(self):
         """Check for a unrecoverable timeout, otherwise resend."""
-        self.later = None
-        delay = datetime.now() - self.start
-        if delay > timedelta(seconds = self.config.get('KRPC_TIMEOUT', 9)):
-            log.msg('%r timed out after %0.2f sec' %
-                    (self.tid, delay.seconds + delay.microseconds/1000000.0))
-            self.protocol.timeOut(self.tid, self.method)
+        self.laterNextTimeout = None
+        if datetime.now() - self.start > timedelta(seconds = self.config.get('KRPC_TIMEOUT', 9)):
+            self.finalTimeout()
         elif self.protocol.stopped:
             log.msg('Timeout but can not resend %r, protocol has been stopped' % self.tid)
         else:
             self.delay *= 2
[…]

### 2008-05-04 apt_p2p/HTTPDownloader.py
Fixed a bug in the HTTP downloader that caused errors with some mirrors that always close the connections.

@@ -358,8 +358,9 @@ class Peer(ClientFactory):
     def _processLastResponse(self):
         """Save the download time of the last request for speed calculations."""
         if self._lastResponse is not None:
-            now = datetime.now()
-            self._downloadSpeeds.append((now, now - self._lastResponse[0], self._lastResponse[1]))
+            if self._lastResponse[1] is not None:
+                now = datetime.now()
+                self._downloadSpeeds.append((now, now - self._lastResponse[0], self._lastResponse[1]))
             self._lastResponse = None
             
     def downloadSpeed(self):

### 2008-05-05 apt_p2p/apt_p2p_conf.py
Update versions and changelog for new release.

@@ -23,7 +23,7 @@ class ConfigError(Exception):
     def __str__(self):
         return repr(self.message)
 
-version = versions.Version('apt-p2p', 0, 1, 1)
+version = versions.Version('apt-p2p', 0, 1, 2)
 
 # Set the home parameter
 home = os.path.expandvars('${HOME}')

### 2008-05-05 setup.py

@@ -5,7 +5,7 @@ from distutils.core import setup
 
 setup(
     name = "apt-p2p",
-    version = "0.1.1",
+    version = "0.1.2",
     author = "Cameron Dale",
     author_email = "<camrdale@gmail.com>",
     url = "http://www.camrdale.org/apt-p2p.html",

### 2008-06-22 apt_p2p/apt_p2p_conf.py
New changelog entry and version update for relase.

@@ -23,7 +23,7 @@ class ConfigError(Exception):
     def __str__(self):
         return repr(self.message)
 
-version = versions.Version('apt-p2p', 0, 1, 4)
+version = versions.Version('apt-p2p', 0, 1, 5)
 
 mapbase64 = '0123456789ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz.-'
 versionID = 'A'

### 2008-06-22 setup.py

@@ -5,7 +5,7 @@ from distutils.core import setup
 
 setup(
     name = "apt-p2p",
-    version = "0.1.4",
+    version = "0.1.5",
     author = "Cameron Dale",
     author_email = "<camrdale@gmail.com>",
     url = "http://www.camrdale.org/apt-p2p.html",

### 2008-06-22 apt_p2p_Khashmir/actions.py
Make sure bad nodes don't get added to the routing table.

@@ -191,21 +191,22 @@ class ActionBase:
 
     def gotResponse(self, dict, node):
         """Receive a response from a remote node."""
-        if node.id != self.caller.node.id:
-            reactor.callLater(0, self.caller.insertNode, node)
-        if self.finished or self.answered.has_key(node.id):
+        if self.finished or self.answered.has_key(node.id) or self.failed.has_key(node.id):
             # a day late and a dollar short
             return
         try:
             # Process the response
             self.processResponse(dict)
-            self.answered[node.id] = 1
         except Exception, e:
             # Unexpected error with the response
             log.msg("action %s failed on %s/%s: %r" % (self.action, node.host, node.port, e))
             if node.id != self.caller.node.id:
                 self.caller.nodeFailed(node)
             self.failed[node.id] = 1
+        else:
+            self.answered[node.id] = 1
+            if node.id != self.caller.node.id:
+                reactor.callLater(0, self.caller.insertNode, node)
         if self.outstanding.has_key(node.id):
             self.outstanding_results -= self.outstanding[node.id]
             del self.outstanding[node.id]

### 2008-06-24 docs/abstract/abstract.tex
Updated the abstract to the ACM class.

@@
-\documentclass[conference]{IEEEtran}
+\documentclass{sig-alt-release2}
 
 \usepackage[noadjust]{cite}
-
-\usepackage[dvips]{graphicx}
-
 \usepackage{url}
-\usepackage[cmex10]{amsmath}
-\interdisplaylinepenalty=2500
-\usepackage{threeparttable}
-\usepackage{multirow}
+\hyphenation{dis-tri-bu-ted}
 
@@
 
+\conferenceinfo{SIGCOMM'08,} {August 17--22, 2008, Seattle, Washington, USA.} 
+\CopyrightYear{2008}
+\crdata{978-1-60558-175-0/08/08}
+
 \title{Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads}
-\author{\IEEEauthorblockN{Cameron Dale}
-\IEEEauthorblockA{School of Computing Science (student)\\
-Simon Fraser University\\
-Burnaby, British Columbia, Canada\\
-Email: camerond@cs.sfu.ca}
-\and
-\IEEEauthorblockN{Jiangchuan Liu}
-\IEEEauthorblockA{School of Computing Science (faculty)\\
-Simon Fraser University\\
-Burnaby, British Columbia, Canada\\
-Email: jcliu@cs.sfu.ca}}
+\numberofauthors{2}
+\author{
+\alignauthor Cameron Dale\\
+	\affaddr{School of Computing Science (student)}\\
+	\affaddr{Simon Fraser University}\\
+	\affaddr{Burnaby, British Columbia, Canada}\\
+	\email{Email: camerond@cs.sfu.ca}
+%\and
+\alignauthor Jiangchuan Liu\\
+	\affaddr{School of Computing Science (faculty)}\\
+	\affaddr{Simon Fraser University}\\
+	\affaddr{Burnaby, British Columbia, Canada}\\
+	\email{Email: jcliu@cs.sfu.ca}
+}
 
 \maketitle
+
+\begin{abstract}
+We introduce the opportunity many free content distributors have to
+reduce their bandwidth costs using a peer-to-peer download
+mechanism. Our model for this peer-to-peer distribution is simple to
+implement, backwards-compatible with the existing infrastructure,
+and can be deployed incrementally. It makes wide use of a customized
+distributed hash table (DHT) to facilitate the finding of peers and
+the efficient downloading of files. We also present the example
+solution we created for one of these distributors, the Debian
+project, which is already in use by some users.
+\end{abstract}
+
+% A category with the (minimum) three required fields
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
+%A category including the fourth, optional field follows...
+\category{C.2.4}{Computer-Communication Networks}{Distributed Systems}[Distributed applications]
+\terms{Design, Performance, Reliability}
+\keywords{BitTorrent, P2P}
 
@@
 We present a new peer-to-peer distribution model to meet these
-demands. It is based on the lessons learned from many previous
-implementations of successful peer-to-peer protocols, such as
+demands, based on other successful peer-to-peer protocols such as
 Distributed Hash Tables and BitTorrent.
@@
 many of the standard Unix/Linux tools in a Windows environment,
-using a package management tool that requests packages from
+using a package management tool that requests \texttt{tar} files from
 websites; two software distribution systems exist for Mac OSX,
@@
 some systems use direct web downloads, but with a hash verification
-file also available for download next to the desired file. These
+file also available. These
 hash files usually have the same file name, but with an added
@@
 the content. The original servers or mirrors thus act as \emph{seeds} for the
-P2P system without any modification to them.
+P2P system without any modification to them, and users are satisfied even when
+there are no peers which allows the system to be deployed incrementally.
 
@@
 package software typically only runs until the download request is complete.
-Alternatively, since many of the package management software
-implementations use HTTP requests to download the files, it
-it possible to implement the P2P aspect as as HTTP caching
+Alternatively, it
+it possible to implement the P2P aspect as a caching
 proxy. The proxy will get uncached requests first from the P2P system,
-and then fallback to the normal HTTP request from a server. For
-methods that don't use HTTP requests, other types of proxies
-(possibly custom-made) may also be possible.
+and then fallback to the normal request from a server.
 
@@
 Linux distributions. \texttt{apt} uses SHA1 hashes to verify most
-downloaded files, including the large index files that contain
+downloaded files, including the large index files that contains detailed
 information such as the hashes of the individual packages. Since all
@@
 accomplished by simple HTTP requests for the hash of the file,
-which is sent to the peers identified from the DHT lookup to
+which is sent to the peers identified in a DHT lookup to
 have the desired file. All peers support HTTP/1.1, both as
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{./IEEEabrv,./all}
+\bibliographystyle{abbrv}
+\bibliography{./all}
 

### 2008-06-24 docs/abstract/sig-alternate.tex
Updated the abstract to the ACM class.

% This is "sig-alternate.tex" V1.8 June 2007
% This file should be compiled with V2.3 of "sig-alternate.cls" June 2007
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.3 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.3) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@acm.org)
% ===============================================================
%
% For tracking purposes - this is V1.8 - June 2007

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Delphi theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (2.54cm [1"] top and
bottom and 1.9cm [.75"] left and right; specified column width
(8.45cm [3.33"]) and gutter size (.083cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
\begin{figure*}
\centering
\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
\psfig{file=rosette.ps, height=1in, width=1in,}
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2008-06-30 docs/abstract/abstract.tex
Final version of abstract.

@@
 \alignauthor Cameron Dale\\
-	\affaddr{School of Computing Science (student)}\\
+	\affaddr{School of Computing Science}\\
 	\affaddr{Simon Fraser University}\\
@@
 \alignauthor Jiangchuan Liu\\
-	\affaddr{School of Computing Science (faculty)}\\
+	\affaddr{School of Computing Science}\\
 	\affaddr{Simon Fraser University}\\
@@
 solution we created for one of these distributors, the Debian
-project, which is already in use by some users.
+project, which is already seeing some use.
 \end{abstract}
@@
 package distribution systems over the Internet to distribute content
-to their users: most Linux distributions (e.g. Debian, Red Hat,
+to their users: e.g. most Linux distributions (Debian, Red Hat,
 Gentoo), Cygwin, CPAN, etc.
@@
 client-server model to satisfy user requests. The contents' size,
-popularity and number of users results in a large number of
+popularity and the number of users results in a large number of
 requests, which usually requires a complicated network of mirrors to handle. Due
-to the free nature of this content, many users are willing and able
+to the free nature of this content, many users are willing
 to contribute upload bandwidth to this distribution, but currently
@@
 We also present a sample implementation based on the Debian package
-distribution system. This implementation is simple, makes use of
-mostly well-known technologies, and so serves as an example for
+distribution system. This implementation is simple and makes use of
+well-known technologies, and so serves as an example for
 other free content distributors of the opportunity that can be easily met
@@
 Most Linux distributions use a software package management system
-that fetches packages to be installed from a network of mirrors.
-Debian-based distributions uses the \texttt{apt} program, which
+that fetches packages from a network of mirrors.
+Debian-based distributions use the \texttt{apt} program, which
 downloads Debian packages from one of many mirrors. RPM-based
@@
 the content. The original servers or mirrors thus act as \emph{seeds} for the
-P2P system without any modification to them, and users are satisfied even when
-there are no peers which allows the system to be deployed incrementally.
+P2P system, without any modification to them. Users are satisfied even when
+there are no peers, allowing the system to be deployed incrementally.
 

### 2008-08-16 docs/paper/paper.tex
Added the new paper targetting INFOCOM to the docs.

\documentclass[conference]{IEEEtran}

%% INFOCOM addition:
\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil\leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother
\pagestyle{headings}

\usepackage[noadjust]{cite}

\usepackage[dvips]{graphicx}

\usepackage{url}

\begin{document}

\title{Peer-to-Peer Distribution of Free Software Packages}
\author{\IEEEauthorblockN{Cameron Dale}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: camerond@cs.sfu.ca}
\and
\IEEEauthorblockN{Jiangchuan Liu}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: jcliu@cs.sfu.ca}}

\maketitle

\begin{abstract}
A large amount of free software packages are available over the
Internet from many different distributors. Most of these
distributors use the traditional client-server model to handle
requests from users. However, there is an excellent opportunity to
use peer-to-peer techniques to reduce the cost of much of this
distribution, especially due to the altruistic nature of many of
these users. There are no existing solution suitable for this
situation, so we present a new technique for satisfying the needs of
this P2P distribution, which is generally applicable to many of
these distributors' systems. Our method makes use of a DHT for
storing the location of peers, using the cryptographic hash of the
package as a key. To show the simplicity and functionality, we
implement a solution for the distribution of Debian software
packages, including the many DHT customizations needed. Finally, we
analyze our system to determine how it is performing and what effect
it is having.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{intro}

There are a large number of free software package distributors using
package distribution systems over the Internet to distribute
software to their users. These distributors have developed many
different methods for this distribution, but they almost exclusively
use a client-server model to satisfy user requests. The popularity
and number of users results in a large number of requests, which
usually requires a network of mirrors to handle. Due to the free
nature of this software, many users are willing and able to
contribute upload bandwidth to this distribution, but have no
current way to do this.

We present a new peer-to-peer distribution model to meet these
demands. It is based on many previous implementations of successful
peer-to-peer protocols, especially distributed hash tables (DHT) and
BitTorrent. The model relies on the pre-existence of cryptographic
hashes of the packages, which should uniquely identify it for a
request from other peers. If the peer-to-peer download fails, then
the original request to the server is used as a fallback to prevent
any dissatisfaction from users. The peer can then share this new
package with others through the P2P system.

First, we examine the opportunity that is available for many of
these free software package distributors. We present an overview of
a system that would efficiently satisfy the demands of a large
number of users, and should significantly reduce the currently
substantial bandwidth requirements of hosting this software. We then
present an example implementation based on the Debian package
distribution system. This implementation will be used by a large
number of users, and serves as an example for other free software
distributors of the opportunity that can be met with such a system.

This paper is organized as follows. The current situation and its
problems are presented in section \ref{situation}. We analyze some
related solutions in section \ref{related}, and then present our
solution in section \ref{opportunity}. We then detail our sample
implementation for Debian-based distributions in section
\ref{implementation}, including an in-depth look at our DHT
customizations in section \ref{custom_dht}. We analyze our deployed
sample implementation in section \ref{analysis}, suggest some future
enhancements in section \ref{future}, and conclude the paper in
section \ref{conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem/Situation}
\label{situation}

There are a large number of groups using the Internet to distribute
their free software packages. These packages are usually available
from a free download site, which usually requires a network of
mirrors to support the large number of requests. These distribution
systems almost always support some kind of file verification,
usually cryptographic hashes, to verify the completed downloads
accuracy or authenticity.

\subsection{Examples}
\label{examples}

Most Linux distributions use a software package management system
that fetches packages to be installed from an archive of packages
hosted on a network of mirrors. The Debian project, and other
Debian-based distributions such as Ubuntu and Knoppix, use the
\texttt{apt} (Advanced Package Tool) program, which downloads Debian
packages in the \texttt{.deb} format from one of many HTTP mirrors.
The program will first download index files that contain a listing
of which packages are available, as well as important information
such as their size, location, and a hash of their content. The user
can then select which packages to install or upgrade, and
\texttt{apt} will download and verify them before installing them.

There are also several similar frontends for the RPM-based
distributions. Red Hat's Fedora project uses the \texttt{yum}
program, SUSE uses \texttt{YAST}, while Mandriva has
\texttt{Rpmdrake}, all of which are used to obtain RPMs from
mirrors. Other distributions use tarballs (\texttt{.tar.gz} or
\texttt{.tar.bz2}) to contain their packages. Gentoo's package
manager is called \texttt{portage}, SlackWare Linux uses
\texttt{pkgtools}, and FreeBSD has a suite of command-line tools,
all of which download these tarballs from web servers.

Other software distributors also use a similar system. CPAN
distributes software packages for the PERL
programming language, using SOAP RPC requests to find and download
files. Cygwin provides many of the
standard Unix/Linux tools in a Windows environment, using a
package management tool that requests packages from websites. There
are two software distribution systems for Mac OSX, fink and
MacPorts, that also retrieve packages in this way.

Also, some systems use direct web downloads, but with a hash
verification file also available for download next to the desired
file. These hash files usually have the same file name, but with an
added extension identifying the hash used (e.g. \texttt{.md5} for
the MD5 hash). This type of file downloading and verification is
typical of free software hosting facilities that are open to anyone
to use, such as SourceForge.

\subsection{Similarities}
\label{similarities}

The important things to note for each of the systems mentioned in
section \ref{examples}, is that they all have the following in
common:
\begin{itemize}
 \item The content is divided into distinct units (packages), each
       of which contains a small independent part of all the
       content.
 \item The packages are avaliable for anyone to freely download.
 \item Users are typically not interested in downloading all of the
       packages available.
 \item Hashes of the packages are available before the download is
       attempted.
 \item Requests to download the packages are sent by a tool, not
       directly by the user (though the tool is responding to
       requests from the user).
\end{itemize}

We also expect that there are a number of users of these systems
that are motivated by altruism to want to help out with this
distribution. This is common in these systems due to the free nature
of the software being delivered, which encourages some to want to
help out in some way. A number of the systems are also used by
groups that are staffed mostly, or sometimes completely, by
volunteers. This also encourages users to want to give back
to the volunteer community that has created the software they are
using.

\subsection{Differences}
\label{differences}

Although at first it might seem that a single all-reaching solution
is possible for this situation, there are some differences in each
system that require independent solutions. The systems all use
different tools for their distribution, so any implementation would
have to be specific to the tool it is to integrate with. In
particular, how to obtain requests from the tool or user, and how to
find a hash that corresponds to the file being requested, is very
specific to each system.

Also, there may be no benefit in creating a single large solution to
integrate all these problems. Although they sometimes distribute
nearly identical packages (e.g. the same software available in
multiple Linux distributions), it is not exactly identical and has
usually been tailored to the system by the distributor. The small
differences will change the hash of the files, and so will make it
impossible to distribute similar packages across systems. And,
although peer-to-peer systems can scale very well with the number of
peers in the system, there is some overhead involved, so having a
much larger system of peers would mean that requests could take
longer to complete.

\subsection{Problems}
\label{problems}

There are some attributes of software package distribution that make
it difficult to use an existing peer-to-peer solution with, or to
design a new solution for this need.

\subsubsection{Archive Dimensions}

The dimensions of the software archive makes it difficult to
distribute efficiently. While most of the packages are very small in
size, there are some that are quite large. There are too many
packages to distribute each individually, but the archive is also
too large to distribute in its entirety. In some archives there are
also divisions of the archive into sections, e.g. by the computer
architecture that the package is intended for.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{apt_p2p_simulation-size_CDF.eps}
\caption{The CDF of the size of packages in a Debian system, both
for the actual size and adjusted size based on the popularity of
the package.}
\label{size_CDF}
\end{figure}

For example, Figure~\ref{size_CDF} shows the size of packages in the
Debian distribution. While 80\% of the packages are less than
512~KB, some of the packages are hundreds of megabytes. The entire
archive consists of 22,298 packages and is approximately 119,000 MB
in size. Some packages are divided by architecture, but there are a
large number of packages that are not architecture specific and so
can be installed on any computer architecture.

\subsubsection{Package Updates}

The software packages being distributed are being constantly
updated. These updates could be the result of the software creators
releasing a new version of the software with improved functionality,
or the result of the distributor updating their packaging of the
software to meet new requirements. Even if the distributor
periodically makes \emph{stable} releases, which are snapshots of
all the packages in the archive at a certain time, updates are still
released for security issues or serious bugs.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{size-quarter.eps}
\caption{The amount of data in the Debian archive that is updated
each day, broken down by architecture.}
\label{update_size}
\end{figure}

For example, Figure~\ref{update_size} shows the amount of data in
the Debian archive that was updated each day over a period of 3
months. Each day approximately 1.5\% of the 119,000 MB archive is
updated with new versions of packages.

\subsubsection{Limited Interest}

Though there are a large number of packages, and a large number of
users, interest in a given version of a package could be very
limited. If the package is necessary for the system (e.g. the
package distribution software), then it is likely that everyone will
have it installed, but there are very few of these packages. Most
packages fall in the category of optional or extra, and so are
interesting to only a limited number of people.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{apt_p2p_popularity-cdf.eps}
\caption{The CDF of the popularity of packages in a Debian system.}
\label{popularity_CDF}
\end{figure}

For example, the Debian distribution tracks the popularity of its
packages using popcon \cite{popcon}. Figure~\ref{popularity_CDF}
shows the cumulative distribution function of the percentage of all
users who install each package. Though some packages are installed
by everyone, 80\% of the packages are installed by less than 1\% of
users.

\subsubsection{Interactive Users}

The package management software that downloads packages displays
some kind of indication of speed and completeness for users to
watch. Since previous client-server downloads occur in a sequential
fashion, the package management software also measures the speed
based on sequential downloading. This requires the peer-to-peer
solution to be reasonably responsive at retrieving packages
sequentially.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work/Existing Solutions}
\label{related}

There are many peer-to-peer implementations available today, but
none is very well suited to this specific problem.

\subsection{BitTorrent}
\label{bittorrent}

Many distributors make their software available in some form using
BitTorrent \cite{COHEN03}, e.g. for the distribution of CD
images. This is not an ideal situation though, as it requires the
peers to download large numbers of packages that they are not
interested in, and prevents them from updating to new packages
without downloading another image containing a lot of the same
packages they already have. Unfortunately, BitTorrent is not ideally
suited to an application such as this one.

First, there is no obvious way to divide the packages into torrents.
Most of the packages are too small, and there are too many packages
in the entire archive, to create individual torrents for each one.
However, all the packages together are too large to track
efficiently as a single torrent. Some division of the archive's
packages into torrents is obviously necessary, but wherever that
split occurs it will cause either some duplication of connections,
or prevent some peers from connecting to others who do have the same
content. Also, a small number of the packages can be updated every
day which would add new files to the torrent, thereby changing its
\emph{infohash} identifier and making it a new torrent. This will
severely fracture the download population, even though peers in the
new torrent share 99\% of the packages in common with peers in the
old torrent.

Other issues also prevent BitTorrent from being a good solution to
this problem. BitTorrent's fixed piece sizes that disregard file
boundaries are bigger than many of the packages in the archive. This
will waste peers' downloading bandwidth as they will end up
downloading parts of other packages just to get the piece that
contains the package they do want. Incentives to share (upload) are
no longer needed, as the software is freely available for anyone to
download without sharing. Seeds are also not needed as the mirrors
can serve in that capacity. Finally, BitTorrent downloads files
randomly, which does not work well with the interactive package
management tools expectation of sequential downloads.

\subsection{Other Solutions}
\label{others}

There have also been other attempted implementations, usually based
on some form of BitTorrent, to address this problem. Some other
implementations have used BitTorrent almost unmodified, while others
have only looked at using DHTs to replace the tracker in a
BitTorrent system. apt-torrent \cite{apttorrent} creates torrents
for some of the larger packages available, but it ignores the
smaller packages, which are often the most popular. DebTorrent
\cite{debtorrent} makes widespread modifications to a traditional
BitTorrent client, to try and fix the drawbacks mentioned in section
\ref{bittorrent}. However, these changes also require some
modifications to the distribution system to support it.

Others have also used DHTs to support this type of functionality.
Kenosis \cite{kenosis} is a P2P Remote Procedure Call
client also based on the Kademlia DHT, but it is meant as a P2P
primitive system on which other tools can be built, and so it has no
file sharing functionality at all. Many have used a DHT as a drop-in
replacement for the tracker in a BitTorrent system
\cite{bittorrent-dht, azureus-dht}, but such systems only use the
DHT to find peers for the same torrent, so the file sharing uses
traditional BitTorrent and so is not ideal for the reasons listed
above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Opportunity/Our Proposal}
\label{opportunity}

The situation described in section \ref{situation} presents a clear
opportunity to use some form of peer-to-peer file-sharing protocol
to allow willing users to contribute upload bandwidth. This sparse
interest in a large number of packages undergoing constant updating
is well suited to the functionality provided by a Distributed Hash
Table (DHT). DHTs require unique keys to store and retrieve strings
of data, for which the cryptographic hashes used by these package
management systems are perfect for. The stored and retrieved strings
can then be pointers to the peers that have the package that hashes
to that key. A downloading peer can lookup the package hash in the
DHT and, if it is found, download the file from those peers and
verify the package with the hash. Once the download is complete, the
peer will add its entry to the DHT indicating that it now has the
package.

The fact that this package is also available to download for free
from a server is very important to our proposal. If the package hash
can not be found in the DHT, the peer can then fallback to
downloading from the original location (i.e. the network of
mirrors). The mirrors thus, with no modification to their
functionality, serve as seeds for the packages in the peer-to-peer
system. Any packages that have just been updated, or that are very
rare, and so don't have any peers available can always be found on
the mirror. Once the peer has completed the download from the mirror
and verified the package, it can then add itself to the DHT as the
first peer for the new package, so that future requests from peers
will not need the mirror.

The trust of the package is also always guaranteed through the use
of the cryptographic hashes. Nothing can be downloaded from a peer
until the hash is looked up in the DHT, so a hash must first come
from a trusted source (i.e. a mirror). Most distributors use index
files that contain hashes for a large number of the packages in
their archive, and which are also hashed. After retrieving the
index's hash from the mirror, the index file can be downloaded from
peers and verified. Then the program has access to all the hashes of
the packages it will be downloading, all of which can be verified
with a \emph{chain of trust} that stretches back to the original
distributor's server.

\subsection{Implementation Options}
\label{imp_options}

There are several ways to implement the desired P2P functionality
into the existing package management software. The functionality can
be directly integrated through modifications to the software, though
this could be difficult as the P2P functionality should be running
at all times. This is needed both for efficient lookups and to
support uploading of already downloaded packages. Unfortunately, the
package management tools typically only run until the download and
install request is complete.

Many of the package management software implementations use HTTP
requests from web servers to download the packages, which makes it
possible to implement the P2P aspect as an almost standard HTTP
caching proxy. This proxy will run as a daemon in the background,
listening for requests from the package management tool for package
files. It will get uncached requests first from the P2P system, or
falling back to the normal HTTP request from a server should it not
be found. For methods that don't use HTTP requests, other types of
proxies may also be possible.

\subsection{Downloading From Peers}
\label{downloading}

Although not necessary, we recommend implementing a download
protocol that is similar to the protocol used to fetch packages from
the distributor's servers. This simplifies the P2P program, as it
can then treat peers and mirrors almost identically when requesting
packages. In fact, the mirrors can be used when there are only a few
slow peers available for a file to help speed up the download
process.

Downloading a file efficiently from a number of peers is where
BitTorrent shines as a peer-to-peer application. Its method of
breaking up larger files into pieces, each with its own hash,
makes it very easy to parallelize the downloading process and
maximize the download speed. For very small packages (i.e. less than
the piece size), this parallel downloading is not necessary, or
even desirable. However, this method should still be used, in
conjunction with the DHT, for the larger packages that are
available.

Since the package management system only stores a hash of the entire
package, and not of pieces of that package, we will need to be able
to store and retrieve these piece hashes using the P2P protocol. In
addition to storing the file download location in the DHT (which
would still be used for small files), a peer will store a
\emph{torrent string} containing the peer's hashes of the pieces of
the larger files. These piece hashes could be compared ahead of time
to determine which peers have the same piece hashes (they all
should), and then used during the download to verify the pieces of
the downloaded package.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Sample Implementation/Implementation}
\label{implementation}

We have created a sample implementation that functions as described
in section \ref{opportunity}, and is freely available for other
distributors to download and modify \cite{apt-p2p}. This software,
called \texttt{apt-p2p}, interacts with the \texttt{apt} tool, which
is found in most Debian-based Linux distributions. \texttt{apt} uses
SHA1 hashes to verify most downloaded files, including the large
index files that contain the hashes of the individual packages. We
chose this distribution system as it is familiar to us, it allows
software contributions, and there are interesting statistics
available for analyzing the popularity of the software packages
\cite{popcon}.

Since all requests from apt are in the form of HTTP downloads from a
server, the implementation takes the form of a caching HTTP proxy.
Making a standard \texttt{apt} implementation use the proxy is then
as simple as prepending the proxy location and port to the front of
the mirror name in \texttt{apt}'s configuration file (i.e.
``localhost:9977/mirrorname.debian.org/\ldots'').

We created a customized DHT based on Khashmir \cite{khashmir}, which
is an implementation of Kademlia \cite{kademlia} using methods
familiar to BitTorrent developers. Khashmir is also the same DHT
implementation used by most of the existing BitTorrent clients to
implement trackerless operation. The communication is all handled by
UDP messages, and RPC (remote procedure call) requests and responses
are all \emph{bencoded} in the same way as BitTorrent's
\texttt{.torrent} files. Khashmir uses the high-level Twisted
event-driven networking engine \cite{twisted}, so we also use
Twisted in our sample implementation for all other networking needs.
More details of this customized DHT can be found below in section
\ref{custom_dht}.

Downloading is accomplished by sending simple HTTP requests to the
peers identified by lookups in the DHT to have the desired file.
Requests for a package are made using the package's hash, properly
encoded, as the URL to request from the peer. The HTTP server used
for the proxy also doubles as the server listening for requests for
downloads from other peers. All peers support HTTP/1.1, both in the
server and the client, which allows for pipelining of multiple
requests to a peer, and the requesting of smaller pieces of a large
file using the Range request header.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Customized DHT}
\label{custom_dht}

A large contribution of our work is in the customization and use of
a Distributed Hash Table (DHT). Although our DHT is based on
Kademlia, we have made many improvements to it to make it suitable
for this application. In addition to a novel storage technique to
support piece hashes, we have improved the response time of looking
up queries, allowed the storage of multiple values for each key, and
incorporated some improvements from BitTorrent's tracker-less DHT
implementation.

\subsection{Kademlia Background}
\label{kademlia}

The Kademlia DHT, like most other DHTs, assigns IDs to peers from
the same space that is used for keys. The peers with IDs closest to
the desired key will then store the values for that key. Lookups are
recursive, as nodes are queried in each step that are exponentially
closer to the key than in the previous step.

Nodes in a Kademlia system support four primitive requests.
\texttt{ping} will cause a peer to return nothing, and is only used
to determine if a node is still alive, while \texttt{store} tells a
node to store a value associated with a given key. The most
important primitives are \texttt{find\_node} and
\texttt{find\_value}, which both function recursively to find nodes
close to a key. The queried nodes will return a list of the nodes
they know about that are closest to the key, allowing the querying
node to quickly traverse the DHT to find the nodes close to the
desired key. The only difference between them is that the
\texttt{find\_value} query will cause a node to return a value, if
it has one for that key, instead of a list of nodes.

\subsection{Piece Hash Storage}
\label{pieces}

Hashes of pieces of the larger package files are needed to support
their efficient downloading from multiple peers.
For large files (5 or more pieces), the torrent strings described in
section \ref{downloading}
are too long to store with the peer's download info in the DHT. This
is due to the limitation that a single UDP packet should be less
than 1472 bytes to avoid fragmentation.
% For example, for a file of 4 pieces, the peer would need to store a
% value 120 bytes in size (IP address, port, four 20-byte pieces and
% some overhead), which would limit a return value from including more
% than 10 values to a request.

Instead, the peers will store the torrent string for large files
separately in the DHT, and only contain a reference to it in their
stored value for the hash of the file. The reference is an SHA1 hash
of the entire concatenated length of the torrent string. If the
torrent string is short enough to store in the DHT (i.e. less than
1472 bytes, or about 70 pieces for the SHA1 hash), then a lookup of
that hash in the DHT will return the torrent string. Otherwise, a
request to the peer for the hash (using the same method as file
downloads, i.e. HTTP), will cause the peer to return the torrent
string.

Figure \ref{size_CDF} shows the package size of the 22,298 packages
available in Debian in January 2008. We can see that most of the
packages are quite small, and so most will therefore not require
piece hash information to download. We have chosen a piece
size of 512 kB, which means that 17,515 (78\%) of the packages will
not require this information. There are 3054 packages that will
require 2 to 4 pieces, for which the torrent string can be stored
directly with the package hash in the DHT. There are 1667 packages
that will require a separate lookup in the DHT for the longer
torrent string as they require 5 to 70 pieces. Finally, there are
only 62 packages that require more than 70 pieces, and so will
require a separate request to a peer for the torrent string.

\subsection{Response Time}
\label{response_time}

Most of our customizations to the DHT have been to try and improve
the time of the recursive \texttt{find\_value} requests, as this can
cause long delays for the user waiting for a package download to
begin. The one problem that slows down such requests is waiting for
timeouts to occur before marking the node as failed and moving on.

Our first improvement is to retransmit a request multiple times
before a timeout occurs, in case the original request or its
response was lost by the unreliable UDP protocol. If it does not
receive a response, the requesting node will retransmit the request
after a short delay. This delay will increase exponentially for
later retransmissions, should the request again fail. Our current
implementation will retransmit the request after 2 seconds and 6
seconds (4 seconds after the first retransmission), and then timeout
after 9 seconds.

We have also added some improvements to the recursive
\texttt{find\_node} and \texttt{find\_value} queries to speed up the
process when nodes fail. If enough nodes have responded to the
current query such that there are many new nodes to query that are
closer to the desired key, then a stalled request to a node further
away will be dropped in favor of a new request to a closer node.
This has the effect of leap-frogging unresponsive nodes and
focussing attention on the nodes that do respond. We will also
prematurely abort a query while there are still oustanding requests,
if enough of the closest nodes have responded and there are no
closer nodes found. This prevents a far away unresponsive node from
making the query's completion wait for it to timeout.

Finally, we made all attempts possible to prevent firewalled and
NATted nodes from being added to the routing table for future
requests. Only a node that has responded to a request from us will
be added to the table. If a node has only sent us a request, we
attempt to send a \texttt{ping} to the node to determine if it is
NATted or not. Unfortunately, due to the delays used by NATs in
allowing UDP packets for a short time if one was recently sent by
the NATted host, the ping is likely to succeed even if the node is
NATted. We therefore also schedule a future ping to the node to make
sure it is still reachable after the NATs delay has hopefully
elapsed. We also schedule future pings of nodes that fail once to
respond to a request, as it takes multiple failures (currently 3)
before a node is removed from the routing table.

\subsection{Multiple Values}
\label{multiple_values}

The original design of Kademlia specified that each keywould have
only a single value associated with it. The RPC to find this value
was called \texttt{find\_value} and worked similarly to
\texttt{find\_node}, iteratively finding nodes with ID's closer to
the desired key. However, if a node had a value stored associated
with the searched for key, it would respond to the request with that
value instead of the list of nodes it knows about that are closer.

While this works well for single values, it can cause a problem when
there are multiple values. If the responding node is no longer one
of the closest to the key being searched for, then the values it is
returning will probably be the staler ones in the system, and it
will not have the latest stored values. However, the search for
closer nodes will stop here, as the queried node only returned the
values and not a list of nodes to recursively query. We could have
the request return both the values and the list of nodes, but that
would severely limit the size and number of the values that could be
returned.

Instead, we have broken up the original \texttt{find\_value}
operation into two parts. The new \texttt{find\_value} request
always returns a list of nodes that the node believes are closest to
the key, as well as a number indicating the number of values that
this node has for the key. Once a querying node has finished its
search for nodes and found the closest ones to the key, it can issue
\texttt{get\_value} requests to some nodes to actually retrieve the
values they have. This allows for much more control of when and how
many nodes to query for values. For example, a querying node could
abort the search once it has found enough values in some nodes, or
it could choose to only request values from the nodes that are
closest to the key being searched for.

\subsection{BitTorrent's Improvements}
\label{bittorrent_dht}

In the many years that some BitTorrent clients have been using a
Kademlia based DHT for tracker-less operation, the developers have
made many enhancements which we can take advantage of. One of the
most important is a security feature added to stop malicious nodes
from subscribing other nodes as downloaders. When a node issues a
request to another node to store its download info for that key, it
must include a \emph{token} returned by the storing node in a recent
\texttt{find\_node} request. The token is created by hashing the IP
address of the requesting peer with a temporary secret that expires
after several minutes. This prevents the requesting peer from faking
its IP address in the store request, since it must first receive a
response from a \texttt{find\_node} on that IP.

We also made some BitTorrent-inspired changes to the parameters of
the DHT originally specified by the authors of Kademlia. Since, as
we will show later, peers stay online in our system for much longer
periods of time, we reduced Kademlia's \emph{k} value from 20 to 8.
The value is supposed to be large enough such that any given
\emph{k} nodes are unlikely to fail within an hour of each other,
which is very unlikely in our system given the long uptimes of
nodes. We also increased the number of concurrent outstanding
requests allowed from 3 to 6 to speed up the recursive key finding
processes.

\subsection{Other Changes}
\label{other_changes}

We added one other new RPC request that nodes can make:
\texttt{join}. This request is only sent on first loading the DHT,
and is usually only sent to the bootstrap nodes that are listed for
the DHT. These bootstrap nodes will respond to the request with the
requesting peer's IP and port, so that the peer can determine what
its oustide IP address is and whether port translation is being
used. In the future, we hope to add functionality similar to STUN
\cite{STUN}, so that nodes can detect whether they are NATted and
take appropriate steps to circumvent it.

In addition, we have allowed peers to store values in the DHT, even
if the hash they are using is not the correct length. Most of the
keys used in the DHT are based on the SHA1 hash, and so they are 20
bytes in length. However, some files in the targetted Debian package
system are only hashed using the MD5 algorithm, and so the hashes
retrieved from the server are 16 bytes in length. The DHT will still
store values using these keys by padding the end of them with zeroes
to the proper length, as the loss of uniqueness from the 4 less
bytes is not an issue. Also, the DHT will happily store hashes that
are longer than 20 bytes, such as those from the 32 byte SHA256
algorithm, by truncating them to the correct length. Since the
requesting peer will have the full length of the hash, this will not
affect its attempt to verify the downloaded file.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Analysis}
\label{analysis}

Our \texttt{apt-p2p} implementation supporting the Debian package
distribution system has been available to all Debian users since May
3rd, 2008 \cite{apt-p2p-debian}, and will also be available in the
next release of Ubuntu \cite{apt-p2p-ubuntu}. We have since created
a \emph{walker} that will navigate the DHT and find all the peers
currently connected to it. This allows us to analyze many aspects of
our implementation.

\subsection{Peer Lifetimes}
\label{peer_life}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PWalker-peers.eps}
\caption{The number of peers found in the system, and how many are
behind a firewall or NAT.}
\label{walker_peers}
\end{figure}

We first began analyzing the DHT on June 24th, and continue to this
day, giving us 2 months of data so far. Figure~\ref{walker_peers}
shows the number of peers we have seen in the DHT during this time.
The peer population is very steady, with just over 50 regular users
participating in the DHT at any time. We also note that we find 100
users who connect regularly (weekly), and we have found 186 unique
users in the 2 months of our analysis. We determined which users are
behind a firewall or NAT, which is one of the main problems of
implementing a peer-to-peer network. These peers will be
unresponsive to DHT requests from peers they have not contacted
recently, which will cause the peer to wait for a timeout to occur
(currently set at 9 seconds) before moving on. They will also be
unable to contribute any upload bandwidth to other peers, as all
requests for packages from them will also timeout. From
Figure~\ref{walker_peers}, we see that approximately half of all
peers suffered from this restriction.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDuration-peers.eps}
\caption{The CDF of how long an average session will last.}
\label{duration_peers}
\end{figure}

Figure~\ref{duration_peers} shows the cumulative distribution of how
long a connection from a peer can be expected to last. Due to our
software being installed as a daemon that is started by default
every time their computer boots up, peers are expected to stay for a
long period in the system. 50\% of connections last longer than 5
hours, and 20\% last longer than 10 hours. These connections are
much longer than those reported by Saroiu et. al. \cite{saroiu2001}
for other P2P systems, which had 50\% of Napster and Gnutella
sessions lasting only 1 hour.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDuration-ind_peers.eps}
\caption{The CDF of the average time individual peers stay in the
system.}
\label{duration_ind_peers}
\end{figure}

We also examined the average time each individual peer spends in the
system. Figure~\ref{duration_peers} shows the cumulative
distribution of how long each individual peer remains in the system.
Here we see that 50\% of peers have average stays in the system
longer than 10 hours.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDuration-online_1.eps}
\caption{The fraction of peers that, given their current duration in
the system, will stay online for another hour.}
\label{duration_online_1}
\end{figure}

Since our DHT is based on Kademlia, which was designed based on the
probability that a node will remain up another hour, we also
analyzed our system for this parameter.
Figure~\ref{duration_online_1} shows the fraction of peers will
remain online for another hour, as a function of how long they have
been online so far. Maymounkov and Mazieres found that the longer a
node has been online, the higher the probability that it will stay
online \cite{kademlia}. Our results also show this behavior. In
addition, similar to the Gnutella peers, over 90\% of our peers that
have been online for 10 hours, will remain online for another hour.
Our results also show that, for our system, over 80\% of all peers
will remain online another hour, compared with around 50\% for
Gnutella.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDuration-online_6.eps}
\caption{The fraction of peers that, given their current duration in
the system, will stay online for another 6 hours.}
\label{duration_online_6}
\end{figure}

Since our peers are much longer-lived than other P2P systems, we
also looked at the fraction of peers that stay online for another 6
hours. Figure~\ref{duration_online_6} shows that over 60\% of peers
that are online for 10 hours will stay online for another 6.
However, we see an interesting decrease in this fraction between 8
and 12 hours, which can also be seen in
Figure~\ref{duration_online_1}. We believe this to be due to desktop
users, who regularly turn off their computers at night.

\subsection{Peer Statistics}
\label{peer_stats}

On July 31st we enhanced our walker to retrieve additional
information from each contacted peer. The peers are configured, by
default, to publish some statistics on how much they are downloading
and uploading, and their measured response times for DHT queries.
Our walker can extract this information if the peer is not
firewalled or NATted, it has not disabled this functionality, and if
it uses the same port for both its DHT (UDP) requests and download
(TCP) requests (which is also a configuration parameter).

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDownloaded-peers.eps}
\caption{The number of peers that were contacted to determine their
bandwidth, and the total number of peers in the system.}
\label{down_peers}
\end{figure}

Figure~\ref{down_peers} shows the total number of peers we have been
able to contact since starting to gather this additional
information, as well as how many total peers were found. We were
only able to contact 30\% of all the peers that connected to the
system during this time.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDownloaded-bw.eps}
\caption{The bandwidth of data that the contacted peers have
downloaded and uploaded.}
\label{down_bw}
\end{figure}

Figure~\ref{down_bw} shows the amount of data the peers we were able
to contact have downloaded. Peers measure their downloads from other
peers and mirrors separately, so we are able to get an idea of how
much savings our system is generating for the mirrors. We see that
the peers are downloading approximately 20\% of their package data
from other peers, which is saving the mirrors from supplying that
bandwidth. The actual numbers are only a lower bound, since we have
only contacted 30\% of the peers in the system, but we can estimate
that \texttt{apt-p2p} has already saved the mirrors 15 GB of
bandwidth, or 1 GB per day.

We also collected the statistics on the measured response time peers
were experiencing when sending requests to the DHT. We found that
the recursive \texttt{find\_value} query, which is necessary before
a download can occur, is taking 17 seconds on average. This
indicates that, on average, requests are experiencing almost 2
stalls while waiting for the 9 second timeouts to occur on
unresponsive peers. This is longer than our target of 10 seconds,
although it will only lead to a slight average delay in downloading
of 1.7 seconds when the default 10 concurrent downloads are
occurring.This increased response time is due to the number of peers
that were behind firewalls or NATs, which was much higher than we
anticipated. We do have plans to improve this through better
informing of users of their NATted status, the use of STUN
\cite{STUN} to circumvent the NATs, and by better exclusion of
NATted peers from the DHT (which will not prevent them from using
the system).

We were also concerned that the constant DHT requests and responses,
even while not downloading, would overwhelm some peers' network
connections. However, we found that peers are using 200 to 300 bytes
per second of bandwidth in servicing the DHT. These numbers are
small enough to not affect any other network services the peer would
be running.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Future Work}
\label{future}

We feel that our P2P software package distribution model is
well-suited to be used by many free software distributors. We hope
to convince them to adopt such a model in their distribution, or we
may port our existing system to some of the other groups for them to
try.

One aspect missing from our model is the removal of old packages
from the cache. Since our implementation is still relatively young,
we have not had to deal with the problems of a growing cache of
obsolete packages consuming all of a user's hard drive. We plan to
implement some form of least recently used (LRU) cache removal
technique, in which packages that are no longer available on the
server, no longer requested by peers, or simply are the oldest in
the cache, will be removed.

The most significant area of improvement still needed in our sample
implementation is to further speed up some of the slower recursive
DHT requests. We hope to accomplish this by further tuning the
parameters of our current system, better exclusion of NATted peers
from the routing tables, and through the use of STUN \cite{STUN} to
circumvent the NATs of the 50\% of the peers that have not
configured port forwarding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{conclusions}

We have designed a generally applicable peer-to-peer content
distribution model to be used by many of the free software
distributors operating today. It makes use of already existing
features of most package management systems to create a
peer-to-peer distribution that should substantially reduce the costs
of hosting the software packages.

We have also implemented our design in freely available software to
be used in conjuction with Debian-based distribution of Linux
software packages. It is currently in use by some users of the
Debian project's distribution, and so serves as an example of the
possibilities that exist.

\bibliographystyle{IEEEtran}
\bibliography{./IEEEabrv,./all}

\end{document}

### 2008-08-26 docs/paper/paper.tex
Added JC's changes to the paper.

@@
 
-\title{Peer-to-Peer Distribution of Free Software Packages}
+\title{\texttt{apt-p2p}: A Peer-to-Peer Distributor for Free Software Package Release and Update}
 \author{\IEEEauthorblockN{Cameron Dale}
@@
 
-This paper is organized as follows. The current situation and its
-problems are presented in section \ref{situation}. We analyze some
-related solutions in section \ref{related}, and then present our
-solution in section \ref{opportunity}. We then detail our sample
-implementation for Debian-based distributions in section
+The rest of this paper is organized as follows. The background and motivation are presented in Section \ref{situation}, together with related works in  Section \ref{related}. We propose
+our solution in Section \ref{opportunity}. We then detail our sample
+implementation for Debian-based distributions in Section
 \ref{implementation}, including an in-depth look at our DHT
-customizations in section \ref{custom_dht}. We analyze our deployed
-sample implementation in section \ref{analysis}, suggest some future
-enhancements in section \ref{future}, and conclude the paper in
-section \ref{conclusions}.
+customizations in Section \ref{custom_dht}. Its performance is evaluated in Section \ref{analysis}. Finally, 
+Section \ref{conclusions} concludes the paper and offers some future directions.
 
@@
 
-\section{Problem/Situation}
+\section{Background and Motivations}
 \label{situation}
 
-There are a large number of groups using the Internet to distribute
-their free software packages. These packages are usually available
-from a free download site, which usually requires a network of
-mirrors to support the large number of requests. These distribution
-systems almost always support some kind of file verification,
-usually cryptographic hashes, to verify the completed downloads
-accuracy or authenticity.
+In the free software society, there are a large number of groups using the Internet to 
+collaboratively develop and release their software. The ever increasing power of
+modern programming languages and operating systems has made these software, like commercial software, extremely large and complex, which often
+consists of many small units (packages). Together with their popularity among users, 
+an efficient and reliable management and distribution of these packages over the Internet has become an daunting task. In this section, we offer concrete examples illustrating the 
+unique challenges in this context.
 
-\subsection{Examples}
+\subsection{Free Software Package Distributor: Examples}
 \label{examples}
@@
 
-Other software distributors also use a similar system. CPAN
+Similar tools have been used for other software packages. CPAN
 distributes software packages for the PERL
@@
 
-Also, some systems use direct web downloads, but with a hash
-verification file also available for download next to the desired
-file. These hash files usually have the same file name, but with an
+Direct web downloading is also a common way, often coupled with a hash
+verification file to be downloaded next to the desired
+file. The hash file usually have the same file name, but with an
 added extension identifying the hash used (e.g. \texttt{.md5} for
@@
 
-\subsection{Similarities}
-\label{similarities}
+Given the free nature of these software, there are often a number of users 
+motivated by altruism to want to help out with their distribution. 
+This is particularly true considering that many of these software are used by
+groups that are staffed mostly, or sometimes completely, by
+volunteers. They are thus motivated to contribute their network resources, so as to promote the healthy development 
+of the volunteer community that released the software. As a matter of fact,
+we have see many free mirror sites hosting these software packages for downloading. 
+We also naturally expect that peer-to-peer distribution can be implementation in 
+this context, which scale well with large user bases and can easily explore the network resources made available by
+the volunteers.
 
-The important things to note for each of the systems mentioned in
-section \ref{examples}, is that they all have the following in
-common:
-\begin{itemize}
- \item The content is divided into distinct units (packages), each
-       of which contains a small independent part of all the
-       content.
- \item The packages are avaliable for anyone to freely download.
- \item Users are typically not interested in downloading all of the
-       packages available.
- \item Hashes of the packages are available before the download is
-       attempted.
- \item Requests to download the packages are sent by a tool, not
-       directly by the user (though the tool is responding to
-       requests from the user).
-\end{itemize}
-
-We also expect that there are a number of users of these systems
-that are motivated by altruism to want to help out with this
-distribution. This is common in these systems due to the free nature
-of the software being delivered, which encourages some to want to
-help out in some way. A number of the systems are also used by
-groups that are staffed mostly, or sometimes completely, by
-volunteers. This also encourages users to want to give back
-to the volunteer community that has created the software they are
-using.
-
-\subsection{Differences}
-\label{differences}
-
-Although at first it might seem that a single all-reaching solution
-is possible for this situation, there are some differences in each
-system that require independent solutions. The systems all use
-different tools for their distribution, so any implementation would
-have to be specific to the tool it is to integrate with. In
-particular, how to obtain requests from the tool or user, and how to
-find a hash that corresponds to the file being requested, is very
-specific to each system.
-
-Also, there may be no benefit in creating a single large solution to
-integrate all these problems. Although they sometimes distribute
-nearly identical packages (e.g. the same software available in
-multiple Linux distributions), it is not exactly identical and has
-usually been tailored to the system by the distributor. The small
-differences will change the hash of the files, and so will make it
-impossible to distribute similar packages across systems. And,
-although peer-to-peer systems can scale very well with the number of
-peers in the system, there is some overhead involved, so having a
-much larger system of peers would mean that requests could take
-longer to complete.
-
-\subsection{Problems}
+\subsection{Unique Characteristics}
 \label{problems}
 
-There are some attributes of software package distribution that make
-it difficult to use an existing peer-to-peer solution with, or to
-design a new solution for this need.
+While it seems straightforward to use an existing peer-to-peer file sharing tool like BitTorrent for
+free software package distribution, there are indeed a series of new challenges in this unique scenario:
 
@@
 
-The dimensions of the software archive makes it difficult to
-distribute efficiently. While most of the packages are very small in
+While most of the packages of a software release are very small in
 size, there are some that are quite large. There are too many
@@
 too large to distribute in its entirety. In some archives there are
-also divisions of the archive into sections, e.g. by the computer
+also divisions of the archive into sections, e.g. by the operating system (OS) or computer
 architecture that the package is intended for.
@@
 
-For example, Figure~\ref{size_CDF} shows the size of packages in the
-Debian distribution. While 80\% of the packages are less than
+For example, Figure~\ref{size_CDF} shows the size of the packages in the
+current Debian distribution. While 80\% of the packages are less than
 512~KB, some of the packages are hundreds of megabytes. The entire
 archive consists of 22,298 packages and is approximately 119,000 MB
-in size. Some packages are divided by architecture, but there are a
-large number of packages that are not architecture specific and so
-can be installed on any computer architecture.
+in size. Most of the packages are to be installed in any computer environment, but there are 
+also OS- or architecture-specific packages, as shown by the adjusted sizes based on popularity of the packages. ((((more words on how the size is adjusted by popularity))))
 
@@
 updated. These updates could be the result of the software creators
-releasing a new version of the software with improved functionality,
-or the result of the distributor updating their packaging of the
+releasing a new version with improved functionality,
+or the distributor updating their packaging of the
 software to meet new requirements. Even if the distributor
@@
 the Debian archive that was updated each day over a period of 3
-months. Each day approximately 1.5\% of the 119,000 MB archive is
-updated with new versions of packages.
+months. In every single day, approximately 1.5\% of the 119,000 MB archive is
+updated with new versions of packages. Note that this frequency is much higher than
+that of most commercial software, mainly because many free software are
+developed in a loosely management environment with developers working
+asynchronously from worldwide.
 
@@
 
-Though there are a large number of packages, and a large number of
-users, interest in a given version of a package could be very
-limited. If the package is necessary for the system (e.g. the
-package distribution software), then it is likely that everyone will
-have it installed, but there are very few of these packages. Most
-packages fall in the category of optional or extra, and so are
+Finally, though there are a large number of packages and a large number of
+users, the interest in a particular version of a package can be very 
+limited. Specifically, there are core packages that every user has to download, but most
+packages would fall in the category of optional or extra, and so are
 interesting to only a limited number of people.
@@
 
-\subsubsection{Interactive Users}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The package management software that downloads packages displays
-some kind of indication of speed and completeness for users to
-watch. Since previous client-server downloads occur in a sequential
-fashion, the package management software also measures the speed
-based on sequential downloading. This requires the peer-to-peer
-solution to be reasonably responsive at retrieving packages
-sequentially.
+\subsection{Why BitTorrent Doesn't Work Well}
+\label{related}
+
+Recently, many distributors make their software available using
+BitTorrent \cite{COHEN03}, in particular, for the distribution of CD
+images. This straightforward use however is far ineffective, as it requires the
+peers to download large numbers of packages that they are not
+interested in, and prevents them from updating to new packages
+without downloading another image containing a lot of the same
+packages they already have.
+
+An alternative is to create torrents tracking individual packages. Unfortunately, we find that this enhancement can be
+quite difficult given the unique characteristic of free software packages.
+
+First, there is no obvious way to divide the packages into torrents.
+Most of the packages are too small, and there are too many packages
+in the entire archive to create individual torrents for each one.
+On the other hand, all the packages together are too large to track
+efficiently as a single torrent. Hence, some division of the archive's
+packages into torrents is obviously necessary, but wherever that
+split occurs it will cause either some duplication of connections,
+or prevent some peers from connecting to others who do have the same
+content. In addition, a small number of the packages can be updated every
+day which would add new files to the torrent, thereby changing its
+\emph{infohash} identifier and making it a new torrent. This will
+severely fracture the download population, even though peers in the
+new torrent may share 99\% of the packages in common with peers in the
+old torrent.
+
+Other issues also prevent BitTorrent from being a good solution to
+this problem. In particular, BitTorrent's fixed piece sizes (?KB) that disregard file
+boundaries are bigger than many of the packages in the archive. This
+will waste peers' downloading bandwidth as they will end up
+downloading parts of other packages just to get the piece that
+contains the package they do want.
+
+Finally, note that BitTorrent downloads files
+randomly, which does not work well with the interactive package
+management tools expectation of sequential downloads. On the other hand, with altruistic peers, incentives to share (upload) 
+become a less important issue, and the availability of seeds are not critical, either, as the mirror sites 
+can serve in that capacity.
 
@@
 
-\section{Related Work/Existing Solutions}
-\label{related}
-
-There are many peer-to-peer implementations available today, but
-none is very well suited to this specific problem.
-
-\subsection{BitTorrent}
-\label{bittorrent}
-
-Many distributors make their software available in some form using
-BitTorrent \cite{COHEN03}, e.g. for the distribution of CD
-images. This is not an ideal situation though, as it requires the
-peers to download large numbers of packages that they are not
-interested in, and prevents them from updating to new packages
-without downloading another image containing a lot of the same
-packages they already have. Unfortunately, BitTorrent is not ideally
-suited to an application such as this one.
-
-First, there is no obvious way to divide the packages into torrents.
-Most of the packages are too small, and there are too many packages
-in the entire archive, to create individual torrents for each one.
-However, all the packages together are too large to track
-efficiently as a single torrent. Some division of the archive's
-packages into torrents is obviously necessary, but wherever that
-split occurs it will cause either some duplication of connections,
-or prevent some peers from connecting to others who do have the same
-content. Also, a small number of the packages can be updated every
-day which would add new files to the torrent, thereby changing its
-\emph{infohash} identifier and making it a new torrent. This will
-severely fracture the download population, even though peers in the
-new torrent share 99\% of the packages in common with peers in the
-old torrent.
-
-Other issues also prevent BitTorrent from being a good solution to
-this problem. BitTorrent's fixed piece sizes that disregard file
-boundaries are bigger than many of the packages in the archive. This
-will waste peers' downloading bandwidth as they will end up
-downloading parts of other packages just to get the piece that
-contains the package they do want. Incentives to share (upload) are
-no longer needed, as the software is freely available for anyone to
-download without sharing. Seeds are also not needed as the mirrors
-can serve in that capacity. Finally, BitTorrent downloads files
-randomly, which does not work well with the interactive package
-management tools expectation of sequential downloads.
-
-\subsection{Other Solutions}
-\label{others}
-
-There have also been other attempted implementations, usually based
-on some form of BitTorrent, to address this problem. Some other
-implementations have used BitTorrent almost unmodified, while others
-have only looked at using DHTs to replace the tracker in a
-BitTorrent system. apt-torrent \cite{apttorrent} creates torrents
-for some of the larger packages available, but it ignores the
-smaller packages, which are often the most popular. DebTorrent
-\cite{debtorrent} makes widespread modifications to a traditional
-BitTorrent client, to try and fix the drawbacks mentioned in section
-\ref{bittorrent}. However, these changes also require some
-modifications to the distribution system to support it.
-
-Others have also used DHTs to support this type of functionality.
-Kenosis \cite{kenosis} is a P2P Remote Procedure Call
-client also based on the Kademlia DHT, but it is meant as a P2P
-primitive system on which other tools can be built, and so it has no
-file sharing functionality at all. Many have used a DHT as a drop-in
-replacement for the tracker in a BitTorrent system
-\cite{bittorrent-dht, azureus-dht}, but such systems only use the
-DHT to find peers for the same torrent, so the file sharing uses
-traditional BitTorrent and so is not ideal for the reasons listed
-above.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Opportunity/Our Proposal}
+\section{Peer-to-Peer Assisted Distributor: An Overview}
 \label{opportunity}
@@
 
-\section{Sample Implementation/Implementation}
+\section{\texttt{apt-p2p}: A Practical Implementation}
 \label{implementation}
@@
 
-\section{Analysis}
+\section{Performance Evaluation}
 \label{analysis}
@@
 
+\section{Related Work}
+\label{others}
+
+There have also been preliminary attempts to implement peer-to-peer distributors for
+software packages. apt-torrent \cite{apttorrent} creates torrents
+for some of the larger packages available, but it ignores the
+smaller packages, which are often the most popular. DebTorrent
+\cite{debtorrent} makes widespread modifications to a traditional
+BitTorrent client, to try and fix the drawbacks mentioned in section
+\ref{bittorrent}. However, these changes also require some
+modifications to the distribution system to support it.
+
+Others have also used DHTs to support this type of functionality.
+Kenosis \cite{kenosis} is a P2P Remote Procedure Call
+client also based on the Kademlia DHT, but it is meant as a P2P
+primitive system on which other tools can be built, and so it has no
+file sharing functionality at all. Many have used a DHT as a drop-in
+replacement for the tracker in a BitTorrent system
+\cite{bittorrent-dht, azureus-dht}, but such systems only use the
+DHT to find peers for the same torrent, so the file sharing uses
+traditional BitTorrent and so is not ideal for the reasons listed
+above.
+
+Our solution differs from them in that ...
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Future Work}
-\label{future}
+\section{Conclusion and Future Work}
+\label{conclusions}
+
+We have designed a generally applicable peer-to-peer content
+distribution model to be used by many of the free software
+distributors operating today. It makes use of already existing
+features of most package management systems to create a
+peer-to-peer distribution that should substantially reduce the costs
+of hosting the software packages.
+
+We have also implemented our design in freely available software to
+be used in conjuction with Debian-based distribution of Linux
+software packages. It is currently in use by some users of the
+Debian project's distribution, and so serves as an example of the
+possibilities that exist.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusions}
-\label{conclusions}
-
-We have designed a generally applicable peer-to-peer content
-distribution model to be used by many of the free software
-distributors operating today. It makes use of already existing
-features of most package management systems to create a
-peer-to-peer distribution that should substantially reduce the costs
-of hosting the software packages.
-
-We have also implemented our design in freely available software to
-be used in conjuction with Debian-based distribution of Linux
-software packages. It is currently in use by some users of the
-Debian project's distribution, and so serves as an example of the
-possibilities that exist.
-
 \bibliographystyle{IEEEtran}

### 2008-08-26 docs/paper/paper.tex
More work on the paper, including adding a new response time improvement figure.

@@
 
-The rest of this paper is organized as follows. The background and motivation are presented in Section \ref{situation}, together with related works in  Section \ref{related}. We propose
-our solution in Section \ref{opportunity}. We then detail our sample
-implementation for Debian-based distributions in Section
-\ref{implementation}, including an in-depth look at our DHT
-customizations in Section \ref{custom_dht}. Its performance is evaluated in Section \ref{analysis}. Finally, 
-Section \ref{conclusions} concludes the paper and offers some future directions.
+The rest of this paper is organized as follows.
+The background and motivation are presented in Section~\ref{situation}, and we analyze BitTorrent's use for this purpose in Section~\ref{bittorrent}.
+We propose
+our solution in Section~\ref{opportunity}. We then detail our sample
+implementation for Debian-based distributions in Section~\ref{implementation},
+including an in-depth look at our DHT
+customizations in Section~\ref{custom_dht}. Its performance is evaluated in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
+Section~\ref{conclusions} concludes the paper and offers some future directions.
 
@@
 \subsection{Why BitTorrent Doesn't Work Well}
-\label{related}
+\label{bittorrent}
 
@@
 
-The situation described in section \ref{situation} presents a clear
+The situation described in Section~\ref{situation} presents a clear
 opportunity to use some form of peer-to-peer file-sharing protocol
@@
 Twisted in our sample implementation for all other networking needs.
-More details of this customized DHT can be found below in section
-\ref{custom_dht}.
+More details of this customized DHT can be found below in
+Section~\ref{custom_dht}.
 
@@
 For large files (5 or more pieces), the torrent strings described in
-section \ref{downloading}
+Section~\ref{downloading}
 are too long to store with the peer's download info in the DHT. This
@@
 
-Most of our customizations to the DHT have been to try and improve
+Many of our customizations to the DHT have been to try and improve
 the time of the recursive \texttt{find\_value} requests, as this can
@@
 before a node is removed from the routing table.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{apt_p2p_improvements-find_value.eps}
+\caption{The distribution of average response times PlanetLab nodes
+experience for \texttt{find\_value} queries. The original DHT
+implementation results are shown, as well as the successive
+improvements that we made to reduce the response time.}
+\label{improvements}
+\end{figure}
+
+To test our changes during development, we ran the customized DHT
+for several hours after each major change on 300 PlanetLab nodes
+\cite{planetlab}. Though the nodes are not expected to be firewalled
+or NATted, some can be quite overloaded and so consistently fail to
+respond within a timeout period, similar to NATted peers. The
+resulting distribution of the nodes' average response times is shown
+in Figure~\ref{improvements}. Each improvement successfully reduced
+the response time, for a total reduction of more than half. The
+final distribution is also narrower, as the improvements make the
+system more predictable. However, there are still a large number of
+outliers with higher average response times, which are the
+overloaded nodes on PlanetLab. This was confirmed by examining the
+average time it took for a timeout to occur, which should be
+constant as it is a configuration option, but can be much larger if
+the node is too overloaded for the program to be able to check for a
+timeout very often.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Related Work}
-\label{others}
+\label{related}
 
@@
 \cite{debtorrent} makes widespread modifications to a traditional
-BitTorrent client, to try and fix the drawbacks mentioned in section
-\ref{bittorrent}. However, these changes also require some
-modifications to the distribution system to support it.
+BitTorrent client, to try and fix the drawbacks mentioned in
+Section~\ref{bittorrent}. However, these changes also require some
+modifications to the distribution system to support it. Our system
+considers all the files available to users to download, and makes
+use of the existing infrastructure unmodified.
 
@@
 traditional BitTorrent and so is not ideal for the reasons listed
-above.
+in Section~\ref{bittorrent}.
 
-Our solution differs from them in that ...
+% \cite{deltacast}
+
+There are a number of works dedicated to developing a collaborative
+content distribution network (CDN) using peer-to-peer techniques.
+Freedman et. al. developed Coral \cite{coral} using a distrbitued
+\emph{sloppy} hash table to speed request times. Pierre and van
+Steen developed Globule \cite{globule} which uses typical DNS and
+HTTP redirection techniques to serve requests from a network of
+replica servers, which in turn draw their content from the original
+location (or a backup). Shah et. al. \cite{shah08} analyze an
+existing software delivery system and use the results to design a
+peer-to-peer content distribution network that makes use of
+volunteer servers to help with the load. None of these systems meets
+our goal of an even distribution of load amongst the users of the
+system. Not all users of the systems become peers, and so are not
+able to contribute back to the system after downloading. The
+volunteers that do contribute as servers are required to contribute
+larger amounts of bandwidth, both for uploading to others, and in
+downloading content they are not in need of in order to share them
+with other users. Our system treats all users equally, requiring all
+to become peers in the system, sharing the uploading load equally
+amongst all, but doesn't require any user to download files they
+would not otherwise need.
+
+The most similar works to ours are by Shah et. al. \cite{shah08} and
+Shark by Annapureddy et. al. \cite{shark}.
+Shah's system, in addition to the drawbacks mentioned previously,
+is not focused on the interactivity of downloads, as
+half of all requests were required ``to wait between 8 and 15
+minutes.'' In contrast, lookups in our system take only seconds to
+complete, and all requests can be completed in under a minute.
+Shark makes use of Coral's distributed sloppy hash table to speed
+the lookup time, but their system is more suited to its intended use
+as a distributed file server. It doesn't make use of authoritative
+copies of the original files, allowing instead any users in the
+system to update files and propagate those changes to others. Our
+system is well-tailored to the application of disseminating the
+unchanging software packages from the authoritative sources to all
+users unchanged.
 

### 2008-08-28 docs/paper/paper.tex
Added a new figure and lots of work on the paper.

@@
 
-\title{\texttt{apt-p2p}: A Peer-to-Peer Distributor for Free Software Package Release and Update}
+\title{\texttt{apt-p2p}: A Peer-to-Peer Distribution System for Software Package Releases and Updates}
 \author{\IEEEauthorblockN{Cameron Dale}
@@
 distribution, especially due to the altruistic nature of many of
-these users. There are no existing solution suitable for this
+these users. There are no existing solutions suitable for this
 situation, so we present a new technique for satisfying the needs of
-this P2P distribution, which is generally applicable to many of
+this P2P distribution which is generally applicable to many of
 these distributors' systems. Our method makes use of a DHT for
 storing the location of peers, using the cryptographic hash of the
-package as a key. To show the simplicity and functionality, we
+package as a key. To show the simplicity and functionality of our model, we
 implement a solution for the distribution of Debian software
-packages, including the many DHT customizations needed. Finally, we
+packages, including details on the DHT customizations and system optimizations needed. Finally, we
 analyze our system to determine how it is performing and what effect
@@
 BitTorrent. The model relies on the pre-existence of cryptographic
-hashes of the packages, which should uniquely identify it for a
+hashes of the packages, which uniquely identify it for a
 request from other peers. If the peer-to-peer download fails, then
@@
 distributors of the opportunity that can be met with such a system.
+Finally, we analyze our currently deployed implementation to
+determine how effective it is at meeting our goals, and to see what
+effect it is having on the Debian package distribution system.
 
 The rest of this paper is organized as follows.
-The background and motivation are presented in Section~\ref{situation}, and we analyze BitTorrent's use for this purpose in Section~\ref{bittorrent}.
+The background and motivation are presented in Section~\ref{situation}, including an analysis of BitTorrent's use for this purpose in Section~\ref{bittorrent}.
 We propose
@@
 including an in-depth look at our DHT
-customizations in Section~\ref{custom_dht}. Its performance is evaluated in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
+customizations in Section~\ref{custom_dht}. The performance of our implementation is evaluated in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
 Section~\ref{conclusions} concludes the paper and offers some future directions.
@@
 collaboratively develop and release their software. The ever increasing power of
-modern programming languages and operating systems has made these software, like commercial software, extremely large and complex, which often
-consists of many small units (packages). Together with their popularity among users, 
-an efficient and reliable management and distribution of these packages over the Internet has become an daunting task. In this section, we offer concrete examples illustrating the 
+modern programming languages and operating systems has made this software, like commercial software, extremely large and complex, though it is often
+distributed in many small units (packages). Together with their popularity among users, 
+an efficient and reliable management and distribution of these packages over the Internet has become a daunting task. In this section, we offer concrete examples illustrating the 
 unique challenges in this context.
 
-\subsection{Free Software Package Distributor: Examples}
+\subsection{Free Software Package Distributors}
 \label{examples}
@@
 package management tool that requests packages from websites. There
-are two software distribution systems for Mac OSX, fink and
+are two software distribution systems for software that runs on the Macintosh OS, fink and
 MacPorts, that also retrieve packages in this way.
@@
 verification file to be downloaded next to the desired
-file. The hash file usually have the same file name, but with an
+file. The hash file usually has the same file name, but with an
 added extension identifying the hash used (e.g. \texttt{.md5} for
@@
 
-Given the free nature of these software, there are often a number of users 
-motivated by altruism to want to help out with their distribution. 
-This is particularly true considering that many of these software are used by
+Given the free nature of this software, there are often a number of users 
+motivated by altruism to want to help out with the distribution.
+This is particularly true considering that many of this software is used by
 groups that are staffed mostly, or sometimes completely, by
 volunteers. They are thus motivated to contribute their network resources, so as to promote the healthy development 
-of the volunteer community that released the software. As a matter of fact,
-we have see many free mirror sites hosting these software packages for downloading. 
+of the volunteer community that released the software.
 We also naturally expect that peer-to-peer distribution can be implementation in 
@@
 in size. Most of the packages are to be installed in any computer environment, but there are 
-also OS- or architecture-specific packages, as shown by the adjusted sizes based on popularity of the packages. ((((more words on how the size is adjusted by popularity))))
+also OS- or architecture-specific packages.
 
@@
 \includegraphics[width=\columnwidth]{size-quarter.eps}
-\caption{The amount of data in the Debian archive that is updated
-each day, broken down by architecture.}
+\caption{The amount of data in the 119,000 MB Debian archive that is
+updated each day, broken down by architecture.}
 \label{update_size}
@@
 the Debian archive that was updated each day over a period of 3
-months. In every single day, approximately 1.5\% of the 119,000 MB archive is
-updated with new versions of packages. Note that this frequency is much higher than
-that of most commercial software, mainly because many free software are
-developed in a loosely management environment with developers working
-asynchronously from worldwide.
+months. Every single day, approximately 1.5\% of the 119,000 MB archive is
+updated with new versions of packages. This frequency is much higher than
+that of most commercial software, mainly because much of free software is
+developed in a loosely managed environment of developers working
+asynchronously on a worldwide scale.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Interactive Users}
+
+The package management software that downloads packages displays
+some kind of indication of speed and completeness for users to
+watch. Since previous client-server downloads occurred in a sequential
+fashion, the package management software also measures the speed
+based on sequential downloading. This requires the peer-to-peer
+solution to be reasonably responsive at retrieving packages
+sequentially.
 
@@
 
-Recently, many distributors make their software available using
-BitTorrent \cite{COHEN03}, in particular, for the distribution of CD
-images. This straightforward use however is far ineffective, as it requires the
+Many distributors make their software available using
+BitTorrent \cite{COHEN03}, in particular for the distribution of CD
+images. This straightforward use however can be very ineffective, as it requires the
 peers to download large numbers of packages that they are not
@@
 
-An alternative is to create torrents tracking individual packages. Unfortunately, we find that this enhancement can be
+An alternative is to create torrents tracking smaller groups of packages. Unfortunately, we find that this enhancement can be
 quite difficult given the unique characteristic of free software packages. 
-
 First, there is no obvious way to divide the packages into torrents.
@@
 Other issues also prevent BitTorrent from being a good solution to
-this problem. In particular, BitTorrent's fixed piece sizes (?KB) that disregard file
+this problem. In particular, BitTorrent's fixed piece sizes (usually 512 KB) that disregard file
 boundaries are bigger than many of the packages in the archive. This
@@
 contains the package they do want. 
-
 Finally, note that BitTorrent downloads files
 randomly, which does not work well with the interactive package
-management tools expectation of sequential downloads. On the other hand, with altruistic peers, incentives to share (upload) 
-become a less important issue, and the availability of seeds are not critical, either, as the mirror sites 
-can serve in that capacity.
+management tools expectation of sequential downloads.
+
+On the other hand, there are aspects of BitTorrent that are no
+longer needed in a system such as this one. With altruistic peers
+and all files being available without uploading, incentives to share
+become a less important issue. Also, the availability of seeds are
+not critical either, as the mirror sites serve in that capacity
+already.
 
@@
 
-The situation described in Section~\ref{situation} presents a clear
-opportunity to use some form of peer-to-peer file-sharing protocol
-to allow willing users to contribute upload bandwidth. This sparse
+We assume that the user is attempting to download packages from a
+server, using package management software to find and retrieve the
+packages. Further, we assume that the requests from the user to the
+server are able to be proxied by our peer-to-peer program, and that
+the server is always available and has all of the package files.
+Finally, the cryptographic hash of the packages must be available
+separately from the package itself, and is usually contained in an
+\emph{index} file which also contains all the packages' names,
+locations and sizes.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{model_simple.eps}
+\caption{The different phases of functionality of our peer-to-peer distribution model.}
+\label{model}
+\end{figure}
+
+Our model for using P2P to enhance such a system is shown in
+Figure~\ref{model}. As shown in Phase~1, our program will act as a
+proxy (1,2), downloading (3) and caching all files communicated
+between the user and the server (4). It will therefore also have
+available the index files containing the cryptographic hashes all
+packages. Later, in Phase~2, upon receiving a request from the user
+to download a package (5), our program will search the index files
+for the package being requested and find its hash (6). This hash can
+then be looked up recursively in the DHT (7), which will return a
+list of peers that have the package already (8). The package can
+then be downloaded from the peers (11,12), it can be verified using
+the hash (13), and if valid can be returned to the user (14). The
+current nodes location is also added to the DHT for that hash (15),
+as it is now a source for others to download from.
+
+In steps (11,12), the fact that this package is also available to download for free
+from a server is very important to our proposed model. If the package hash
+can not be found in the DHT, the peer can then fallback to
+downloading from the original location (i.e. the network of
+mirrors). The mirrors thus, with no modification to their
+functionality, serve as seeds for the packages in the peer-to-peer
+system. Any packages that have just been updated, or that are very
+rare, and so don't have any peers available, can always be found on
+the mirror. Once the peer has completed the download from the mirror
+and verified the package, it can then add itself to the DHT as the
+first peer for the new package, so that future requests for the package
+will not need the mirror.
+
+This sparse
 interest in a large number of packages undergoing constant updating
@@
 
-The fact that this package is also available to download for free
-from a server is very important to our proposal. If the package hash
-can not be found in the DHT, the peer can then fallback to
-downloading from the original location (i.e. the network of
-mirrors). The mirrors thus, with no modification to their
-functionality, serve as seeds for the packages in the peer-to-peer
-system. Any packages that have just been updated, or that are very
-rare, and so don't have any peers available can always be found on
-the mirror. Once the peer has completed the download from the mirror
-and verified the package, it can then add itself to the DHT as the
-first peer for the new package, so that future requests from peers
-will not need the mirror.
-
-The trust of the package is also always guaranteed through the use
+Note that, despite downloading the package from untrustworthy peers,
+the trust of the package is always guaranteed through the use
 of the cryptographic hashes. Nothing can be downloaded from a peer
 until the hash is looked up in the DHT, so a hash must first come
-from a trusted source (i.e. a mirror). Most distributors use index
+from a trusted source (i.e. the distributors server). Most distributors use index
 files that contain hashes for a large number of the packages in
@@
 listening for requests from the package management tool for package
-files. It will get uncached requests first from the P2P system, or
-falling back to the normal HTTP request from a server should it not
+files, as well as serving (uploading) cached package to other peers.
+It will get uncached requests first from the P2P system, or
+fall back to the normal HTTP request from a server should it not
 be found. For methods that don't use HTTP requests, other types of
@@
 \emph{torrent string} containing the peer's hashes of the pieces of
-the larger files. These piece hashes could be compared ahead of time
+the larger files. These piece hashes are compared ahead of time
 to determine which peers have the same piece hashes (they all
@@
 distributors to download and modify \cite{apt-p2p}. This software,
-called \texttt{apt-p2p}, interacts with the \texttt{apt} tool, which
+called \texttt{apt-p2p}, interacts with the \texttt{apt} tool which
 is found in most Debian-based Linux distributions. \texttt{apt} uses
@@
 
-Since all requests from apt are in the form of HTTP downloads from a
+Since all requests from \texttt{apt} are in the form of HTTP downloads from a
 server, the implementation takes the form of a caching HTTP proxy.
@@
 the mirror name in \texttt{apt}'s configuration file (i.e.
-``localhost:9977/mirrorname.debian.org/\ldots'').
+``http://localhost:9977/mirrorname.debian.org/\ldots'').
 
@@
 requests to a peer, and the requesting of smaller pieces of a large
-file using the Range request header.
+file using the HTTP Range request header.
 
@@
 
-\section{Customized DHT}
+\section{Customized DHT / System Optimization}
 \label{custom_dht}
@@
 
-\subsection{Kademlia Background}
-\label{kademlia}
+\subsection{DHT Background}
+\label{dht}
+
+DHT's operate by storing (\emph{key}, \emph{value}) pairs in a
+distributed fashion such that no node will, on average, store more
+than any other node. They support two primitive operations:
+\emph{put}, which takes a key and a value and stores it in the DHT;
+and \emph{get}, which takes a key and returns a value (or values)
+that was previously stored with that key. These operations are
+recursive, as each node does not know about all the other nodes in a
+DHT, and so must recursively search for the correct node to put to
+or get from.
 
@@
 node to quickly traverse the DHT to find the nodes close to the
-desired key. The only difference between them is that the
+desired key. The only difference between \texttt{find\_node} and \texttt{find\_value} is that the
 \texttt{find\_value} query will cause a node to return a value, if
@@
 
-Figure \ref{size_CDF} shows the package size of the 22,298 packages
+Figure \ref{size_CDF} shows the size of the 22,298 packages
 available in Debian in January 2008. We can see that most of the
@@
 To test our changes during development, we ran the customized DHT
-for several hours after each major change on 300 PlanetLab nodes
+for several hours after each major change on over 300 PlanetLab nodes
 \cite{planetlab}. Though the nodes are not expected to be firewalled
@@
 
-The original design of Kademlia specified that each keywould have
+The original design of Kademlia specified that each key would have
 only a single value associated with it. The RPC to find this value
@@
 of the closest to the key being searched for, then the values it is
-returning will probably be the staler ones in the system, and it
+returning will probably be the staler ones in the system, as it
 will not have the latest stored values. However, the search for
-closer nodes will stop here, as the queried node only returned the
+closer nodes will stop here, as the queried node only returned
 values and not a list of nodes to recursively query. We could have
@@
 would severely limit the size and number of the values that could be
-returned.
+returned in a single UDP packet.
 
@@
 
-\subsection{BitTorrent's Improvements}
+\subsection{\textbf{OPTIONAL}: BitTorrent's Improvements}
 \label{bittorrent_dht}
 
-In the many years that some BitTorrent clients have been using a
-Kademlia based DHT for tracker-less operation, the developers have
+In the several years that some BitTorrent clients have been using a
+Kademlia-based DHT for tracker-less operation, the developers have
 made many enhancements which we can take advantage of. One of the
@@
 which is very unlikely in our system given the long uptimes of
-nodes. We also increased the number of concurrent outstanding
+nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
+We also increased the number of concurrent outstanding
 requests allowed from 3 to 6 to speed up the recursive key finding
@@
 
-\subsection{Other Changes}
+\subsection{\textbf{OPTIONAL}: Other Changes}
 \label{other_changes}
@@
 users who connect regularly (weekly), and we have found 186 unique
-users in the 2 months of our analysis. We determined which users are
+users in the 2 months of our analysis.
+
+We also determined which users are
 behind a firewall or NAT, which is one of the main problems of
@@
 recently, which will cause the peer to wait for a timeout to occur
-(currently set at 9 seconds) before moving on. They will also be
+(currently 9 seconds) before moving on. They will also be
 unable to contribute any upload bandwidth to other peers, as all
@@
 every time their computer boots up, peers are expected to stay for a
-long period in the system. 50\% of connections last longer than 5
+long period in the system.
+Indeed, we find that 50\% of connections last longer than 5
 hours, and 20\% last longer than 10 hours. These connections are
@@
 \includegraphics[width=\columnwidth]{AptP2PDuration-ind_peers.eps}
-\caption{The CDF of the average time individual peers stay in the
+\caption{\textbf{OPTIONAL}: The CDF of the average time individual peers stay in the
 system.}
@@
 
-We also examined the average time each individual peer spends in the
+\textbf{OPTIONAL}: We also examined the average time each individual peer spends in the
 system. Figure~\ref{duration_peers} shows the cumulative
@@
 analyzed our system for this parameter.
-Figure~\ref{duration_online_1} shows the fraction of peers will
+Figure~\ref{duration_online_1} shows the fraction of peers that will
 remain online for another hour, as a function of how long they have
@@
 \includegraphics[width=\columnwidth]{AptP2PDuration-online_6.eps}
-\caption{The fraction of peers that, given their current duration in
+\caption{\textbf{OPTIONAL}: The fraction of peers that, given their current duration in
 the system, will stay online for another 6 hours.}
@@
 
-Since our peers are much longer-lived than other P2P systems, we
+\textbf{OPTIONAL}: Since our peers are much longer-lived than other P2P systems, we
 also looked at the fraction of peers that stay online for another 6
@@
 it uses the same port for both its DHT (UDP) requests and download
-(TCP) requests (which is also a configuration parameter).
+(TCP) requests (which is also the default configuration parameter).
 
@@
 \includegraphics[width=\columnwidth]{AptP2PDownloaded-bw.eps}
-\caption{The bandwidth of data that the contacted peers have
+\caption{\textbf{OPTIONAL}: The bandwidth of data that the contacted peers have
 downloaded and uploaded.}
@@
 \cite{STUN} to circumvent the NATs, and by better exclusion of
-NATted peers from the DHT (which will not prevent them from using
+NATted peers from the DHT (which does not prevent them from using
 the system).
@@
 
-There have also been preliminary attempts to implement peer-to-peer distributors for
+There have been other preliminary attempts to implement peer-to-peer distributors for
 software packages. apt-torrent \cite{apttorrent} creates torrents
@@
 
-Others have also used DHTs to support this type of functionality.
+\textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
 Kenosis \cite{kenosis} is a P2P Remote Procedure Call
@@
 content distribution network (CDN) using peer-to-peer techniques.
-Freedman et. al. developed Coral \cite{coral} using a distrbitued
+Freedman et. al. developed Coral \cite{coral} using a distributed
 \emph{sloppy} hash table to speed request times. Pierre and van
@@
 Shark by Annapureddy et. al. \cite{shark}.
-Shah's system, in addition to the drawbacks mentioned previously,
+Shah's system, in addition to the drawbacks mentioned above,
 is not focused on the interactivity of downloads, as
@@
 unchanging software packages from the authoritative sources to all
-users unchanged.
+users.
 
@@
 
-One aspect missing from our model is the removal of old packages
+\textbf{OPTIONAL}: One aspect missing from our model is the removal of old packages
 from the cache. Since our implementation is still relatively young,

### 2008-08-29 docs/paper/paper.tex
JC's updated changes to the INFOCOM paper.

@@
 \begin{abstract}
-A large amount of free software packages are available over the
-Internet from many different distributors. Most of these
-distributors use the traditional client-server model to handle
-requests from users. However, there is an excellent opportunity to
-use peer-to-peer techniques to reduce the cost of much of this
-distribution, especially due to the altruistic nature of many of
-these users. There are no existing solutions suitable for this
-situation, so we present a new technique for satisfying the needs of
-this P2P distribution which is generally applicable to many of
-these distributors' systems. Our method makes use of a DHT for
-storing the location of peers, using the cryptographic hash of the
-package as a key. To show the simplicity and functionality of our model, we
-implement a solution for the distribution of Debian software
-packages, including details on the DHT customizations and system optimizations needed. Finally, we
-analyze our system to determine how it is performing and what effect
-it is having.
+The Internet has become an cost-effective
+vehicle for software development and release, particular in the free software society.
+Given the free nature of these software, there are often a number of users
+motivated by altruism to help out with the distribution, so as to promote the healthy development
+of this voluntary society. It is thus naturally expected that peer-to-peer distribution can be implemented,
+which scale well with large user bases and can easily explore the network resources made available by
+the volunteers.
+
+Unfortunately, this application scenario has many unique characteristics, which
+make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
+a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
+too large to be distributed in its entirety. The packages are also being constantly
+updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
+limited depending on the computer platforms and operating systems used.
+
+In this paper, we propose a novel peer-to-peer assisted distributor design that
+addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
+for software packages.
+Our design leads to \texttt{apt-p2p}, a practical implementation that extends the popular \texttt{apt} distributor.
+\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
+software packages and will also be available in the next release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
+response time reduction, and multi-value extension. They together ensure
+that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
+sheer bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
 \end{abstract}
@@
 
-There are a large number of free software package distributors using
-package distribution systems over the Internet to distribute
-software to their users. These distributors have developed many
-different methods for this distribution, but they almost exclusively
-use a client-server model to satisfy user requests. The popularity
-and number of users results in a large number of requests, which
-usually requires a network of mirrors to handle. Due to the free
-nature of this software, many users are willing and able to
-contribute upload bandwidth to this distribution, but have no
-current way to do this.
+With the widespread penetration of broadband accesses, the Internet has become an cost-effective
+vehicle for software development and release. This is particularly true for the free software
+society whose developers and users are worldwide distributed and work asynchronously. The ever increasing power of
+modern programming languages, computer platforms, and operating systems has made these software extremely large and complex,
+which are often divided into a huge number of small packages.
+Together with their popularity among users,
+an efficient and reliable management and distribution of these software over the Internet has become a challenging task.
 
-We present a new peer-to-peer distribution model to meet these
-demands. It is based on many previous implementations of successful
-peer-to-peer protocols, especially distributed hash tables (DHT) and
-BitTorrent. The model relies on the pre-existence of cryptographic
-hashes of the packages, which uniquely identify it for a
-request from other peers. If the peer-to-peer download fails, then
-the original request to the server is used as a fallback to prevent
-any dissatisfaction from users. The peer can then share this new
-package with others through the P2P system.
+The existing distributor for free software are mostly based on the client/server model, e.g.,
+the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
+Given the free nature of these software, there are often a number of users
+motivated by altruism to help out with the distribution, so as to promote the healthy development
+of this voluntary society.
+We thus naturally expect that peer-to-peer distribution can be implemented in
+this context, which scale well with large user bases and can easily explore the network resources made available by
+the volunteers.
 
-First, we examine the opportunity that is available for many of
-these free software package distributors. We present an overview of
-a system that would efficiently satisfy the demands of a large
-number of users, and should significantly reduce the currently
-substantial bandwidth requirements of hosting this software. We then
-present an example implementation based on the Debian package
-distribution system. This implementation will be used by a large
-number of users, and serves as an example for other free software
-distributors of the opportunity that can be met with such a system.
-Finally, we analyze our currently deployed implementation to
+Unfortunately, this application scenario has many unique characteristics, which
+make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
+there are too many packages to distribute each individually, but the archive is
+too large to distribute in its entirety. The packages are also being constantly
+updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
+limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
+sequential order used in existing software package distributors. This in turn suppresses interaction with users
+given the difficulty in tracking speed and downloading progress.
+
+In this paper, we propose a novel peer-to-peer assisted distributor design that
+addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
+for software packages.
+Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian \footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
+distribution system.  We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
+response time reduction, and multi-value extension. They together ensure
+that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
+sheer bandwidth requirements of hosting the software.
+
+\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
+software packages and will also be available in the next release of Ubuntu. We have evaluated our currently deployment  to
 determine how effective it is at meeting our goals, and to see what
-effect it is having on the Debian package distribution system.
+effect it is having on the Debian package distribution system. In particular, our existing real user statistics
+have suggested that it responsively interacts with clients and substantially reduces server cost.
 
@@
 
-The package management software that downloads packages displays
+Given the relatively long time for software package downloading, existing package management systems generally display
 some kind of indication of speed and completeness for users to
-watch. Since previous client-server downloads occurred in a sequential
+monitor. Since previous client-server downloads occurred in a sequential
 fashion, the package management software also measures the speed
-based on sequential downloading. This requires the peer-to-peer
-solution to be reasonably responsive at retrieving packages
-sequentially.
+based on sequential downloading. To offer comparable user experience, it is naturally to expect that
+the new peer-to-peer solution be reasonably responsive at retrieving packages,
+preferably enabling sequential downloading order, too.
 
@@
 On the other hand, there are aspects of BitTorrent that are no
-longer needed in a system such as this one. With altruistic peers
+longer critical. Specifically, with altruistic peers
 and all files being available without uploading, incentives to share
@@
 
-We assume that the user is attempting to download packages from a
-server, using package management software to find and retrieve the
-packages. Further, we assume that the requests from the user to the
-server are able to be proxied by our peer-to-peer program, and that
-the server is always available and has all of the package files.
-Finally, the cryptographic hash of the packages must be available
+We now present the design of our peer-to-peer assisted distributor for free software package releases and
+updates. A key principle in our design is that the new functionalities implemented in our distributor should be transparent to users,
+thus offering the same experience as using conventional software management systems, despite enhanced efficiency.
+That said, we assume that the user is still attempting to download packages from a
+server, but the requests will be proxied by our peer-to-peer program. The server is always available and has all of the package files.
+In addition, the cryptographic hash of the packages will be available
 separately from the package itself, and is usually contained in an
@@
 
-Our model for using P2P to enhance such a system is shown in
+Our model for using peer-to-peer to enhance package distribution is shown in
 Figure~\ref{model}. As shown in Phase~1, our program will act as a
@@
 for the package being requested and find its hash (6). This hash can
-then be looked up recursively in the DHT (7), which will return a
+then be looked up recursively in an indexing structure (a Distributed Hash Table [] in our implementation) (7), which will return a
 list of peers that have the package already (8). The package can
@@
 
-There are several ways to implement the desired P2P functionality
+There are several ways to implement the desired peer-to-peer functionality
 into the existing package management software. The functionality can
 be directly integrated through modifications to the software, though
-this could be difficult as the P2P functionality should be running
+this could be difficult as the peer-to-peer functionality should be running
 at all times. This is needed both for efficient lookups and to
@@
 requests from web servers to download the packages, which makes it
-possible to implement the P2P aspect as an almost standard HTTP
+possible to implement the peer-to-peer aspect as an almost standard HTTP
 caching proxy. This proxy will run as a daemon in the background,
@@
 files, as well as serving (uploading) cached package to other peers.
-It will get uncached requests first from the P2P system, or
+It will get uncached requests first from the peer-to-peer system, or
 fall back to the normal HTTP request from a server should it not
@@
 protocol that is similar to the protocol used to fetch packages from
-the distributor's servers. This simplifies the P2P program, as it
+the distributor's servers. This simplifies the peer-to-peer program, as it
 can then treat peers and mirrors almost identically when requesting
@@
 package, and not of pieces of that package, we will need to be able
-to store and retrieve these piece hashes using the P2P protocol. In
+to store and retrieve these piece hashes using the peer-to-peer protocol. In
 addition to storing the file download location in the DHT (which
@@
 distributors to download and modify \cite{apt-p2p}. This software,
-called \texttt{apt-p2p}, interacts with the \texttt{apt} tool which
-is found in most Debian-based Linux distributions. \texttt{apt} uses
-SHA1 hashes to verify most downloaded files, including the large
-index files that contain the hashes of the individual packages. We
-chose this distribution system as it is familiar to us, it allows
-software contributions, and there are interesting statistics
+called \texttt{apt-p2p}, interacts with the popular \texttt{apt} tool. This tool 
+is found in most Debian-based Linux distributions, with related statistics
 available for analyzing the popularity of the software packages
@@
 Since all requests from \texttt{apt} are in the form of HTTP downloads from a
-server, the implementation takes the form of a caching HTTP proxy.
+server, our implementation takes the form of a caching HTTP proxy.
 Making a standard \texttt{apt} implementation use the proxy is then
@@
 We created a customized DHT based on Khashmir \cite{khashmir}, which
-is an implementation of Kademlia \cite{kademlia} using methods
-familiar to BitTorrent developers. Khashmir is also the same DHT
+is an implementation of Kademlia \cite{kademlia}. Khashmir is also the same DHT
 implementation used by most of the existing BitTorrent clients to
@@
 requests to a peer, and the requesting of smaller pieces of a large
-file using the HTTP Range request header.
+file using the HTTP Range request header. Like in \texttt{apt}, 
+SHA1 hashes are then used to verify downloaded files, including the large
+index files that contain the hashes of the individual packages.
 
@@
 
-\subsection{DHT Background}
+\subsection{Kademlia DHT Background}
 \label{dht}
@@
 
-\subsection{Response Time}
+\subsection{Response Time Optimization}
 \label{response_time}
@@
 
-\subsection{Multiple Values}
+\subsection{Extension with Multiple Values}
 \label{multiple_values}
@@
 closest to the key being searched for.
-
-\subsection{\textbf{OPTIONAL}: BitTorrent's Improvements}
-\label{bittorrent_dht}
-
-In the several years that some BitTorrent clients have been using a
-Kademlia-based DHT for tracker-less operation, the developers have
-made many enhancements which we can take advantage of. One of the
-most important is a security feature added to stop malicious nodes
-from subscribing other nodes as downloaders. When a node issues a
-request to another node to store its download info for that key, it
-must include a \emph{token} returned by the storing node in a recent
-\texttt{find\_node} request. The token is created by hashing the IP
-address of the requesting peer with a temporary secret that expires
-after several minutes. This prevents the requesting peer from faking
-its IP address in the store request, since it must first receive a
-response from a \texttt{find\_node} on that IP.
-
-We also made some BitTorrent-inspired changes to the parameters of
-the DHT originally specified by the authors of Kademlia. Since, as
-we will show later, peers stay online in our system for much longer
-periods of time, we reduced Kademlia's \emph{k} value from 20 to 8.
-The value is supposed to be large enough such that any given
-\emph{k} nodes are unlikely to fail within an hour of each other,
-which is very unlikely in our system given the long uptimes of
-nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
-We also increased the number of concurrent outstanding
-requests allowed from 3 to 6 to speed up the recursive key finding
-processes.
-
-\subsection{\textbf{OPTIONAL}: Other Changes}
-\label{other_changes}
-
-We added one other new RPC request that nodes can make:
-\texttt{join}. This request is only sent on first loading the DHT,
-and is usually only sent to the bootstrap nodes that are listed for
-the DHT. These bootstrap nodes will respond to the request with the
-requesting peer's IP and port, so that the peer can determine what
-its oustide IP address is and whether port translation is being
-used. In the future, we hope to add functionality similar to STUN
-\cite{STUN}, so that nodes can detect whether they are NATted and
-take appropriate steps to circumvent it.
-
-In addition, we have allowed peers to store values in the DHT, even
-if the hash they are using is not the correct length. Most of the
-keys used in the DHT are based on the SHA1 hash, and so they are 20
-bytes in length. However, some files in the targetted Debian package
-system are only hashed using the MD5 algorithm, and so the hashes
-retrieved from the server are 16 bytes in length. The DHT will still
-store values using these keys by padding the end of them with zeroes
-to the proper length, as the loss of uniqueness from the 4 less
-bytes is not an issue. Also, the DHT will happily store hashes that
-are longer than 20 bytes, such as those from the 32 byte SHA256
-algorithm, by truncating them to the correct length. Since the
-requesting peer will have the full length of the hash, this will not
-affect its attempt to verify the downloaded file.
 
@@
 Figure~\ref{walker_peers}, we see that approximately half of all
-peers suffered from this restriction.
+peers suffered from this restriction. To address this problem, we added one other new RPC request that nodes can make:
+\texttt{join}. This request is only sent on first loading the DHT,
+and is usually only sent to the bootstrap nodes that are listed for
+the DHT. These bootstrap nodes will respond to the request with the
+requesting peer's IP and port, so that the peer can determine what
+its oustide IP address is and whether port translation is being
+used. In the future, we hope to add functionality similar to STUN
+\cite{STUN}, so that nodes can detect whether they are NATted and
+take appropriate steps to circumvent it.
 
@@
 much longer than those reported by Saroiu et. al. \cite{saroiu2001}
-for other P2P systems, which had 50\% of Napster and Gnutella
+for other peer-to-peer systems, which had 50\% of Napster and Gnutella
 sessions lasting only 1 hour.
@@
 
-\textbf{OPTIONAL}: Since our peers are much longer-lived than other P2P systems, we
+\textbf{OPTIONAL}: Since our peers are much longer-lived than other peer-to-peer systems, we
 also looked at the fraction of peers that stay online for another 6
@@
 \textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
-Kenosis \cite{kenosis} is a P2P Remote Procedure Call
-client also based on the Kademlia DHT, but it is meant as a P2P
+Kenosis \cite{kenosis} is a peer-to-peer Remote Procedure Call
+client also based on the Kademlia DHT, but it is meant as a peer-to-peer
 primitive system on which other tools can be built, and so it has no
@@
 
-We have designed a generally applicable peer-to-peer content
-distribution model to be used by many of the free software
-distributors operating today. It makes use of already existing
-features of most package management systems to create a
-peer-to-peer distribution that should substantially reduce the costs
-of hosting the software packages.
+In this paper, we have provided strong evidence that free software package distribution 
+and update exhibit many distinct characteristics, which call for new designs other
+than the existing peer-to-peer systems for file sharing. To this end, we have 
+present \texttt{apt-p2p}, a novel peer-to-peer distributor that sits between
+client and server, providing efficient and transparent downloading and updating services
+for software packages. We have addressed the key design issues in \texttt{apt-p2p}, including DHT customization,
+response time reduction, and multi-value extension. \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
+software packages and will also be available in the next release of Ubuntu. Existing real user statistics
+have suggested that it interacts well with clients and substantially reduces server cost.
 
-We have also implemented our design in freely available software to
-be used in conjuction with Debian-based distribution of Linux
-software packages. It is currently in use by some users of the
-Debian project's distribution, and so serves as an example of the
-possibilities that exist.
+There are many future avenues toward improving our implementation. Besides 
+evaluating its performance in larger scales, we are particularly interest in further speeding up some of the slower recursive
+DHT requests. We expect to accomplish this by fine tuning the
+parameters of our current system, better exclusion of NATted peers
+from the routing tables, and through the use of STUN \cite{STUN} to
+circumvent the NATs of the 50\% of the peers that have not
+configured port forwarding.
 
-We feel that our P2P software package distribution model is
-well-suited to be used by many free software distributors. We hope
-to convince them to adopt such a model in their distribution, or we
-may port our existing system to some of the other groups for them to
-try.
-
-\textbf{OPTIONAL}: One aspect missing from our model is the removal of old packages
+One aspect missing from our model is the removal of old packages
 from the cache. Since our implementation is still relatively young,
@@
 
-The most significant area of improvement still needed in our sample
-implementation is to further speed up some of the slower recursive
-DHT requests. We hope to accomplish this by further tuning the
-parameters of our current system, better exclusion of NATted peers
-from the routing tables, and through the use of STUN \cite{STUN} to
-circumvent the NATs of the 50\% of the peers that have not
-configured port forwarding.
-
 \bibliographystyle{IEEEtran}

### 2008-08-29 docs/paper/paper.tex
More work on the paper, almost done now.

@@
 \begin{abstract}
-The Internet has become an cost-effective
-vehicle for software development and release, particular in the free software society.
+The Internet has become a cost-effective
+vehicle for software development and release, particular in the free software community.
 Given the free nature of these software, there are often a number of users
@@
 Unfortunately, this application scenario has many unique characteristics, which
-make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
+make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
 a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
 too large to be distributed in its entirety. The packages are also being constantly
-updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
+updated by the loosely-managed developers, and the interest in a particular version of a package can be very
 limited depending on the computer platforms and operating systems used.
 
-In this paper, we propose a novel peer-to-peer assisted distributor design that
+In this paper, we propose a novel peer-to-peer assisted distribution system design that
 addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
@@
 response time reduction, and multi-value extension. They together ensure
-that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
-sheer bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
+that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
+large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
 \end{abstract}
@@
 
-With the widespread penetration of broadband accesses, the Internet has become an cost-effective
+With the widespread penetration of broadband access, the Internet has become a cost-effective
 vehicle for software development and release. This is particularly true for the free software
-society whose developers and users are worldwide distributed and work asynchronously. The ever increasing power of
-modern programming languages, computer platforms, and operating systems has made these software extremely large and complex,
-which are often divided into a huge number of small packages.
+community whose developers and users are distributed worldwide and work asynchronously. The ever increasing power of
+modern programming languages, computer platforms, and operating systems has made this software extremely large and complex,
+though it is often divided into a huge number of small packages.
 Together with their popularity among users,
-an efficient and reliable management and distribution of these software over the Internet has become a challenging task.
-
-The existing distributor for free software are mostly based on the client/server model, e.g.,
+an efficient and reliable management and distribution of these software packages over the Internet has become a challenging task.
+
+The existing distribution for free software is mostly based on the client/server model, e.g.,
 the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
-Given the free nature of these software, there are often a number of users
+Given the free nature of this software, there are often a number of users
 motivated by altruism to help out with the distribution, so as to promote the healthy development
@@
 We thus naturally expect that peer-to-peer distribution can be implemented in
-this context, which scale well with large user bases and can easily explore the network resources made available by
+this context, which will scale well with the currently large user bases and can easily explore the network resources made available by
 the volunteers.
@@
 Unfortunately, this application scenario has many unique characteristics, which
-make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
+make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
 there are too many packages to distribute each individually, but the archive is
 too large to distribute in its entirety. The packages are also being constantly
-updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
+updated by the loosely-managed developers, and the interest in a particular version of a package can be very
 limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
-sequential order used in existing software package distributors. This in turn suppresses interaction with users
+sequential order used in existing software package distributions. This in turn suppresses interaction with users
 given the difficulty in tracking speed and downloading progress.
 
-In this paper, we propose a novel peer-to-peer assisted distributor design that
+In this paper, we propose a novel peer-to-peer assisted distribution system design that
 addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
@@
 response time reduction, and multi-value extension. They together ensure
-that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
-sheer bandwidth requirements of hosting the software.
+that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
+large bandwidth requirements of hosting the software.
 
 \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
-software packages and will also be available in the next release of Ubuntu. We have evaluated our currently deployment  to
+software packages and will also be available in the next release of Ubuntu. We have evaluated our current deployment  to
 determine how effective it is at meeting our goals, and to see what
@@
 implementation for Debian-based distributions in Section~\ref{implementation},
-including an in-depth look at our DHT
-customizations in Section~\ref{custom_dht}. The performance of our implementation is evaluated in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
+including an in-depth look at our system optimization
+in Section~\ref{custom_dht}. The performance of our implementation is evaluated
+in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
 Section~\ref{conclusions} concludes the paper and offers some future directions.
@@
 
-Similar tools have been used for other software packages. CPAN
-distributes software packages for the PERL
+Similar tools have been used for other types of software packages. CPAN
+distributes packaged software for the PERL
 programming language, using SOAP RPC requests to find and download
@@
 motivated by altruism to want to help out with the distribution.
-This is particularly true considering that many of this software is used by
+This is particularly true considering that much of this software is used by
 groups that are staffed mostly, or sometimes completely, by
@@
 of the volunteer community that released the software.
-We also naturally expect that peer-to-peer distribution can be implementation in
-this context, which scale well with large user bases and can easily explore the network resources made available by
+We also naturally expect that peer-to-peer distribution can be implemented in
+this context, which will scale well with the currently large user bases
+and can easily explore the network resources made available by
 the volunteers.
@@
 \centering
-\includegraphics[width=\columnwidth]{apt_p2p_simulation-size_CDF.eps}
+\includegraphics[width=0.80\columnwidth]{apt_p2p_simulation-size_CDF.eps}
 \caption{The CDF of the size of packages in a Debian system, both
@@
 archive consists of 22,298 packages and is approximately 119,000 MB
-in size. Most of the packages are to be installed in any computer environment, but there are
+in size. Many of the packages are to be installed in any computer environment, but there are
 also OS- or architecture-specific packages.
@@
 periodically makes \emph{stable} releases, which are snapshots of
-all the packages in the archive at a certain time, updates are still
+all the packages in the archive at a certain time, later updates are still
 released for security issues or serious bugs.
@@
 \centering
-\includegraphics[width=\columnwidth]{size-quarter.eps}
+\includegraphics[width=0.9\columnwidth]{size-quarter.eps}
 \caption{The amount of data in the 119,000 MB Debian archive that is
@@
 Finally, though there are a large number of packages and a large number of
-users, the interest in a particular version of a package can be very
+users, the interest in a particular package, or version of a package, can be very
 limited. Specifically, there are core packages that every user has to download, but most
@@
 \centering
-\includegraphics[width=\columnwidth]{apt_p2p_popularity-cdf.eps}
+\includegraphics[width=0.80\columnwidth]{apt_p2p_popularity-cdf.eps}
 \caption{The CDF of the popularity of packages in a Debian system.}
@@
 fashion, the package management software also measures the speed
-based on sequential downloading. To offer comparable user experience, it is naturally to expect that
+based on sequential downloading. To offer comparable user experience, it is natural to expect that
 the new peer-to-peer solution be reasonably responsive at retrieving packages,
-preferably enabling sequential downloading order, too.
+preferably in a sequential downloading order too.
 
@@
 Most of the packages are too small, and there are too many packages
-in the entire archive to create individual torrents for each one.
+in the entire archive, to create individual torrents for each one.
 On the other hand, all the packages together are too large to track
@@
 \emph{infohash} identifier and making it a new torrent. This will
-severely fracture the download population, even though peers in the
+severely fracture the download population, since even though peers in the
 new torrent may share 99\% of the packages in common with peers in the
-old torrent.
+old torrent, they will be unable to communicate.
 
@@
 longer critical. Specifically, with altruistic peers
-and all files being available without uploading, incentives to share
+and all files being available to download without uploading, incentives to share
 become a less important issue. Also, the availability of seeds are
-not critical either, as the mirror sites serve in that capacity
-already.
+not critical either, as the serves are already available to serve in
+that capacity.
 
@@
 
-We now present the design of our peer-to-peer assisted distributor for free software package releases and
+We now present the design of our peer-to-peer assisted distribution system for free software package releases and
 updates. A key principle in our design is that the new functionalities implemented in our distributor should be transparent to users,
-thus offering the same experience as using conventional software management systems, despite enhanced efficiency.
+thus offering the same experience as using conventional software management systems, but with enhanced efficiency.
 That said, we assume that the user is still attempting to download packages from a
-server, but the requests will be proxied by our peer-to-peer program. The server is always available and has all of the package files.
+server, but the requests will be proxied by our peer-to-peer program.
+We further assume that the server is always available and has all of the package files.
 In addition, the cryptographic hash of the packages will be available
@@
 \centering
-\includegraphics[width=\columnwidth]{model_simple.eps}
+\includegraphics[width=0.9\columnwidth]{model_simple.eps}
 \caption{The different phases of functionality of our peer-to-peer distribution model.}
@@
 for the package being requested and find its hash (6). This hash can
-then be looked up recursively in an indexing structure (a Distributed Hash Table [] in our implementation) (7), which will return a
+then be looked up recursively in an indexing structure (a Distributed Hash Table in our implementation) (7), which will return a
 list of peers that have the package already (8). The package can
@@
 can not be found in the DHT, the peer can then fallback to
-downloading from the original location (i.e. the network of
-mirrors). The mirrors thus, with no modification to their
-functionality, serve as seeds for the packages in the peer-to-peer
+downloading from the original location (i.e. the server).
+The server thus, with no modification to its
+functionality, serves as a seed for the packages in the peer-to-peer
 system. Any packages that have just been updated, or that are very
 rare, and so don't have any peers available, can always be found on
-the mirror. Once the peer has completed the download from the mirror
+the server. Once the peer has completed the download from the server
 and verified the package, it can then add itself to the DHT as the
 first peer for the new package, so that future requests for the package
-will not need the mirror.
+will not need the server.
 
@@
 can then be pointers to the peers that have the package that hashes
-to that key. A downloading peer can lookup the package hash in the
-DHT and, if it is found, download the file from those peers and
-verify the package with the hash. Once the download is complete, the
-peer will add its entry to the DHT indicating that it now has the
-package.
+to that key.
 
@@
 their archive, and which are also hashed. After retrieving the
-index's hash from the mirror, the index file can be downloaded from
+index's hash from the server, the index file can be downloaded from
 peers and verified. Then the program has access to all the hashes of
@@
 
-\subsection{Implementation Options}
-\label{imp_options}
-
-There are several ways to implement the desired peer-to-peer functionality
-into the existing package management software. The functionality can
-be directly integrated through modifications to the software, though
-this could be difficult as the peer-to-peer functionality should be running
-at all times. This is needed both for efficient lookups and to
-support uploading of already downloaded packages. Unfortunately, the
-package management tools typically only run until the download and
-install request is complete.
-
-Many of the package management software implementations use HTTP
-requests from web servers to download the packages, which makes it
-possible to implement the peer-to-peer aspect as an almost standard HTTP
-caching proxy. This proxy will run as a daemon in the background,
-listening for requests from the package management tool for package
-files, as well as serving (uploading) cached package to other peers.
-It will get uncached requests first from the peer-to-peer system, or
-fall back to the normal HTTP request from a server should it not
-be found. For methods that don't use HTTP requests, other types of
-proxies may also be possible.
-
-\subsection{Downloading From Peers}
+% \subsection{Implementation Options}
+% \label{imp_options}
+% 
+% There are several ways to implement the desired peer-to-peer functionality
+% into the existing package management software. The functionality can
+% be directly integrated through modifications to the software, though
+% this could be difficult as the peer-to-peer functionality should be running
+% at all times. This is needed both for efficient lookups and to
+% support uploading of already downloaded packages. Unfortunately, the
+% package management tools typically only run until the download and
+% install request is complete.
+% 
+% Many of the package management software implementations use HTTP
+% requests from web servers to download the packages, which makes it
+% possible to implement the peer-to-peer aspect as an almost standard HTTP
+% caching proxy. This proxy will run as a daemon in the background,
+% listening for requests from the package management tool for package
+% files, as well as serving (uploading) cached package to other peers.
+% It will get uncached requests first from the peer-to-peer system, or
+% fall back to the normal HTTP request from a server should it not
+% be found. For methods that don't use HTTP requests, other types of
+% proxies may also be possible.
+
+\subsection{Peer Downloading Protocol}
 \label{downloading}
@@
 protocol that is similar to the protocol used to fetch packages from
-the distributor's servers. This simplifies the peer-to-peer program, as it
-can then treat peers and mirrors almost identically when requesting
-packages. In fact, the mirrors can be used when there are only a few
+the distributor's server. This simplifies the peer-to-peer program, as it
+can then treat peers and the server almost identically when requesting
+packages. In fact, the server can be used when there are only a few
 slow peers available for a file to help speed up the download
@@
 \emph{torrent string} containing the peer's hashes of the pieces of
-the larger files. These piece hashes are compared ahead of time
+the larger files (similar to (15) in Phase 3 of Figure~\ref{model}).
+These piece hashes will be retrieved and compared ahead of time by
+the downloading peer ((9,10) in Phase 2 of Figure~\ref{model})
 to determine which peers have the same piece hashes (they all
@@
 are all \emph{bencoded} in the same way as BitTorrent's
-\texttt{.torrent} files. Khashmir uses the high-level Twisted
-event-driven networking engine \cite{twisted}, so we also use
-Twisted in our sample implementation for all other networking needs.
+\texttt{.torrent} files.
+% Khashmir uses the high-level Twisted
+% event-driven networking engine \cite{twisted}, so we also use
+% Twisted in our sample implementation for all other networking needs.
 More details of this customized DHT can be found below in
@@
 
-\section{Customized DHT / System Optimization}
+\section{System Optimization}
 \label{custom_dht}
 
-A large contribution of our work is in the customization and use of
+Another contribution of our work is in the customization and use of
 a Distributed Hash Table (DHT). Although our DHT is based on
@@
 
-\subsection{Kademlia DHT Background}
+\subsection{DHT Details}
 \label{dht}
 
-DHT's operate by storing (\emph{key}, \emph{value}) pairs in a
-distributed fashion such that no node will, on average, store more
+DHTs operate by storing (\emph{key}, \emph{value}) pairs in a
+distributed fashion such that no node will, on average, store more or have to work harder
 than any other node. They support two primitive operations:
@@
 that was previously stored with that key. These operations are
-recursive, as each node does not know about all the other nodes in a
+recursive, as each node does not know about all the other nodes in the
 DHT, and so must recursively search for the correct node to put to
@@
 
-The Kademlia DHT, like most other DHTs, assigns IDs to peers from
+The Kademlia DHT, like most other DHTs, assigns IDs to peers randomly from
 the same space that is used for keys. The peers with IDs closest to
-the desired key will then store the values for that key. Lookups are
-recursive, as nodes are queried in each step that are exponentially
-closer to the key than in the previous step.
-
-Nodes in a Kademlia system support four primitive requests.
+the desired key will then store the values for that key.
+Nodes support four primitive requests.
 \texttt{ping} will cause a peer to return nothing, and is only used
-to determine if a node is still alive, while \texttt{store} tells a
+to determine if a node is still alive. \texttt{store} tells a
 node to store a value associated with a given key. The most
@@
 they know about that are closest to the key, allowing the querying
-node to quickly traverse the DHT to find the nodes close to the
+node to quickly traverse the DHT to find the nodes closest to the
 desired key. The only difference between \texttt{find\_node} and \texttt{find\_value} is that the
@@
 \centering
-\includegraphics[width=\columnwidth]{apt_p2p_improvements-find_value.eps}
+\includegraphics[width=0.80\columnwidth]{apt_p2p_improvements-find_value.eps}
 \caption{The distribution of average response times PlanetLab nodes
@@
 
-To test our changes during development, we ran the customized DHT
+To test our changes during development, we ran our customized DHT
 for several hours after each major change on over 300 PlanetLab nodes
@@
 in Figure~\ref{improvements}. Each improvement successfully reduced
-the response time, for a total reduction of more than half. The
+the response time, for a total reduction of more than 50\%. The
 final distribution is also narrower, as the improvements make the
@@
 
-\subsection{Extension with Multiple Values}
+\subsection{Multiple Values Extension}
 \label{multiple_values}
@@
 
+% \subsection{\textbf{OPTIONAL}: BitTorrent's Improvements}
+% \label{bittorrent_dht}
+% 
+% In the several years that some BitTorrent clients have been using a
+% Kademlia-based DHT for tracker-less operation, the developers have
+% made many enhancements which we can take advantage of. One of the
+% most important is a security feature added to stop malicious nodes
+% from subscribing other nodes as downloaders. When a node issues a
+% request to another node to store its download info for that key, it
+% must include a \emph{token} returned by the storing node in a recent
+% \texttt{find\_node} request. The token is created by hashing the IP
+% address of the requesting peer with a temporary secret that expires
+% after several minutes. This prevents the requesting peer from faking
+% its IP address in the store request, since it must first receive a
+% response from a \texttt{find\_node} on that IP.
+% 
+% We also made some BitTorrent-inspired changes to the parameters of
+% the DHT originally specified by the authors of Kademlia. Since, as
+% we will show later, peers stay online in our system for much longer
+% periods of time, we reduced Kademlia's \emph{k} value from 20 to 8.
+% The value is supposed to be large enough such that any given
+% \emph{k} nodes are unlikely to fail within an hour of each other,
+% which is very unlikely in our system given the long uptimes of
+% nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
+% We also increased the number of concurrent outstanding
+% requests allowed from 3 to 6 to speed up the recursive key finding
+% processes.
+% 
+% \subsection{\textbf{OPTIONAL}: Other Changes}
+% \label{other_changes}
+% 
+% In addition, we have allowed peers to store values in the DHT, even
+% if the hash they are using is not the correct length. Most of the
+% keys used in the DHT are based on the SHA1 hash, and so they are 20
+% bytes in length. However, some files in the targetted Debian package
+% system are only hashed using the MD5 algorithm, and so the hashes
+% retrieved from the server are 16 bytes in length. The DHT will still
+% store values using these keys by padding the end of them with zeroes
+% to the proper length, as the loss of uniqueness from the 4 less
+% bytes is not an issue. Also, the DHT will happily store hashes that
+% are longer than 20 bytes, such as those from the 32 byte SHA256
+% algorithm, by truncating them to the correct length. Since the
+% requesting peer will have the full length of the hash, this will not
+% affect its attempt to verify the downloaded file.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-\includegraphics[width=\columnwidth]{AptP2PWalker-peers.eps}
+\includegraphics[width=0.80\columnwidth]{AptP2PWalker-peers.eps}
 \caption{The number of peers found in the system, and how many are
@@
 \centering
-\includegraphics[width=\columnwidth]{AptP2PDuration-peers.eps}
+\includegraphics[width=0.80\columnwidth]{AptP2PDuration-peers.eps}
 \caption{The CDF of how long an average session will last.}
@@
 
+% \begin{figure}
+% \centering
+% \includegraphics[width=0.80\columnwidth]{AptP2PDuration-ind_peers.eps}
+% \caption{\textbf{OPTIONAL}: The CDF of the average time individual peers stay in the
+% system.}
+% \label{duration_ind_peers}
+% \end{figure}
+% 
+% \textbf{OPTIONAL}: We also examined the average time each individual peer spends in the
+% system. Figure~\ref{duration_peers} shows the cumulative
+% distribution of how long each individual peer remains in the system.
+% Here we see that 50\% of peers have average stays in the system
+% longer than 10 hours.
+
 \begin{figure}
 \centering
-\includegraphics[width=\columnwidth]{AptP2PDuration-ind_peers.eps}
-\caption{\textbf{OPTIONAL}: The CDF of the average time individual peers stay in the
-system.}
-\label{duration_ind_peers}
-\end{figure}
-
-\textbf{OPTIONAL}: We also examined the average time each individual peer spends in the
-system. Figure~\ref{duration_peers} shows the cumulative
-distribution of how long each individual peer remains in the system.
-Here we see that 50\% of peers have average stays in the system
-longer than 10 hours.
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{AptP2PDuration-online_1.eps}
+\includegraphics[width=0.80\columnwidth]{AptP2PDuration-online_1.eps}
 \caption{The fraction of peers that, given their current duration in
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{AptP2PDuration-online_6.eps}
-\caption{\textbf{OPTIONAL}: The fraction of peers that, given their current duration in
-the system, will stay online for another 6 hours.}
-\label{duration_online_6}
-\end{figure}
-
-\textbf{OPTIONAL}: Since our peers are much longer-lived than other peer-to-peer systems, we
-also looked at the fraction of peers that stay online for another 6
-hours. Figure~\ref{duration_online_6} shows that over 60\% of peers
-that are online for 10 hours will stay online for another 6.
-However, we see an interesting decrease in this fraction between 8
-and 12 hours, which can also be seen in
-Figure~\ref{duration_online_1}. We believe this to be due to desktop
-users, who regularly turn off their computers at night.
+% \begin{figure}
+% \centering
+% \includegraphics[width=0.80\columnwidth]{AptP2PDuration-online_6.eps}
+% \caption{\textbf{OPTIONAL}: The fraction of peers that, given their current duration in
+% the system, will stay online for another 6 hours.}
+% \label{duration_online_6}
+% \end{figure}
+% 
+% \textbf{OPTIONAL}: Since our peers are much longer-lived than other peer-to-peer systems, we
+% also looked at the fraction of peers that stay online for another 6
+% hours. Figure~\ref{duration_online_6} shows that over 60\% of peers
+% that are online for 10 hours will stay online for another 6.
+% However, we see an interesting decrease in this fraction between 8
+% and 12 hours, which can also be seen in
+% Figure~\ref{duration_online_1}. We believe this to be due to desktop
+% users, who regularly turn off their computers at night.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{AptP2PDownloaded-peers.eps}
+\includegraphics[width=0.80\columnwidth]{AptP2PDownloaded-peers.eps}
 \caption{The number of peers that were contacted to determine their
@@
 \centering
-\includegraphics[width=\columnwidth]{AptP2PDownloaded-bw.eps}
-\caption{\textbf{OPTIONAL}: The bandwidth of data that the contacted peers have
+\includegraphics[width=0.80\columnwidth]{AptP2PDownloaded-bw.eps}
+\caption{The bandwidth of data that the contacted peers have
 downloaded and uploaded.}
@@
 that \texttt{apt-p2p} has already saved the mirrors 15 GB of
-bandwidth, or 1 GB per day.
+bandwidth, or 1 GB per day. Considering the current small number of
+users this savings is quite large, and is expected to grow
+considerably as more users participate in the P2P system.
 
@@
 a download can occur, is taking 17 seconds on average. This
-indicates that, on average, requests are experiencing almost 2
+indicates that, on average, requests are experiencing almost 2 full
 stalls while waiting for the 9 second timeouts to occur on
-unresponsive peers. This is longer than our target of 10 seconds,
+unresponsive peers. This time is longer than our target of 10 seconds,
 although it will only lead to a slight average delay in downloading
@@
 even while not downloading, would overwhelm some peers' network
-connections. However, we found that peers are using 200 to 300 bytes
-per second of bandwidth in servicing the DHT. These numbers are
+connections. However, we found that peers are using 200 to 300 bytes/sec
+of bandwidth in servicing the DHT. These numbers are
 small enough to not affect any other network services the peer would
@@
 
-\textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
-Kenosis \cite{kenosis} is a peer-to-peer Remote Procedure Call
-client also based on the Kademlia DHT, but it is meant as a peer-to-peer
-primitive system on which other tools can be built, and so it has no
-file sharing functionality at all. Many have used a DHT as a drop-in
-replacement for the tracker in a BitTorrent system
-\cite{bittorrent-dht, azureus-dht}, but such systems only use the
-DHT to find peers for the same torrent, so the file sharing uses
-traditional BitTorrent and so is not ideal for the reasons listed
-in Section~\ref{bittorrent}.
+% \textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
+% Kenosis \cite{kenosis} is a peer-to-peer Remote Procedure Call
+% client also based on the Kademlia DHT, but it is meant as a peer-to-peer
+% primitive system on which other tools can be built, and so it has no
+% file sharing functionality at all. Many have used a DHT as a drop-in
+% replacement for the tracker in a BitTorrent system
+% \cite{bittorrent-dht, azureus-dht}, but such systems only use the
+% DHT to find peers for the same torrent, so the file sharing uses
+% traditional BitTorrent and so is not ideal for the reasons listed
+% in Section~\ref{bittorrent}.
 

### 2008-09-28 docs/paper/paper.tex
Final version of paper.

@@
 that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
-large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
+large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics over the Internet.
 \end{abstract}
@@
 With the widespread penetration of broadband access, the Internet has become a cost-effective
-vehicle for software development and release. This is particularly true for the free software
+vehicle for software development and release \cite{feller2000fao}. This is particularly true for the free software
 community whose developers and users are distributed worldwide and work asynchronously. The ever increasing power of
@@
 Together with their popularity among users,
-an efficient and reliable management and distribution of these software packages over the Internet has become a challenging task.
+an efficient and reliable management and distribution of these software packages over the Internet has become a challenging task \cite{ubuntu-blueprint}.
 
 The existing distribution for free software is mostly based on the client/server model, e.g.,
-the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
-Given the free nature of this software, there are often a number of users
+the Advanced Package Tool (\texttt{apt}) for Linux \cite{apt}, which suffers from the well-known bottleneck problem.
+Given the free nature of these software, there are often a number of users
 motivated by altruism to help out with the distribution, so as to promote the healthy development
@@
 We thus naturally expect that peer-to-peer distribution can be implemented in
-this context, which will scale well with the currently large user bases and can easily explore the network resources made available by
+this context, which will scale well with the currently large user bases and can easily explore the resources made available by
 the volunteers.
@@
 
-In the free software society, there are a large number of groups using the Internet to
-collaboratively develop and release their software. The ever increasing power of
-modern programming languages and operating systems has made this software, like commercial software, extremely large and complex, though it is often
-distributed in many small units (packages). Together with their popularity among users,
-an efficient and reliable management and distribution of these packages over the Internet has become a daunting task. In this section, we offer concrete examples illustrating the
+In the free software community, there are a large number of groups using the Internet to
+collaboratively develop and release their software. Efficient and reliable management and distribution 
+of these software packages over the Internet thus has become a critical task. In this section, we offer concrete examples illustrating the
 unique challenges in this context.
@@
 periodically makes \emph{stable} releases, which are snapshots of
-all the packages in the archive at a certain time, later updates are still
+all the packages in the archive at a certain time, many later updates are still
 released for security issues or serious bugs.
@@
 updated with new versions of packages. This frequency is much higher than
-that of most commercial software, mainly because much of free software is
+that of most commercial software \cite{microsoft-update}, mainly because much of free software is
 developed in a loosely managed environment of developers working
@@
 
-Finally, though there are a large number of packages and a large number of
+Though there are a large number of packages and a large number of
 users, the interest in a particular package, or version of a package, can be very
@@
 packages would fall in the category of optional or extra, and so are
-interesting to only a limited number of people.
+interesting to only a limited number of users.
 
@@
 
-Given the relatively long time for software package downloading, existing package management systems generally display
+Finally, given the relatively long time for software package downloading, existing package management systems generally display
 some kind of indication of speed and completeness for users to
@@
 
-\subsection{Why BitTorrent Doesn't Work Well}
+\subsection{Why BitTorrent Does Not Work Well}
 \label{bittorrent}
@@
 
+\subsection{System Overview}
 Our model for using peer-to-peer to enhance package distribution is shown in
@@
 for the package being requested and find its hash (6). This hash can
-then be looked up recursively in an indexing structure (a Distributed Hash Table in our implementation) (7), which will return a
+then be looked up recursively in an indexing structure (a Distributed Hash Table, or DHT \cite{kademlia}, in our implementation) (7), which will return a
 list of peers that have the package already (8). The package can
@@
 system. Any packages that have just been updated, or that are very
-rare, and so don't have any peers available, can always be found on
+rare, and so do not have any peers available, can always be found on
 the server. Once the peer has completed the download from the server
@@
 interest in a large number of packages undergoing constant updating
-is well suited to the functionality provided by a Distributed Hash
-Table (DHT). DHTs require unique keys to store and retrieve strings
+is well suited to the functionality provided by a DHT. A DHT requires unique keys to store and retrieve strings
 of data, for which the cryptographic hashes used by these package
@@
 % fall back to the normal HTTP request from a server should it not
-% be found. For methods that don't use HTTP requests, other types of
+% be found. For methods that do not use HTTP requests, other types of
 % proxies may also be possible.
@@
 currently connected to it. This allows us to analyze many aspects of
-our implementation.
+our implementation in the real Internet environment.
 
@@
 to become peers in the system, sharing the uploading load equally
-amongst all, but doesn't require any user to download files they
+amongst all, but does not require any user to download files they
 would not otherwise need.
@@
 the lookup time, but their system is more suited to its intended use
-as a distributed file server. It doesn't make use of authoritative
+as a distributed file server. It does not make use of authoritative
 copies of the original files, allowing instead any users in the

### 2009-01-19 docs/paper/paper.tex
WIP on final version of accepted INFOCOM paper.

@@
 %% INFOCOM addition:
-\makeatletter
-\def\ps@headings{%
-\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
-\def\@evenhead{\scriptsize\thepage \hfil\leftmark\mbox{}}%
-\def\@oddfoot{}%
-\def\@evenfoot{}}
-\makeatother
-\pagestyle{headings}
+% \makeatletter
+% \def\ps@headings{%
+% \def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
+% \def\@evenhead{\scriptsize\thepage \hfil\leftmark\mbox{}}%
+% \def\@oddfoot{}%
+% \def\@evenfoot{}}
+% \makeatother
+% \pagestyle{headings}
 
@@
 and all files being available to download without uploading, incentives to share
-become a less important issue. Also, the availability of seeds are
-not critical either, as the serves are already available to serve in
+become a less important issue. Also, the availability of seeders is
+not critical either, as the servers are already available to serve in
 that capacity.
@@
 hours, and 20\% last longer than 10 hours. These connections are
-much longer than those reported by Saroiu et. al. \cite{saroiu2001}
+much longer than those reported by Saroiu et al. \cite{saroiu2001}
 for other peer-to-peer systems, which had 50\% of Napster and Gnutella
@@
 \includegraphics[width=0.80\columnwidth]{AptP2PDownloaded-bw.eps}
-\caption{The bandwidth of data that the contacted peers have
+\caption{The bandwidth of data (total number of bytes) that the contacted peers have
 downloaded and uploaded.}
@@
 content distribution network (CDN) using peer-to-peer techniques.
-Freedman et. al. developed Coral \cite{coral} using a distributed
+Freedman et al. developed Coral \cite{coral} using a distributed
 \emph{sloppy} hash table to speed request times. Pierre and van
@@
 replica servers, which in turn draw their content from the original
-location (or a backup). Shah et. al. \cite{shah08} analyze an
+location (or a backup). Shah et al. \cite{shah08} analyze an
 existing software delivery system and use the results to design a
@@
 
-The most similar works to ours are by Shah et. al. \cite{shah08} and
-Shark by Annapureddy et. al. \cite{shark}.
+The most similar works to ours are by Shah et al. \cite{shah08} and
+Shark by Annapureddy et al. \cite{shark}.
 Shah's system, in addition to the drawbacks mentioned above,

### 2009-01-21 docs/paper/paper.tex
Final version of INFOCOM paper.

@@
 vehicle for software development and release, particular in the free software community.
-Given the free nature of these software, there are often a number of users
+Given the free nature of this software, there are often a number of users
 motivated by altruism to help out with the distribution, so as to promote the healthy development
-of this voluntary society. It is thus naturally expected that peer-to-peer distribution can be implemented,
-which scale well with large user bases and can easily explore the network resources made available by
+of this voluntary society. It is thus naturally expected that a peer-to-peer distribution can be implemented,
+which will scale well with large user bases, and can easily explore the network resources made available by
 the volunteers.
@@
 Unfortunately, this application scenario has many unique characteristics, which
-make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
-a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
+make a straightforward adoption of existing peer-to-peer systems for file sharing (such as BitTorrent) suboptimal. In particular,
+a software release often consists of a large number of packages, which are difficult to distribute individually, but the archive is
 too large to be distributed in its entirety. The packages are also being constantly
@@
 Our design leads to \texttt{apt-p2p}, a practical implementation that extends the popular \texttt{apt} distributor.
-\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
-software packages and will also be available in the next release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
+\texttt{apt-p2p}  has been used in conjunction with Debian-based distribution of Linux
+software packages and is also available in the latest release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
 response time reduction, and multi-value extension. They together ensure
 that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
-large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics over the Internet.
+large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics gathered over the Internet.
 \end{abstract}
@@
 the Advanced Package Tool (\texttt{apt}) for Linux \cite{apt}, which suffers from the well-known bottleneck problem.
-Given the free nature of these software, there are often a number of users
+Given the free nature of this software, there are often a number of users
 motivated by altruism to help out with the distribution, so as to promote the healthy development
@@
 Unfortunately, this application scenario has many unique characteristics, which
-make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
+make a straightforward adoption of existing peer-to-peer systems for file sharing (such as BitTorrent) suboptimal. In particular,
 there are too many packages to distribute each individually, but the archive is
@@
 updated by the loosely-managed developers, and the interest in a particular version of a package can be very
-limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
-sequential order used in existing software package distributions. This in turn suppresses interaction with users
+limited. They together make it very difficult to efficiently create and manage torrents and trackers. The random downloading nature of BitTorrent-like systems is also different from the
+sequential order used in existing software package distributors. This in turn suppresses interaction with users
 given the difficulty in tracking speed and downloading progress.
@@
 for software packages.
-Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian \footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
+Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian\footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
 distribution system.  We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
@@
 
-\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
-software packages and will also be available in the next release of Ubuntu. We have evaluated our current deployment  to
+\texttt{apt-p2p}  has been used in conjunction with the Debian-based distribution of Linux
+software packages and is also available in the latest release of Ubuntu. We have evaluated our current deployment  to
 determine how effective it is at meeting our goals, and to see what
@@
 
-\section{Background and Motivations}
+\section{Background and Motivation}
 \label{situation}
@@
 
-Direct web downloading is also a common way, often coupled with a hash
+Direct web downloading by users is also common, often coupled with a hash
 verification file to be downloaded next to the desired
@@
 While it seems straightforward to use an existing peer-to-peer file sharing tool like BitTorrent for
-free software package distribution, there are indeed a series of new challenges in this unique scenario:
+this free software package distribution, there are indeed a series of new challenges in this unique scenario:
 
@@
 between the user and the server (4). It will therefore also have
-available the index files containing the cryptographic hashes all
+available the index files containing the cryptographic hashes of all
 packages. Later, in Phase~2, upon receiving a request from the user
@@
 then be looked up recursively in an indexing structure (a Distributed Hash Table, or DHT \cite{kademlia}, in our implementation) (7), which will return a
-list of peers that have the package already (8). The package can
-then be downloaded from the peers (11,12), it can be verified using
+list of peers that have the package already (8). Then, in Phase~3, the package
+can be downloaded from the peers (11,12), it can be verified using
 the hash (13), and if valid can be returned to the user (14). The
-current nodes location is also added to the DHT for that hash (15),
+current node's location is also added to the DHT for that hash (15),
 as it is now a source for others to download from.
@@
 functionality, serves as a seed for the packages in the peer-to-peer
-system. Any packages that have just been updated, or that are very
-rare, and so do not have any peers available, can always be found on
+system. Any packages that have just been updated or that are very
+rare, and so do not yet have any peers available, can always be found on
 the server. Once the peer has completed the download from the server
@@
 first peer for the new package, so that future requests for the package
-will not need the server.
+will not need to use the server.
 
@@
 until the hash is looked up in the DHT, so a hash must first come
-from a trusted source (i.e. the distributors server). Most distributors use index
+from a trusted source (i.e. the distributor's server). Most distributors use index
 files that contain hashes for a large number of the packages in
 their archive, and which are also hashed. After retrieving the
-index's hash from the server, the index file can be downloaded from
+index file's hash from the server, the index file can also be downloaded from
 peers and verified. Then the program has access to all the hashes of
@@
 implement trackerless operation. The communication is all handled by
-UDP messages, and RPC (remote procedure call) requests and responses
+UDP messages, and RPC (remote procedure call) requests and responses between nodes
 are all \emph{bencoded} in the same way as BitTorrent's
@@
 peers identified by lookups in the DHT to have the desired file.
-Requests for a package are made using the package's hash, properly
-encoded, as the URL to request from the peer. The HTTP server used
+Requests for a package are made using the package's hash (properly
+encoded) as the URL to request from the peer. The HTTP server used
 for the proxy also doubles as the server listening for requests for
@@
 
-The Kademlia DHT, like most other DHTs, assigns IDs to peers randomly from
-the same space that is used for keys. The peers with IDs closest to
+The Kademlia DHT, like most other DHTs, assigns Ids to peers randomly from
+the same space that is used for keys. The peers with Ids closest to
 the desired key will then store the values for that key.
@@
 of the entire concatenated length of the torrent string. If the
-torrent string is short enough to store in the DHT (i.e. less than
+torrent string is short enough to store separately in the DHT (i.e. less than
 1472 bytes, or about 70 pieces for the SHA1 hash), then a lookup of
@@
 that will require a separate lookup in the DHT for the longer
-torrent string as they require 5 to 70 pieces. Finally, there are
+torrent string, as they require 5 to 70 pieces. Finally, there are
 only 62 packages that require more than 70 pieces, and so will
@@
 This has the effect of leap-frogging unresponsive nodes and
-focussing attention on the nodes that do respond. We will also
+focussing attention on the closer nodes that do respond. We will also
 prematurely abort a query while there are still oustanding requests,
@@
 was called \texttt{find\_value} and worked similarly to
-\texttt{find\_node}, iteratively finding nodes with ID's closer to
+\texttt{find\_node}, iteratively finding nodes with Id's closer to
 the desired key. However, if a node had a value stored associated
@@
 % \emph{k} nodes are unlikely to fail within an hour of each other,
-% which is very unlikely in our system given the long uptimes of
+% which is very unlikely in our system given the long uptime of
 % nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
@@
 distribution system has been available to all Debian users since May
-3rd, 2008 \cite{apt-p2p-debian}, and will also be available in the
-next release of Ubuntu \cite{apt-p2p-ubuntu}. We have since created
+3rd, 2008 \cite{apt-p2p-debian}, and is also available in the
+latest release of Ubuntu \cite{apt-p2p-ubuntu}. We created
 a \emph{walker} that will navigate the DHT and find all the peers
@@
 
-We first began analyzing the DHT on June 24th, and continue to this
-day, giving us 2 months of data so far. Figure~\ref{walker_peers}
+We first began analyzing the DHT on June 24th, 2008, and continued
+until we had gathered almost 2 months of data. Figure~\ref{walker_peers}
 shows the number of peers we have seen in the DHT during this time.
@@
 requesting peer's IP and port, so that the peer can determine what
-its oustide IP address is and whether port translation is being
+its outside IP address is and whether port translation is being
 used. In the future, we hope to add functionality similar to STUN
@@
 it uses the same port for both its DHT (UDP) requests and download
-(TCP) requests (which is also the default configuration parameter).
+(TCP) requests (which is also the default configuration behavior).
 
@@
 than the existing peer-to-peer systems for file sharing. To this end, we have
-present \texttt{apt-p2p}, a novel peer-to-peer distributor that sits between
+presented \texttt{apt-p2p}, a novel peer-to-peer distributor that sits between
 client and server, providing efficient and transparent downloading and updating services
 for software packages. We have addressed the key design issues in \texttt{apt-p2p}, including DHT customization,
-response time reduction, and multi-value extension. \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
-software packages and will also be available in the next release of Ubuntu. Existing real user statistics
+response time reduction, and multi-value extension. \texttt{apt-p2p}  has been used in conjunction with Debian-based distribution of Linux
+software packages and is also available in the latest release of Ubuntu. Existing real user statistics
 have suggested that it interacts well with clients and substantially reduces server cost.
